### 2024-07-29 documents/paper.tex
Clean up template.

\documentclass[11pt, a4paper, leqno]{article}
\usepackage{a4wide}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{float, afterpage, rotating, graphicx}
\usepackage{epstopdf}
\usepackage{longtable, booktabs, tabularx}
\usepackage{fancyvrb, moreverb, relsize}
\usepackage{eurosym, calc}
% \usepackage{chngcntr}
\usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
\usepackage{caption}
\usepackage{mdwlist}
\usepackage{xfrac}
\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{minibox}
% \usepackage{pdf14} % Enable for Manuscriptcentral -- can't handle pdf 1.5
% \usepackage{endfloat} % Enable to move tables / figures to the end. Useful for some
% submissions.

\usepackage[
    natbib=true,
    bibencoding=inputenc,
    bibstyle=authoryear-ibid,
    citestyle=authoryear-comp,
    maxcitenames=3,
    maxbibnames=10,
    useprefix=false,
    sortcites=true,
    backend=biber
]{biblatex}
\AtBeginDocument{\toggletrue{blx@useprefix}}
\AtBeginBibliography{\togglefalse{blx@useprefix}}
\setlength{\bibitemsep}{1.5ex}
\addbibresource{refs.bib}

\usepackage[unicode=true]{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    anchorcolor=black,
    citecolor=NavyBlue,
    filecolor=black,
    menucolor=black,
    runcolor=black,
    urlcolor=NavyBlue
}

\widowpenalty=10000
\clubpenalty=10000

\setlength{\parskip}{1ex}
\setlength{\parindent}{0ex}
\setstretch{1.5}

\begin{document}

\title{Topics Causal Inference}
\thanks{EXAMPLE UNIVERSITY. Email: janedoe[at]example-university.de}

\author{Julian Budde}

\date{
    {\bf Preliminary -- please do not quote}
    \\[1ex]
    \today
}

\maketitle

\begin{abstract}
    Some abstract here.
\end{abstract}

\clearpage

\section{Introduction} % (fold)
\label{sec:introduction}

If you are using this template, please cite this item from the references:
\citet{GaudeckerEconProjectTemplates}.

The data set for the template project is taken from
\url{https://www.stem.org.uk/resources/elibrary/resource/28452/large-datasets-stats4schools}.
It contains data on smoking habits in the UK, with 1691 observations and 12 variables.
We consider only 4 of the 12 features for the prediction of the variable
\texttt{smoking}: \texttt{marital\_status}, \texttt{highest\_qualification},
\texttt{gender} and \texttt{age}. We model the dependence using a Logistic model. All
numerical features are included linearly, while categorical features are expanded into
dummy variables. Figures below illustrate the model predictions over the lifetime. You
will find one figure and one estimation summary table for each installed programming
language.

\begin{figure}[H]

    \centering
    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_python}

    \caption{Model predictions of the smoking probability over the lifetime. Each
        colored line represents a case where marital status is fixed to one of the
        values present in the data set.}
    \label{fig:predictions}

\end{figure}

\begin{table}[!h]
    \input{../tables/estimation_results.tex}
    \label{tab:summary}
    \caption{Estimation results of the linear Logistic regression.}
\end{table}

\begin{figure}[H]

    \centering
    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_r}

    \caption{\emph{R:} Model predictions of the smoking probability over
        the lifetime. Each colored line represents a case where marital status
        is fixed to one of the values present in the data set.}
    \label{fig:r-predictions}

\end{figure}

section introduction (end)

\setstretch{1}
\printbibliography
\setstretch{1.5}

% \appendix

% The chngctr package is needed for the following lines.
% \counterwithin{table}{section}
% \counterwithin{figure}{section}

\end{document}

### 2024-07-29 documents/presentation.tex
Clean up template.

--- 2024-07-29 documents/paper.tex
@@
-\documentclass[11pt, a4paper, leqno]{article}
-\usepackage{a4wide}
+\documentclass[11pt, aspectratio=169]{beamer}
+% \documentclass[11pt,handout]{beamer}
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
+\usepackage{textcomp}
 \usepackage{float, afterpage, rotating, graphicx}
@@
 \usepackage{eurosym, calc}
-% \usepackage{chngcntr}
 \usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
-\usepackage{caption}
-\usepackage{mdwlist}
-\usepackage{xfrac}
-\usepackage{setspace}
-\usepackage[dvipsnames]{xcolor}
-\usepackage{subcaption}
-\usepackage{minibox}
-% \usepackage{pdf14} % Enable for Manuscriptcentral -- can't handle pdf 1.5
-% \usepackage{endfloat} % Enable to move tables / figures to the end. Useful for some
-% submissions.
-
 \usepackage[
@@
 
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
-    colorlinks=true,
-    linkcolor=black,
-    anchorcolor=black,
-    citecolor=NavyBlue,
-    filecolor=black,
-    menucolor=black,
-    runcolor=black,
-    urlcolor=NavyBlue
-}
+\hypersetup{colorlinks=true, linkcolor=black, anchorcolor=black, citecolor=black, filecolor=black, menucolor=black, runcolor=black, urlcolor=black}
 
-\widowpenalty=10000
-\clubpenalty=10000
-
-\setlength{\parskip}{1ex}
-\setlength{\parindent}{0ex}
-\setstretch{1.5}
+\setbeamertemplate{footline}[frame number]
+\setbeamertemplate{navigation symbols}{}
+\setbeamertemplate{frametitle}{\centering\vspace{1ex}\insertframetitle\par}
 
@@
 \title{Topics Causal Inference}
-\thanks{EXAMPLE UNIVERSITY. Email: janedoe[at]example-university.de}
 
-\author{Julian Budde}
-
-\date{
-    {\bf Preliminary -- please do not quote}
-    \\[1ex]
-    \today
+\author[Julian Budde]
+{
+{\bf Julian Budde}\\
+{\small EXAMPLE UNIVERSITY }\\[1ex]
 }
 
-\maketitle
+\begin{frame}
+    \titlepage
+    \note{~}
+\end{frame}
 
-\begin{abstract}
-    Some abstract here.
-\end{abstract}
+\begin{frame}[t]
+    \frametitle{heading here}
+    \begin{itemize}
+        \item<+-> Please cite this template as: \citet{GaudeckerEconProjectTemplates}
+        \item<+-> Example data is taken from \url{https://www.stem.org.uk/resources/elibrary/resource/28452/large-datasets-stats4schools}
+    \end{itemize}
+    \note{~}
+\end{frame}
 
-\clearpage
-
-\section{Introduction} % (fold)
-\label{sec:introduction}
-
-If you are using this template, please cite this item from the references:
-\citet{GaudeckerEconProjectTemplates}.
-
-The data set for the template project is taken from
-\url{https://www.stem.org.uk/resources/elibrary/resource/28452/large-datasets-stats4schools}.
-It contains data on smoking habits in the UK, with 1691 observations and 12 variables.
-We consider only 4 of the 12 features for the prediction of the variable
-\texttt{smoking}: \texttt{marital\_status}, \texttt{highest\_qualification},
-\texttt{gender} and \texttt{age}. We model the dependence using a Logistic model. All
-numerical features are included linearly, while categorical features are expanded into
-dummy variables. Figures below illustrate the model predictions over the lifetime. You
-will find one figure and one estimation summary table for each installed programming
-language.
-
+\begin{frame}[t]
     \begin{figure}[H]
@@
         \centering
-    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_python}
+        \includegraphics[width=0.5\textwidth]{../figures/smoking_by_marital_status_using_python}
 
-    \caption{Model predictions of the smoking probability over the lifetime. Each
-        colored line represents a case where marital status is fixed to one of the
-        values present in the data set.}
-    \label{fig:predictions}
+        \caption{\emph{Python:} Model predictions of the smoking probability over the
+            lifetime. Each colored line represents a case where marital status is fixed to
+            one of the values present in the data set.}
+        \label{fig:python-predictions}
 
     \end{figure}
+\end{frame}
 
-\begin{table}[!h]
-    \input{../tables/estimation_results.tex}
-    \label{tab:summary}
-    \caption{Estimation results of the linear Logistic regression.}
-\end{table}
-
+\begin{frame}[t]
     \begin{figure}[H]
@@
         \centering
-    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_r}
+        \includegraphics[width=0.5\textwidth]{../figures/smoking_by_marital_status_using_r}
 
-    \caption{\emph{R:} Model predictions of the smoking probability over
-        the lifetime. Each colored line represents a case where marital status
-        is fixed to one of the values present in the data set.}
+        \caption{\emph{R:} Model predictions of the smoking probability over the
+        lifetime. Each colored line represents a case where marital status is fixed to
+        one of the values present in the data set.}
         \label{fig:r-predictions}
@@
     \end{figure}
+\end{frame}
 
-section introduction (end)
+% Print black screen only in presentation mode for finishing up.
+\mode<beamer> {
+    \beamersetaveragebackground{black}
+    \begin{frame}
+        \frametitle{}
+    \end{frame}
 
-\setstretch{1}
+    \beamersetaveragebackground{white}
+}
+
+\begin{frame}[allowframebreaks]
+    \frametitle{References}
+    \renewcommand{\bibfont}{\normalfont\footnotesize}
     \printbibliography
-\setstretch{1.5}
-
-% \appendix
-
-% The chngctr package is needed for the following lines.
-% \counterwithin{table}{section}
-% \counterwithin{figure}{section}
+\end{frame}
 

### 2024-07-29 docs_template/scripts/task_create_visualization_for_docs.py

@@ -1,151 +1,0 @@
-import shutil
-import subprocess
-import time
-from pathlib import Path
-
-import plotly.graph_objects as go
-import pytask
-
-DOCS = Path(__file__).parent.parent.resolve()  # root/docs
-DOCS_SCRIPTS = DOCS / "scripts"
-FIGURES = DOCS / "source" / "figures" / "generated"
-
-
-for case in ("model_steps_full", "model_steps_select", "steps_only_full"):
-
-    @pytask.task(id=case)
-    def task_visualize_organisational_steps(  # noqa: D103
-        case=case,
-        produces=FIGURES / f"{case}.png",
-    ):
-        fig = _visualize_organisational_steps(case)
-        fig.write_image(produces)
-
-
-for tex_file in ("root_bld_src", "src"):
-
-    @pytask.task(id=tex_file)
-    def task_compile_latex(  # noqa: D103
-        depends_on=DOCS_SCRIPTS / "latex" / f"{tex_file}.tex",
-        produces=DOCS_SCRIPTS / "latex" / f"{tex_file}.png",
-    ):
-        subprocess.run(  # noqa: S603
-            ("pdflatex", "--shell-escape", depends_on.name),
-            cwd=depends_on.parent,
-            check=True,
-        )
-        time.sleep(1)
-        subprocess.run(  # noqa: S603
-            ("pdflatex", "--shell-escape", depends_on.name),
-            cwd=depends_on.parent,
-            check=True,
-        )
-        time.sleep(1)
-
-    @pytask.task(id=tex_file)
-    def task_copy_png_to_figures(  # noqa: D103
-        depends_on=DOCS_SCRIPTS / "latex" / f"{tex_file}.png",
-        produces=FIGURES / f"{tex_file}.png",
-    ):
-        shutil.copyfile(str(depends_on), str(produces))
-
-
-def _visualize_organisational_steps(case):
-    """Create pipeline sketch dependent on {case}.
-
-    Args:
-        case (str): Must be in {'steps_only_full', 'model_steps_full',
-            'model_steps_select'}
-
-    Returns:
-        fig (plotly.graph_objects.Figure): Figure object.
-
-    """
-    fig = go.Figure()
-
-    # define coordinate dimensions
-    fig.add_trace(
-        go.Scatter(
-            x=[0.75, 4.3],
-            y=[0.60, 2.5],
-            mode="text",  # this tells plotly to hide the 'data points'
[…]

### 2024-07-29 docs_template/source/conf.py

@@ -1,164 +1,0 @@
-# documentation build configuration file, created by sphinx-quickstart
-#
-# This file is execfile()d with the current directory set to its
-# containing dir.
-#
-# Note that not all possible configuration values are present in this
-# autogenerated file.
-#
-# All configuration values have a default; values that are commented out
-# serve to show the default.
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-#
-import datetime as dt
-import os
-
-year = dt.datetime.now(tz=None).year  # noqa: DTZ005, ignores missing of timezone info
-
-author = "JANE DOE"
-
-# Set variable so that todos are shown in local build
-on_rtd = os.environ.get("READTHEDOCS") == "True"
-
-
-# -- General configuration ------------------------------------------------
-
-# If your documentation needs a minimal Sphinx version, state it here.
-#
-# Add any Sphinx extension module names here, as strings. They can be
-# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
-# ones.
-extensions = [
-    "sphinx.ext.autodoc",
-    "sphinx.ext.todo",
-    "sphinx.ext.coverage",
-    "sphinx.ext.extlinks",
-    "sphinx.ext.intersphinx",
-    "sphinx.ext.mathjax",
-    "sphinx.ext.viewcode",
-    "sphinx.ext.napoleon",
-    "sphinxcontrib.bibtex",
-    "sphinx_panels",
-    "myst_parser",
-]
-
-# MyST
-myst_enable_extensions = []
-
-
-autoapi_dirs = ["../../hooks"]
-
-bibtex_bibfiles = ["refs.bib"]
-
-autodoc_member_order = "bysource"
-
-autodoc_mock_imports = [
-    "bokeh",
-    "cloudpickle",
-    "cyipopt",
-    "fides",
-    "fuzzywuzzy",
-    "joblib",
-    "nlopt",
-    "pandas",
-    "pytest",
-    "pygmo",
-    "scipy",
-    "sqlalchemy",
-    "tornado",
-    "petsc4py",
-    "statsmodels",
-    "chaospy",
-]
-
-
-intersphinx_mapping = {
-    "numpy": ("https://docs.scipy.org/doc/numpy", None),
-    "pandas": ("https://pandas.pydata.org/pandas-docs/stable", None),
[…]

### 2024-07-29 documents/task_documents.py

@@ -4,12 +4,13 @@ import shutil
 
 import pytask
 from pytask_latex import compilation_steps as cs
-from template_project.config import BLD, DOCUMENTS, ROOT
+from topics_causal_inf.config import BLD, DOCUMENTS, ROOT
 
 documents = ["paper", "presentation"]
 
 for document in documents:
 
+    @pytask.mark.skip()
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
         document=BLD / "documents" / f"{document}.pdf",
@@ -26,6 +27,7 @@ for document in documents:
         "produces": ROOT / f"{document}.pdf",
     }
 
+    @pytask.mark.skip()
     @pytask.task(id=document, kwargs=kwargs)
     def task_copy_to_root(depends_on, produces):
         """Copy a document to the root directory for easier retrieval."""

### 2024-07-29 tests/analysis/test_model.py

@@ -1,27 +1,0 @@
-"""Tests for the regression model."""
-
-import numpy as np
-import pandas as pd
-import pytest
-from template_project.analysis.model_template import fit_logit_model
-
-DESIRED_PRECISION = 10e-2
-
-
-@pytest.fixture()
-def data():
-    rng = np.random.default_rng(seed=0)
-    x = rng.normal(size=100_000)
-    coef = 2.0
-    prob = 1 / (1 + np.exp(-coef * x))
-    return pd.DataFrame(
-        {"outcome_numerical": rng.binomial(1, prob), "covariate": x},
-    )
-
-
-def test_fit_logit_model_recover_coefficients(data):
-    formula = "outcome_numerical ~ covariate"
-    model = fit_logit_model(data, formula=formula)
-    params = model.params
-    assert np.abs(params["Intercept"]) < DESIRED_PRECISION
-    assert np.abs(params["covariate"] - 2.0) < DESIRED_PRECISION

### 2024-07-29 tests/analysis/test_predict.py

@@ -1,46 +1,0 @@
-"""Tests for the prediction model."""
-
-import pandas as pd
-import pytest
-from pandas.testing import assert_frame_equal
-from template_project.analysis.predict_template import predict_prob_by_age
-
-
-@pytest.fixture()
-def data():
-    out = pd.DataFrame([1, 2, 3], columns=["age"])
-    out["education"] = ["high-school", "high-school", "university"]
-    out["income"] = ["high", "low", "low"]
-    return out
-
-
-@pytest.fixture()
-def model():
-    class ModelClass:
-        @staticmethod
-        def predict(data):
-            if "high-school" in data["education"].to_numpy():
-                prob = 0.1 if "low" in data["income"].to_numpy() else 0.2
-            else:
-                prob = 0.3 if "low" in data["income"].to_numpy() else 0.4
-            return prob * data["age"]
-
-    return ModelClass
-
-
-@pytest.mark.parametrize("group", ["education", "income"])
-def test_predict_prob_over_age(data, model, group):
-    got = predict_prob_by_age(data, model, group)
-
-    if group == "education":
-        expected = pd.DataFrame(
-            [[1, 0.1, 0.3], [2, 0.2, 0.6], [3, 0.3, 0.9]],
-            columns=["age", "high-school", "university"],
-        )
-    else:
-        expected = pd.DataFrame(
-            [[1, 0.2, 0.1], [2, 0.4, 0.2], [3, 0.6, 0.3]],
-            columns=["age", "high", "low"],
-        )
-
-    assert_frame_equal(got, expected)

### 2024-07-29 tests/data_management/test_stats4schools_smoking.py

@@ -1,142 +1,0 @@
-import numpy as np
-import pandas as pd
-import pytest
-from pandas.testing import assert_series_equal
-from template_project.data_management.stats4schools_smoking_template import (
-    _clean_current_smoker,
-    _clean_gender,
-    _clean_highest_qualification,
-    _clean_marital_status,
-    clean_stats4schools_smoking,
-)
-
-
-def assert_categorical_equal(left, right):
-    assert_series_equal(pd.Series(left), pd.Series(right))
-
-
-@pytest.fixture()
-def data():
-    data = {
-        "highest_qualification": [
-            "No Qualification",
-            "GCSE/CSE",
-            "GCSE/O Level",
-            "ONC/BTEC",
-            "Other/Sub Degree",
-            "Higher/Sub Degree",
-            "A Levels",
-            "Degree",
-        ],
-        "gender": 4 * ["Male"] + 4 * ["Female"],
-        "age": 4 * [20] + 4 * [20.5],
-        "marital_status": [
-            "Divorced",
-            "Widowed",
-            "Separated",
-            "Single",
-            "Married",
-            "Widowed",
-            "Separated",
-            "Married",
-        ],
-        "smoke": 4 * ["Yes"] + 4 * ["No"],
-    }
-    return pd.DataFrame(data)
-
-
-def test_age_is_integer(data):
-    got = clean_stats4schools_smoking(data)["age"]
-    exp = pd.Series(8 * [20], name="age")
-    assert_series_equal(got, exp)
-
-
-def test_current_smoker_numerical_is_numerical(data):
-    got = clean_stats4schools_smoking(data)["current_smoker_numerical"]
-    exp = pd.Series(4 * [1] + 4 * [0], name="current_smoker_numerical", dtype="int8")
-    assert_series_equal(got, exp)
-
-
-def test_clean_gender(data):
-    got = _clean_gender(data["gender"])
-    exp = pd.Categorical(
-        4 * ["Male"] + 4 * ["Female"],
-        categories=["Female", "Male"],
-        ordered=False,
-    )
-    assert_categorical_equal(got, exp)
-
-
-def test_clean_marital_status(data):
-    got = _clean_marital_status(data["marital_status"])
-    exp = pd.Categorical(
-        [
-            "Divorced",
-            "Widowed",
-            "Separated",
-            "Single",
-            "Married",
-            "Widowed",
-            "Separated",
-            "Married",
-        ],
[…]

### 2024-07-29 tests/test_template.py

@@ -1,15 +1,0 @@
-import pytask
-from _pytask.outcomes import ExitCode
-from template_project import config
-from template_project.config import ROOT
-
-
-def test_pytask_build(monkeypatch, tmp_path):
-    monkeypatch.setattr(config, "ROOT", tmp_path)
-    monkeypatch.setattr(config, "BLD", tmp_path / "bld")
-
-    session = pytask.build(
-        config=ROOT / "pyproject.toml",
-        force=True,
-    )
-    assert session.exit_code == ExitCode.OK

### 2024-07-29 src/topics_causal_inf/grf/__init__.py
Add first simulation functions and tasks.

@@ -1,0 +1,1 @@
+"""Test some grf features and run simple examples."""

### 2024-07-29 src/topics_causal_inf/grf/grf_example.R

@@ -1,0 +1,60 @@
+# ======================================================================================
+# Get config from pytask
+# ======================================================================================
+
+args <- commandArgs(trailingOnly = TRUE)
+path_to_yaml <- args[length(args)]
+config <- yaml::yaml.load_file(path_to_yaml)
+
+# ======================================================================================
+# Main
+# ======================================================================================
+library(grf)
+
+# Generate data.
+n <- 2000
+p <- 10
+X <- matrix(rnorm(n * p), n, p)
+X.test <- matrix(0, 101, p)
+X.test[, 1] <- seq(-2, 2, length.out = 101)
+
+# Train a causal forest.
+W <- rbinom(n, 1, 0.4 + 0.2 * (X[, 1] > 0))
+Y <- pmax(X[, 1], 0) * W + X[, 2] + pmin(X[, 3], 0) + rnorm(n)
+tau.forest <- causal_forest(X, Y, W)
+
+# Estimate treatment effects for the training data using out-of-bad prediction.
+tau.hat.oob <- predict(tau.forest)
+hist(tau.hat.oob$predictions)
+
+# Estimate treatment effects for the test sample.
+tau.hat <- predict(tau.forest, X.test)
+plot(X.test[, 1], tau.hat$predictions, ylim = range(tau.hat$predictions, 0, 2), xlab = "x", ylab = "tau", type = "l")
+lines(X.test[, 1], pmax(0, X.test[, 1]), col = 2, lty = 2)
+
+# Estimate the conditional (sic?) average treatment effect on the full sample (CATE).
+average_treatment_effect(tau.forest, target.sample = "all")
+
+# Estimate the conditional average treatment effect on the treated sample (CATT).
+average_treatment_effect(tau.forest, target.sample = "treated")
+
+# Add confidence intervals for heterogeneous treatment effects;
+# growing more trees is now recommended.
+tau.forest <- causal_forest(X, Y, W, num.trees = 4000)
+tau.hat <- predict(tau.forest, X.test, estimate.variance = TRUE)
+sigma.hat <- sqrt(tau.hat$variance.estimates)
+
+png(paste(config[["produces"]]))
+par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs = "i", yaxs = "i")
+
+plot(X.test[, 1], tau.hat$predictions,
[…]

### 2024-07-29 src/topics_causal_inf/grf/task_grf_example.py

@@ -1,0 +1,26 @@
+"""Example task for grf example in R."""
+
+import pytask
+
+from topics_causal_inf.config import BLD, SRC
+
+
+@pytask.mark.r(
+    script=SRC / "grf" / "grf_example.R",
+    serializer="yaml",
+)
+def task_grf_example_in_r(
+    produces=BLD / "grf" / "grf_example_in_r.png",
+):
+    """Produce example figure using grf in R."""
+
+
+@pytask.mark.r(
+    script=SRC / "grf" / "bruhn_2016.R",
+    serializer="yaml",
+)
+def task_grf_example_bruhn_2016(
+    produces=BLD / "grf" / "bruhn_2016.rds",
+    bruhn_2016_data=SRC / "grf" / "bruhn2016.csv",
+):
+    """Produce example figure using grf in R."""

### 2024-07-29 src/topics_causal_inf/wa_replication/__init__.py

@@ -1,0 +1,1 @@
+"""Replication of Wager and Athey (2018) results."""

### 2024-07-29 src/topics_causal_inf/wa_replication/task_wa_replication.py

@@ -1,0 +1,25 @@
+"""Tasks for replication of Wager and Athey (2018) results."""
+
+from pathlib import Path
+
+import pytask
+
+from topics_causal_inf.config import BLD, SRC
+
+SIM_KWARGS = {
+    "n_obs": 500,
+    "n_sim": 100,
+    "dim": 2,
+}
+
+for dgp in ["dgp1", "dgp2", "dgp3"]:
+
+    @pytask.task(kwargs={"dgp": dgp, "return_grid": False, **SIM_KWARGS}, id=dgp)
+    @pytask.mark.r(
+        script=SRC / "wa_replication" / "wa_replication.R",
+        serializer="yaml",
+    )
+    def task_wa_replication(
+        produces: Path = BLD / "wa_replication" / f"wa_replication_{dgp}.rds",
+    ):
+        """Produce example figure using grf in R."""

### 2024-07-29 src/topics_causal_inf/wa_replication/task_wa_replication_grid_only.py

@@ -1,0 +1,20 @@
+"""Tasks for replication of Wager and Athey (2018) results."""
+
+from pathlib import Path
+
+import pytask
+
+from topics_causal_inf.config import BLD, SRC
+from topics_causal_inf.wa_replication.task_wa_replication import SIM_KWARGS
+
+for dgp in ["dgp1", "dgp2", "dgp3"]:
+
+    @pytask.task(kwargs={"dgp": dgp, "return_grid": True, **SIM_KWARGS}, id=dgp)
+    @pytask.mark.r(
+        script=SRC / "wa_replication" / "wa_replication.R",
+        serializer="yaml",
+    )
+    def task_wa_replication_grid(
+        produces: Path = BLD / "wa_replication" / f"wa_replication_{dgp}_grid_only.rds",
+    ):
+        """Produce example figure using grf in R."""

### 2024-07-29 src/topics_causal_inf/wa_replication/wa_replication.R

@@ -1,0 +1,148 @@
+library(grf)
+
+# ======================================================================================
+# Functions
+# ======================================================================================
+
+# Simulation
+simulation <- function(
+    n_sim, n_obs, dim, main_effect, treatment_effect, propensity,
+    return_grid = FALSE) {
+  # Run experiments
+  if (return_grid == TRUE) {
+    return(experiment(n_obs, dim, main_effect, treatment_effect, propensity,
+      return_grid = return_grid
+    ))
+  }
+
+  mse <- rep(0, n_sim)
+  for (i in 1:n_sim) {
+    mse[i] <- experiment(n_obs, dim, main_effect, treatment_effect, propensity,
+      return_grid = return_grid
+    )
+  }
+
+  return(mse)
+}
+
+# Single experiment
+experiment <- function(
+    n_obs, dim, main_effect, treatment_effect, propensity,
+    return_grid = FALSE) {
+  # Draw data
+  data <- data(n_obs, dim, main_effect, treatment_effect, propensity)
+  data_test <- data(n_obs, dim, main_effect, treatment_effect, propensity)
+
+  # Fit causal forest
+  cf <- causal_forest(data$x, data$y, data$w, num.trees = 2000, sample.fraction = 0.5)
+
+  if (!return_grid) {
+    # Predict on test-data
+    pred <- predict(cf, data_test$x)
+    true <- treatment_effect(data_test$x)
+
+    # Evaluate fit by mse
+    mse <- mean((pred$predictions - true)^2)
+
+    # Make prediction on x1, x2 from 0 to 1
+    return(mse)
+  }
+
+  data_eval <- data(10000, dim, main_effect, treatment_effect, propensity)
+
+  pred <- predict(cf, data_eval$x)
+
+  pred_eval <- pred$predictions
+  true_eval <- treatment_effect(data_eval$x)
+
+  # Put into one dataframe
+  mse <- mean((pred_eval - true_eval)^2)
+
+  out <- cbind(data_eval$x, pred_eval, true_eval, mse)
+
+  colnames(out) <- c(paste0("x", 1:dim), "pred", "true", "mse")
+
+  return(out)
+}
+
+# Generate data
+data <- function(n_obs, dim, main_effect, treatment_effect, propensity) {
+  x <- matrix(0, n_obs, dim)
[…]

### 2024-07-29 src/topics_causal_inf/wa_replication/wa_replication_debug.R

@@ -1,0 +1,148 @@
+library(grf)
+
+# ======================================================================================
+# Functions
+# ======================================================================================
+
+# Simulation
+simulation <- function(
+    n_sim, n_obs, dim, main_effect, treatment_effect, propensity,
+    return_grid = FALSE) {
+  # Run experiments
+  if (return_grid == TRUE) {
+    return(experiment(n_obs, dim, main_effect, treatment_effect, propensity,
+      return_grid = return_grid
+    ))
+  }
+
+  mse <- rep(0, n_sim)
+  for (i in 1:n_sim) {
+    mse[i] <- experiment(n_obs, dim, main_effect, treatment_effect, propensity,
+      return_grid = return_grid
+    )
+  }
+
+  return(mse)
+}
+
+# Single experiment
+experiment <- function(
+    n_obs, dim, main_effect, treatment_effect, propensity,
+    return_grid = FALSE) {
+  # Draw data
+  data <- data(n_obs, dim, main_effect, treatment_effect, propensity)
+  data_test <- data(n_obs, dim, main_effect, treatment_effect, propensity)
+
+  # Fit causal forest
+  cf <- causal_forest(data$x, data$y, data$w, num.trees = 2000, sample.fraction = 0.5)
+
+  if (!return_grid) {
+    # Predict on test-data
+    pred <- predict(cf, data_test$x)
+    true <- treatment_effect(data_test$x)
+
+    # Evaluate fit by mse
+    mse <- mean((pred$predictions - true)^2)
+
+    # Make prediction on x1, x2 from 0 to 1
+    return(mse)
+  }
+
+  data_eval <- data(10000, dim, main_effect, treatment_effect, propensity)
+
+  pred <- predict(cf, data_eval$x)
+
+  pred_grid <- pred$predictions
+  true_grid <- treatment_effect(data_eval$x)
+
+  # Put into one dataframe
+  mse <- mean((pred_grid - true_grid)^2)
+
+  out <- cbind(data_eval$x, pred_grid, true_grid, mse)
+
+  colnames(out) <- c(paste0("x", 1:dim), "pred", "true", "mse")
+
+  return(out)
+}
+
+# Generate data
+data <- function(n_obs, dim, main_effect, treatment_effect, propensity) {
+  x <- matrix(0, n_obs, dim)
[…]

### 2024-07-29 src/topics_causal_inf/grf/task_grf_example.py
Repeat simulations for different ambient dimensions.

@@ -15,6 +15,7 @@ def task_grf_example_in_r(
     """Produce example figure using grf in R."""
 
 
+@pytask.mark.skip()
 @pytask.mark.r(
     script=SRC / "grf" / "bruhn_2016.R",
     serializer="yaml",

### 2024-07-29 src/topics_causal_inf/wa_replication/task_tables.py

@@ -1,0 +1,46 @@
+"""Generate tables for the WA replication."""
+
+from pathlib import Path
+
+import pandas as pd  # type: ignore[import-untyped]
+import pyreadr
+import pytask
+
+from topics_causal_inf.config import BLD
+from topics_causal_inf.wa_replication.sim_config import DGP_VALS, DIM_VALS
+
+RESULTS_PATHS = {
+    dgp: [
+        BLD / "wa_replication" / f"wa_replication_{dgp}_dim{dim}.rds"
+        for dim in DIM_VALS
+    ]
+    for dgp in DGP_VALS
+}
+
+for dgp in DGP_VALS:
+    # Generate list of all paths corresponding to a dgp and each value of dim
+
+    @pytask.task(id=f"{dgp}_table")
+    def task_wa_replication_table(
+        depends_on=RESULTS_PATHS[dgp],
+        produces: Path = BLD / "wa_replication" / f"wa_replication_{dgp}.tex",
+    ):
+        """Generate tables for the WA replication."""
+        # Combine all results into a single dataframe
+        dfs = [pyreadr.read_r(result)[None] for result in depends_on]
+
+        for df in dfs:
+            df.columns = ["mse"]
+
+        # To each add column with dim
+        for i, df in enumerate(dfs):
+            df["dim"] = DIM_VALS[i]
+
+        # Concatenate all dataframes
+        res = pd.concat(dfs)
+
+        # Group by dim and calculate mean and std of the mse
+        res_grouped = res.groupby("dim")["mse"].agg(["mean", "std"])
+
+        # Save table
+        res_grouped.to_latex(produces)

### 2024-08-14 documents/paper.tex
Start writing paper.

@@
 % submissions.
+\DeclareMathOperator{\E}{E}
 
@@
 \title{Topics Causal Inference}
-\thanks{EXAMPLE UNIVERSITY. Email: janedoe[at]example-university.de}
+\thanks{Bonn University. Email: s93jbudd[at]uni-bonn.de}
 
@@
 \date{
-    {\bf Preliminary -- please do not quote}
-    \\[1ex]
     \today
@@
 \begin{abstract}
-    Some abstract here.
+    This term paper presents simulation results for two connected approaches to estimating conditional average treatment effects (CATE):
+    The \textit{causal forest} approach originally proposed in \cite{wager2018estimation} and the \textit{generic machine learning} approach discussed in \cite{chernozhukov2023genml}.
+    In a first part, I replicate simulation results for causal forest reported in the original paper that demonstrate the ability to estimate a sparse CATE function and perform valid inference.
+    I then continue to extend their analysis by lowering the degree of sparseness of the DGP and show that their method breaks down: RMSE is significantly increased and coverage is far below the nominal level.
+    I show that inference on the BLP is still possible using the generic machine learning approach.
+However, for the DGP I consider the RMSE is still considerably worse, illustrating a trade-off between estimation properties and informativeness of the inferential target.
+    As an additional point of interest I consider the design of high-dimensional monte carlo studies using the approach suggested by \cite{athey2024wgan}.
 \end{abstract}
@@
 
-\section{Introduction} % (fold)
-\label{sec:introduction}
+\section{Introduction}
 
-If you are using this template, please cite this item from the references:
-\citet{GaudeckerEconProjectTemplates}.
+\section{Methods}
 
-The data set for the template project is taken from
-\url{https://www.stem.org.uk/resources/elibrary/resource/28452/large-datasets-stats4schools}.
-It contains data on smoking habits in the UK, with 1691 observations and 12 variables.
-We consider only 4 of the 12 features for the prediction of the variable
-\texttt{smoking}: \texttt{marital\_status}, \texttt{highest\_qualification},
-\texttt{gender} and \texttt{age}. We model the dependence using a Logistic model. All
-numerical features are included linearly, while categorical features are expanded into
-dummy variables. Figures below illustrate the model predictions over the lifetime. You
-will find one figure and one estimation summary table for each installed programming
-language.
+\subsection{Problem Setup}
+Throughout we consider a setting with a binary treatment denoted by $D$ taking on values in $\{0, 1\}$.
+$Z$ will denote a vector of covariates that has dimension $p$. The sample size will be noted by $n$.
+We will consider generally consider a setting of \textit{unconfoundedness}, meaning that conditional on covariates the treatment is as-good-as randomly assigned.
+Denoting the outcomes obtained in states of the world with treatment realization $D=d$ by $Y(d)$, this means
+\begin{equation*}
+    Y(0), Y(1) \perp D | Z.
+\end{equation*}
+In the context of \cite{chernozhukov2023genml} we will additionally make the assumption that the \textit{propensity score} $p(z) \equiv \Pr(D=1|Z=z)$ is a known function.
+The leading case of this is a stratified experiment, with $p(z) = 0.5$ being the special case of a completely randomized experiment.
+Generally, the literature also often allows for an unknown propensity score which poses the problem of --- high-dimensional --- first-step estimation.
+Note this is not the ``dimensionality'' issue considered here, which is about the complexity of conditional average treatment effects.
+The target of this literature is the \textit{conditional average treatment effect} (CATE), which is the average treatment effect for individuals with covariates $Z=z$:
+\begin{equation*}
+    CATE(Z) = s_0(Z) = \E[Y(1) - Y(0)|Z].
+\end{equation*}
+Our goal is to estimate and perform inference on this quantity.
 
-\begin{figure}[H]
+Ideally, we would like to have a procedure with the following \textit{properties}:
+\begin{itemize}
+    \item[(1)] Consistent estimation in the sense that the mean-squared error (MSE) converges to 0;
+    \item[(2)] valid inference, i.e. confidence intervals, for $CATE(z)$;
+    \item[(3)] ability to adapt to high-dimensional covariate spaces (regimes where $p$ grows with $n$) without making sparsity assumptions;
+    \item[(4)] no functional form restrictions on $s_0(z)$.
+\end{itemize}
+However, we know from general results on optimal achievable non-parametric rates that such a procedure cannot exist.
+All proposed approaches then proceed by relaxing one or more of the requirements stated above.
+The following briefly summarizes some of the ideas, a more comprehensive list of reference can be found in \cite{causalml2024}.
+I focus in particular on sparsity assumptions and approximations to the CATE function, because these are two ideas that are taken up in the generic machine learning approach discussed later.
+\paragraph*{Sparsity}
+A first strategy is to place sparsity assumptions on the CATE function.
+For example, a common set of approaches postulates a linear conditional mean function:
+\begin{equation*}
+    \E[Y|Z] = \sum_{l=1}^{p_1}\alpha_l DX_l + \sum_{j=1}^{p_2}\beta_j W_j,
+\end{equation*}
+where the covariates $Z$ are split into those determining systematic heterogeneity of treatment effects, $X$, and controls, $W$.
+In this model we have hence have $s_0(Z) = S_0(X) = \alpha'X$ under randomization. In these setups we can allow for both $p_1$ and $p_2$ to grow with the sample size.
+Hence, the CATE function can be high-dimensional at least for a subset of the covariates.
+For these linear models Lasso-type approaches adapted to causal inference (so called "double" lasso) will perform well, under a condition called \textit{approximate sparsity}.
+Intuitively, this condition is fulfilled by DGPs where the magnitude of the (ordered) coefficients $\alpha_l, \beta_l$, decays sufficiently fast.
+One such DGP is for example given by
+\begin{equation*}
+    \alpha_l = \frac{1}{l^2}, \qquad \beta_j = \frac{1}{j^2}.
+\end{equation*}
+Assuming $p_1$, the dimension of heterogeneity, does not grow too fast and we have approximate sparsity, it is possible to show an asymptotic normal approximation, namely
+\begin{equation*}
+    \sqrt{n}(\hat{\alpha}-\alpha) \sim_a N(0,V)
+\end{equation*}
+where $V$ can be consistently estimated. These results are shown in \cite{belloni2015uniform} and \cite{belloni2018uniformly}.
 
-    \centering
-    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_python}
+\paragraph*{Approximations to the CATE}
+Another approach is taken in \cite{semenova2021debiased}:
+They propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
+Then, the new target for inference becomes the \textit{partial} CATE $s_0(X)$.
+Their paper then simultaneously allows for \textit{unconfoundedness} which poses the challenge of estimating $p(Z)$ --- potentially using machine learning methods --- in a first step.
+In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE.
+Hence, they essentially use the following approximations to make the problem solvable:
+\begin{equation*}
+    s_0(Z) \approx s_0(X) \approx \beta'X.
+\end{equation*}
+The best linear predictor (BLP) problem in this case is given by
+\begin{equation*}
+    \min_{b\in\mathbb{R}^p} \E [(s_o(X) - X'b)^2].
+\end{equation*}
+We can then make use of the following identification result (which is sometimes referred to as the Horvitz-Thomson transform):
+\begin{equation*}
+    s_0(Z) = \E[YH|Z]
+\end{equation*}
+where $H = \frac{D-p(Z)}{p(Z)(1-p(Z))}$.
+It is then easy to derive the solution of the BLP problem above as
+\begin{equation*}
+    \beta = E[XX']^{-1}E[XHY]
+\end{equation*}
+which instead of the BLP of $Y$ uses the transformed outcome $HY$.
+For this we can then make use of standard least squares theory --- given that the nuisance parameter estimation in the first step satisfies certain regularity conditions --- and show
+\begin{equation*}
+    \hat{\beta} - \beta_0 \sim_a N(0, \Omega/N).
+\end{equation*}
+\subsection{Generic Machine Learning}
 
-    \caption{Model predictions of the smoking probability over the lifetime. Each
-        colored line represents a case where marital status is fixed to one of the
-        values present in the data set.}
-    \label{fig:predictions}
+\section{Simulation and Results}
 
-\end{figure}
+\clearpage
+\newpage
 
-\begin{table}[!h]
-    \input{../tables/estimation_results.tex}
-    \label{tab:summary}
-    \caption{Estimation results of the linear Logistic regression.}
-\end{table}
-
-\begin{figure}[H]
-
-    \centering
-    \includegraphics[width=0.85\textwidth]{../figures/smoking_by_marital_status_using_r}
-
-    \caption{\emph{R:} Model predictions of the smoking probability over
-        the lifetime. Each colored line represents a case where marital status
-        is fixed to one of the values present in the data set.}
-    \label{fig:r-predictions}
-
-\end{figure}
-
-section introduction (end)
-
-\setstretch{1}
+% \setstretch{1}
 \printbibliography

### 2024-08-14 documents/task_documents.py

@@ -6,11 +6,10 @@ import pytask
 from pytask_latex import compilation_steps as cs
 from topics_causal_inf.config import BLD, DOCUMENTS, ROOT
 
-documents = ["paper", "presentation"]
+documents = ["paper"]
 
 for document in documents:
 
-    @pytask.mark.skip()
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
         document=BLD / "documents" / f"{document}.pdf",
@@ -27,7 +26,6 @@ for document in documents:
         "produces": ROOT / f"{document}.pdf",
     }
 
-    @pytask.mark.skip()
     @pytask.task(id=document, kwargs=kwargs)
     def task_copy_to_root(depends_on, produces):
         """Copy a document to the root directory for easier retrieval."""

### 2024-08-14 documents/paper.tex
Add first sections in generic_ml to paper.

@@
     This term paper presents simulation results for two connected approaches to estimating conditional average treatment effects (CATE):
-    The \textit{causal forest} approach originally proposed in \cite{wager2018estimation} and the \textit{generic machine learning} approach discussed in \cite{chernozhukov2023genml}.
+    The \textit{causal forest} approach originally proposed in~\cite{wager2018estimation} and the \textit{generic machine learning} approach discussed in~\cite{chernozhukov2023genml}.
     In a first part, I replicate simulation results for causal forest reported in the original paper that demonstrate the ability to estimate a sparse CATE function and perform valid inference.
@@
 However, for the DGP I consider the RMSE is still considerably worse, illustrating a trade-off between estimation properties and informativeness of the inferential target.
-    As an additional point of interest I consider the design of high-dimensional monte carlo studies using the approach suggested by \cite{athey2024wgan}.
+    As an additional point of interest I consider the design of high-dimensional monte carlo studies using the approach suggested by~\cite{athey2024wgan}.
 \end{abstract}
@@
 \end{equation*}
-In the context of \cite{chernozhukov2023genml} we will additionally make the assumption that the \textit{propensity score} $p(z) \equiv \Pr(D=1|Z=z)$ is a known function.
+In the context of~\cite{chernozhukov2023genml} we will additionally make the assumption that the \textit{propensity score} $p(z) \equiv \Pr(D=1|Z=z)$ is a known function.
 The leading case of this is a stratified experiment, with $p(z) = 0.5$ being the special case of a completely randomized experiment.
@@
     \item[(1)] Consistent estimation in the sense that the mean-squared error (MSE) converges to 0;
-    \item[(2)] valid inference, i.e. confidence intervals, for $CATE(z)$;
+    \item[(2)] valid inference, i.e.\ confidence intervals, for $CATE(z)$;
     \item[(3)] ability to adapt to high-dimensional covariate spaces (regimes where $p$ grows with $n$) without making sparsity assumptions;
@@
 All proposed approaches then proceed by relaxing one or more of the requirements stated above.
-The following briefly summarizes some of the ideas, a more comprehensive list of reference can be found in \cite{causalml2024}.
+The following briefly summarizes some of the ideas, a more comprehensive list of reference can be found in~\cite{causalml2024}.
 I focus in particular on sparsity assumptions and approximations to the CATE function, because these are two ideas that are taken up in the generic machine learning approach discussed later.
@@
 Hence, the CATE function can be high-dimensional at least for a subset of the covariates.
-For these linear models Lasso-type approaches adapted to causal inference (so called "double" lasso) will perform well, under a condition called \textit{approximate sparsity}.
+For these linear models Lasso-type approaches adapted to causal inference (so called ``double'' lasso) will perform well, under a condition called \textit{approximate sparsity}.
 Intuitively, this condition is fulfilled by DGPs where the magnitude of the (ordered) coefficients $\alpha_l, \beta_l$, decays sufficiently fast.
@@
 \end{equation*}
-where $V$ can be consistently estimated. These results are shown in \cite{belloni2015uniform} and \cite{belloni2018uniformly}.
+where $V$ can be consistently estimated. These results are shown in~\cite{belloni2015uniform} and~\cite{belloni2018uniformly}.
 
 \paragraph*{Approximations to the CATE}
-Another approach is taken in \cite{semenova2021debiased}:
+Another approach is taken in~\cite{semenova2021debiased}:
 They propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
@@
 Their paper then simultaneously allows for \textit{unconfoundedness} which poses the challenge of estimating $p(Z)$ --- potentially using machine learning methods --- in a first step.
-In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE.
+In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE\@.
 Hence, they essentially use the following approximations to make the problem solvable:
@@
 \end{equation*}
+
 \subsection{Generic Machine Learning}
+The generic ML approach allows want to place as little assumptions as possible on the data generating process, while maintaining valid inference.
+This is achieved by (a) moving the inferential target to \textit{features} of the CATE (like the BLP) and (b) using sample splitting to derive inferential guarantees.
+Hence, this comes at the cost of potentially being able to only make statements about a (bad) approximation to the original inferential target as well as a loss in precision due to sample splitting.
+However, there are a number of clear advantages: The researcher does not have to commit to a specific --- small --- set of covariates ex-ante or make sparsity assumptions.
+At the same time, the approach as discussed in \cite{chernozhukov2023genml} does not allow for nuisance parameter estimation, i.e.\ requires a known propensity score.
+
+\paragraph*{Two Stages}
+A key feature of the approach is \textit{sample splitting}. Let $(M,A)$ denote two partitions of the sample into an auxiliary and a main dataset.
+In Stage 1, we then use the auxiliary dataset $A$ to generate machine predictors for the maps $z\to B(z)$ (the BCA) and $z\to S(z)$ (the CATE).
+In Stage 2, we then proceed to use $S(z), B(z)$ applied to the main dataset $M$ to make inference on \textit{features} of $z\to s_0(z)$, for example a best linear predictor.
+
+\paragraph*{BLP of CATE}
+The first feature is the best linear predictor of the CATE function \textit{by the machine learning proxy} $S(Z)$.
+In particular, we want to solve
+\begin{equation*}
+\min_{(b1,b2)\in \mathbb{R}^2} \E [(s_0(Z) - b_1 - b_2 S(Z))^2]
+\end{equation*}
+which, if it exists, is given by
+\begin{equation*}
+    BLP[s_0(Z)|S(Z)] := \beta_1 + \beta_2(S(Z) - \E S(Z)).
+\end{equation*}
+Here, the solutions are given by $\beta_1 = \E s_0(Z)$ and $\beta_2 = Cov[s_0(Z), S(Z)]/Var[S(Z)]$.
+
+A simple approach to estimating $\beta_1, \beta_2$ is again the Horvitz-Thompson transform discussed above.
+In particular, it is easy to show that the projection
+\begin{equation*}
+    HY = \mu_1 + \mu_2(S(Z) - \E S(Z)) + \varepsilon, \qquad \E[(1, S(Z) - \E S(Z))'\varphi] = 0
+\end{equation*}
+identifies the BLP parameters, i.e.\ $\mu_1=\beta_1$ and $\mu_2 = \beta_2$.
+Estmation can then proceed based on the empirical analogue.
+While inference \textit{conditional} on the split is standard, unconditional inference has to rely on median aggregation of many splits (discussed below).
+
+\paragraph*{Other Targets}
+Other potential targets beside the BLP of the CATE are sorted group average treatment effects (GATES) and classification analysis (CLAN).
+GATES are average effects for a finite number of groups that are induced by the first stage machine learning proxy.
+For example, we define five groups by the size of their predicted treatment effect $S(z)$.
+More generally, we can consider any finite partition of the real line $-\infty = I_0 < I_1 < \cdots < I_K = \infty$ and the define GATES by
+\begin{equation*}
+    \gamma_k := \E[s_0(Z)|G_k], \qquad k=1,\ldots,K,
+\end{equation*}
+where group membership is determined by $G_k := I\{S(Z) \in I_k\}$.
+In CLAN, we are interested in characteristics $g(Y, D, Z)$ and we want to compare those for the least ($G_1$) and most ($G_K$) affected groups:
+\begin{equation*}
+    \delta_g := \E[g(Y, D, Z)|G_g], \qquad g \in \{1, K\}.
+\end{equation*}
+
+\paragraph*{ML Method Selection}
+[Tbd.]
+
+\paragraph*{Split-based Inference}
 
@@
 % \setstretch{1}
-\printbibliography
+\printbibliography{}
 \setstretch{1.5}

### 2024-08-14 documents/paper.tex
Add section on software; add WGAN results.

@@
 However, there are a number of clear advantages: The researcher does not have to commit to a specific --- small --- set of covariates ex-ante or make sparsity assumptions.
-At the same time, the approach as discussed in \cite{chernozhukov2023genml} does not allow for nuisance parameter estimation, i.e.\ requires a known propensity score.
+At the same time, the approach as discussed in~\cite{chernozhukov2023genml} does not allow for nuisance parameter estimation, i.e.\ requires a known propensity score.
 
@@
 
+\subsection{Software Packages}
+For the baseline causal forest estimator I rely on the `CausalForest' implementation provided by the \href{https://econml.azurewebsites.net/}{EconML} Python package.
+This package implements a range of machine learning techniques for causal based on the statistical literature and making use of existing machine learning implementations in Python.
+For the generic ML part, I implemented most algorithms by hand since these only involve specifying linear regressions and sample splitting.
+Again, it is possible to flexibly use any of the machine learning estimators in this case, for example implemented in EconML or sklearn directly.
+For estimating DGPs using Wasserstein Generative Adversarial Networks (WGAN) I use the \href{https://github.com/gsbDBI/ds-wgan}{ds-wgan} implementation provided by the authors.
+Unfortunately, the code is not maintained and not compatible with recent versions of most basic Python libraries (e.g. NumPy 2.0).
+To avoid dependency conflicts with the other packages or manually debugging their code, I instead trained the generators in a Google Colab environment, which also comes at a significant speedup due to GPU availability.
+
 \clearpage

### 2024-08-14 src/topics_causal_inf/grf/__init__.py
Delete old code.

@@ -1,1 +1,0 @@
-"""Test some grf features and run simple examples."""

### 2024-08-14 src/topics_causal_inf/grf/bruhn_2016.R

@@ -1,48 +1,0 @@
-library(grf)
-
-# ======================================================================================
-# Get config from pytask
-# ======================================================================================
-
-args <- commandArgs(trailingOnly = TRUE)
-path_to_yaml <- args[length(args)]
-config <- yaml::yaml.load_file(path_to_yaml)
-
-# ======================================================================================
-# Main
-# ======================================================================================
-
-data <- read.csv(config[["bruhn_2016_data"]])
-
-y <- data$outcome
-
-w <- data$treatment
-
-school <- data$school
-
-x <- data[-(1:3)]
-
-sum(!complete.cases(x)) / nrow(x)
-
-t.test(w ~ !complete.cases(x))
-
-cf <- causal_forest(x, y, w, W.hat = 0.5, clusters = school)
-
-ate <- average_treatment_effect(cf)
-ate
-
-varimp <- variable_importance(cf)
-ranked.vars <- order(varimp, decreasing = TRUE)
-
-# Top 5 variables according to this measure
-colnames(x)[ranked.vars[1:5]]
-
-# Q: How does this know what to cluster by?
-# --> specified clusters in the causal_forest function
-best_linear_projection(cf, x[ranked.vars[1:5]])
-
-pred <- predict(cf, x)
-
-hist(pred$predictions)
-
-saveRDS(pred, file = config[["produces"]])

### 2024-08-14 src/topics_causal_inf/grf/grf_example.R

@@ -1,60 +1,0 @@
-# ======================================================================================
-# Get config from pytask
-# ======================================================================================
-
-args <- commandArgs(trailingOnly = TRUE)
-path_to_yaml <- args[length(args)]
-config <- yaml::yaml.load_file(path_to_yaml)
-
-# ======================================================================================
-# Main
-# ======================================================================================
-library(grf)
-
-# Generate data.
-n <- 2000
-p <- 10
-X <- matrix(rnorm(n * p), n, p)
-X.test <- matrix(0, 101, p)
-X.test[, 1] <- seq(-2, 2, length.out = 101)
-
-# Train a causal forest.
-W <- rbinom(n, 1, 0.4 + 0.2 * (X[, 1] > 0))
-Y <- pmax(X[, 1], 0) * W + X[, 2] + pmin(X[, 3], 0) + rnorm(n)
-tau.forest <- causal_forest(X, Y, W)
-
-# Estimate treatment effects for the training data using out-of-bad prediction.
-tau.hat.oob <- predict(tau.forest)
-hist(tau.hat.oob$predictions)
-
-# Estimate treatment effects for the test sample.
-tau.hat <- predict(tau.forest, X.test)
-plot(X.test[, 1], tau.hat$predictions, ylim = range(tau.hat$predictions, 0, 2), xlab = "x", ylab = "tau", type = "l")
-lines(X.test[, 1], pmax(0, X.test[, 1]), col = 2, lty = 2)
-
-# Estimate the conditional (sic?) average treatment effect on the full sample (CATE).
-average_treatment_effect(tau.forest, target.sample = "all")
-
-# Estimate the conditional average treatment effect on the treated sample (CATT).
-average_treatment_effect(tau.forest, target.sample = "treated")
-
-# Add confidence intervals for heterogeneous treatment effects;
-# growing more trees is now recommended.
-tau.forest <- causal_forest(X, Y, W, num.trees = 4000)
-tau.hat <- predict(tau.forest, X.test, estimate.variance = TRUE)
-sigma.hat <- sqrt(tau.hat$variance.estimates)
-
-png(paste(config[["produces"]]))
-par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs = "i", yaxs = "i")
-
-plot(X.test[, 1], tau.hat$predictions,
[…]

### 2024-08-14 src/topics_causal_inf/grf/task_grf_example.py

@@ -1,27 +1,0 @@
-"""Example task for grf example in R."""
-
-import pytask
-
-from topics_causal_inf.config import BLD, SRC
-
-
-@pytask.mark.r(
-    script=SRC / "grf" / "grf_example.R",
-    serializer="yaml",
-)
-def task_grf_example_in_r(
-    produces=BLD / "grf" / "grf_example_in_r.png",
-):
-    """Produce example figure using grf in R."""
-
-
-@pytask.mark.skip()
-@pytask.mark.r(
-    script=SRC / "grf" / "bruhn_2016.R",
-    serializer="yaml",
-)
-def task_grf_example_bruhn_2016(
-    produces=BLD / "grf" / "bruhn_2016.rds",
-    bruhn_2016_data=SRC / "grf" / "bruhn2016.csv",
-):
-    """Produce example figure using grf in R."""

### 2024-08-14 src/topics_causal_inf/wa_replication/__init__.py

@@ -1,1 +1,0 @@
-"""Replication of Wager and Athey (2018) results."""

### 2024-08-14 src/topics_causal_inf/wa_replication/sim_config.py

@@ -1,47 +1,0 @@
-"""Define baseline configurations for WA 2018 replication simulations."""
-
-from typing import NamedTuple
-
-
-class SimulationDGP(NamedTuple):
-    """DGP for WA 2018 style simulations."""
-
-    dgp: str
-    n_obs: int
-    n_sim: int
-    num_trees: int
-    sample_fraction: float
-
-
-DIM_VALS = [10, 15, 30]
-
-SIMULATION2 = SimulationDGP(
-    dgp="dgp2",
-    n_obs=5_000,
-    n_sim=20,
-    num_trees=2_000,
-    sample_fraction=0.5,
-)
-SIMULATION3 = SimulationDGP(
-    dgp="dgp3",
-    n_obs=10_000,
-    n_sim=20,
-    num_trees=10_000,
-    sample_fraction=0.2,
-)
-SIMULATION4 = SimulationDGP(
-    dgp="dgp4",
-    n_obs=10_000,
-    n_sim=20,
-    num_trees=10_000,
-    sample_fraction=0.2,
-)
-SIMULATION5 = SimulationDGP(
-    dgp="dgp5",
-    n_obs=10_000,
-    n_sim=20,
-    num_trees=10_000,
-    sample_fraction=0.2,
-)
-
-SIMS_TO_RUN = [SIMULATION3, SIMULATION4, SIMULATION5]

### 2024-08-14 src/topics_causal_inf/wa_replication/task_tables.py

@@ -1,51 +1,0 @@
-"""Generate tables for the WA replication."""
-
-from pathlib import Path
-
-import pandas as pd  # type: ignore[import-untyped]
-import pyreadr  # type: ignore[import-untyped]
-import pytask
-
-from topics_causal_inf.config import BLD
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS, SIMS_TO_RUN
-
-DGP_VALS = [sim.dgp for sim in SIMS_TO_RUN]
-
-RESULTS_PATHS = {
-    dgp: [
-        BLD / "wa_replication" / "sims" / f"wa_replication_{dgp}_dim{dim}.rds"
-        for dim in DIM_VALS
-    ]
-    for dgp in DGP_VALS
-}
-
-for dgp in DGP_VALS:
-    # Generate list of all paths corresponding to a dgp and each value of dim
-    @pytask.mark.skip()
-    @pytask.task(id=f"{dgp}_table")
-    def task_wa_replication_table(
-        depends_on=RESULTS_PATHS[dgp],
-        produces: Path = BLD
-        / "wa_replication"
-        / "tables"
-        / f"wa_replication_{dgp}.tex",
-    ):
-        """Generate tables for the WA replication."""
-        # Combine all results into a single dataframe
-        dfs = [pyreadr.read_r(result)[None] for result in depends_on]
-
-        for df in dfs:
-            df.columns = ["mse"]
-
-        # To each add column with dim
-        for i, df in enumerate(dfs):
-            df["dim"] = DIM_VALS[i]
-
-        # Concatenate all dataframes
-        res = pd.concat(dfs)
-
-        # Group by dim and calculate mean and std of the mse
-        res_grouped = res.groupby("dim")["mse"].agg(["mean", "std"])
-
-        # Save table
-        res_grouped.to_latex(produces)

### 2024-08-14 src/topics_causal_inf/wa_replication/task_wa_replication.py

@@ -1,28 +1,0 @@
-"""Tasks for replication of Wager and Athey (2018) results."""
-
-from pathlib import Path
-
-import pytask
-
-from topics_causal_inf.config import BLD, SRC
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS, SIMS_TO_RUN
-
-for sim in SIMS_TO_RUN:
-    for dim in DIM_VALS:
-
-        @pytask.mark.skip()
-        @pytask.task(
-            kwargs={"dim": dim, "return_grid": False, **sim._asdict()},
-            id=f"{sim.dgp}_dim{dim}",
-        )
-        @pytask.mark.r(
-            script=SRC / "wa_replication" / "wa_replication.R",
-            serializer="yaml",
-        )
-        def task_wa_replication(
-            produces: Path = BLD
-            / "wa_replication"
-            / "sims"
-            / f"wa_replication_{sim.dgp}_dim{dim}.rds",
-        ):
-            """Run simulations for WA 2018 DGPs."""

### 2024-08-14 src/topics_causal_inf/wa_replication/task_wa_replication_grid_only.py

@@ -1,28 +1,0 @@
-"""Tasks for replication of Wager and Athey (2018) results."""
-
-from pathlib import Path
-
-import pytask
-
-from topics_causal_inf.config import BLD, SRC
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS, SIMS_TO_RUN
-
-for sim in SIMS_TO_RUN:
-    for dim in DIM_VALS:
-
-        @pytask.mark.skip()
-        @pytask.task(
-            kwargs={"dim": dim, "return_grid": True, **sim._asdict()},
-            id=f"{sim.dgp}_dim{dim}",
-        )
-        @pytask.mark.r(
-            script=SRC / "wa_replication" / "wa_replication.R",
-            serializer="yaml",
-        )
-        def task_wa_replication_grid(
-            produces: Path = BLD
-            / "wa_replication"
-            / "sims"
-            / f"wa_replication_{sim.dgp}_dim{dim}_grid_only.rds",
-        ):
-            """Produce example figure using grf in R."""

### 2024-08-14 src/topics_causal_inf/wa_replication/wa_replication.R

@@ -1,187 +1,0 @@
-library(grf)
-
-# ======================================================================================
-# Functions
-# ======================================================================================
-
-# Simulation
-simulation <- function(
-    n_sim, n_obs, dim, main_effect, treatment_effect, propensity, num_trees,
-    sample_fraction, return_grid = FALSE) {
-  # Run experiments
-  if (return_grid == TRUE) {
-    return(experiment(n_obs, dim, main_effect, treatment_effect, propensity, num_trees,
-      sample_fraction,
-      return_grid = return_grid
-    ))
-  }
-
-  mse <- rep(0, n_sim)
-  for (i in 1:n_sim) {
-    mse[i] <- experiment(n_obs, dim, main_effect, treatment_effect, propensity,
-      num_trees,
-      sample_fraction,
-      return_grid = return_grid
-    )
-  }
-
-  return(mse)
-}
-
-# Single experiment
-experiment <- function(
-    n_obs, dim, main_effect, treatment_effect, propensity, num_trees, sample_fraction,
-    return_grid = FALSE) {
-  # Draw data
-  data <- data(n_obs, dim, main_effect, treatment_effect, propensity)
-  data_test <- data(n_obs, dim, main_effect, treatment_effect, propensity)
-
-  # Fit causal forest
-  cf <- causal_forest(data$x, data$y, data$w,
-    num.trees = num_trees,
-    sample.fraction = sample_fraction
-  )
-
-  if (!return_grid) {
-    # Predict on test-data
-    pred <- predict(cf, data_test$x)
-    true <- treatment_effect(data_test$x)
-
-    # Evaluate fit by mse
-    mse <- mean((pred$predictions - true)^2)
-
-    # Make prediction on x1, x2 from 0 to 1
-    return(mse)
-  }
-
-  data_eval <- data(10000, dim, main_effect, treatment_effect, propensity)
-
-  pred <- predict(cf, data_eval$x)
-
-  pred_eval <- pred$predictions
-  true_eval <- treatment_effect(data_eval$x)
-
-  # Put into one dataframe
-  mse <- mean((pred_eval - true_eval)^2)
-
[…]

### 2024-08-14 src/topics_causal_inf/wa_replication/wa_replication_debug.R

@@ -1,148 +1,0 @@
-library(grf)
-
-# ======================================================================================
-# Functions
-# ======================================================================================
-
-# Simulation
-simulation <- function(
-    n_sim, n_obs, dim, main_effect, treatment_effect, propensity,
-    return_grid = FALSE) {
-  # Run experiments
-  if (return_grid == TRUE) {
-    return(experiment(n_obs, dim, main_effect, treatment_effect, propensity,
-      return_grid = return_grid
-    ))
-  }
-
-  mse <- rep(0, n_sim)
-  for (i in 1:n_sim) {
-    mse[i] <- experiment(n_obs, dim, main_effect, treatment_effect, propensity,
-      return_grid = return_grid
-    )
-  }
-
-  return(mse)
-}
-
-# Single experiment
-experiment <- function(
-    n_obs, dim, main_effect, treatment_effect, propensity,
-    return_grid = FALSE) {
-  # Draw data
-  data <- data(n_obs, dim, main_effect, treatment_effect, propensity)
-  data_test <- data(n_obs, dim, main_effect, treatment_effect, propensity)
-
-  # Fit causal forest
-  cf <- causal_forest(data$x, data$y, data$w, num.trees = 2000, sample.fraction = 0.5)
-
-  if (!return_grid) {
-    # Predict on test-data
-    pred <- predict(cf, data_test$x)
-    true <- treatment_effect(data_test$x)
-
-    # Evaluate fit by mse
-    mse <- mean((pred$predictions - true)^2)
-
-    # Make prediction on x1, x2 from 0 to 1
-    return(mse)
-  }
-
-  data_eval <- data(10000, dim, main_effect, treatment_effect, propensity)
-
-  pred <- predict(cf, data_eval$x)
-
-  pred_grid <- pred$predictions
-  true_grid <- treatment_effect(data_eval$x)
-
-  # Put into one dataframe
-  mse <- mean((pred_grid - true_grid)^2)
-
-  out <- cbind(data_eval$x, pred_grid, true_grid, mse)
-
-  colnames(out) <- c(paste0("x", 1:dim), "pred", "true", "mse")
-
-  return(out)
-}
-
-# Generate data
-data <- function(n_obs, dim, main_effect, treatment_effect, propensity) {
-  x <- matrix(0, n_obs, dim)
[…]

### 2024-08-14 src/topics_causal_inf/wgan/__init__.py

@@ -1,1 +1,0 @@
-"""Design of Simulations using WGAN using ds-wgan package."""

### 2024-08-14 src/topics_causal_inf/classes.py
Add GATES to generic_ml.

@@ -19,6 +19,12 @@ class GenericMLResult:
     blp_ci_hi: np.ndarray
     ml_fitted_d0: RegressorMixin
     ml_fitted_d1: RegressorMixin
+    gates_params: np.ndarray
+    gates_se: np.ndarray
+    gates_tvals: np.ndarray
+    gates_pvals: np.ndarray
+    gates_ci_lo: np.ndarray
+    gates_ci_hi: np.ndarray
 
 
 class DGP(NamedTuple):

### 2024-08-14 src/topics_causal_inf/config.py

@@ -4,6 +4,8 @@ from pathlib import Path
 
 import numpy as np
 
+from topics_causal_inf.define_dgps import DGP3, DGP4, DGP5
+
 SRC = Path(__file__).parent.resolve()
 ROOT = SRC.joinpath("..", "..").resolve()
 
@@ -16,3 +18,7 @@ DOCUMENTS = ROOT.joinpath("documents").resolve()
 TEMPLATE_GROUPS = ["marital_status", "highest_qualification"]
 
 RNG = np.random.default_rng()
+
+DIMS_TO_RUN = [10, 15, 30]
+
+DGPS_TO_RUN = [DGP3, DGP4, DGP5]

### 2024-08-14 src/topics_causal_inf/generic_ml/example_GenericML.R

@@ -1,123 +1,0 @@
-# install.packages("GenericML")
-# install.packages("glmnet")
-# install.packages("ranger")
-# install.packages("e1071")
-
-library("GenericML")
-
-n <- 1000
-p <- 3
-d <- rbinom(n, 1, 0.5)
-z <- matrix(runif(n * p), n, p)
-
-colnames(z) <- paste("z", 1:p)
-
-y0 <- as.numeric(z %*% rexp(p))
-
-# treatment effect increases with Z1, has a level shift along Z2 (at 0.5)
-# and has no pattern along Z3
-cate <- 2 * z[, 1] + ifelse(z[, 2] >= 0.5, 1, -1)
-ate <- mean(cate) # average treatment effect
-y1 <- cate + y0 # potential outcome under treatment
-y <- ifelse(d == 1, y1, y0) # observed outcome
-
-# ======================================================================================
-# 2 Prepare arguments for GenericML function
-# ======================================================================================
-# quantile cutoffs for the GATES grouping of the estimated CATEs
-quantile_cutoffs <- c(0.25, 0.5, 0.75)
-
-# specify the learner of the propensity score (non-penalized logistic regression here).
-# Propensity scores can also directly be supplied.
-learner_propensity_score <- "mlr3::lrn('glmnet', lambda = 0, alpha = 1)"
-
-# specify the considered learners of the BCA and the CATE
-# (here: lasso, random forest, and SVM)
-learners_genericml <- c(
-  "lasso", "mlr3::lrn('ranger', num.trees = 100)",
-  "mlr3::lrn('svm')"
-)
-
-# specify the data that shall be used for the CLAN
-# here, we use all variables of Z and uniformly distributed random noise
-z_clan <- cbind(z, random = runif(n))
-
-# specify the number of splits (many to rule out seed-dependence of results)
-num_splits <- 1000
-
-# specify if a HT transformation shall be used when estimating BLP and GATES
-ht <- FALSE
-
-# A list controlling the vars used in the matrix X1 for the BLP and GATES regressions.
-x1_blp <- setup_X1()
-x1_gates <- setup_X1()
-
-# consider differences between group K (most affected) with groups 1, 2, and 3.
-diff_gates <- setup_diff(
-  subtract_from = "most",
-  subtracted = 1:3
-)
-diff_clan <- setup_diff(
-  subtract_from = "most",
-  subtracted = 1:3
-)
-
-# specify the significance level
-significance_level <- 0.05
-
[…]

### 2024-08-14 src/topics_causal_inf/generic_ml/generic_ml.py

@@ -28,20 +28,28 @@ def generic_ml(
     # Prepare inputs
 
     # Inference algorithm
-    blp_params, blp_se, ml_fitted_d0, ml_fitted_d1 = inference_algorithm(
-        data,
-        n_splits,
-        strategy,
-        ml_learner,
+    blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se = (
+        inference_algorithm(
+            data,
+            n_splits,
+            strategy,
+            ml_learner,
+        )
     )
 
-    # Compute t-values, lower and upper (1-alpha) confidence intervals
+    # BLP Compute t-values, lower and upper (1-alpha) confidence intervals
     blp_tvals = blp_params / blp_se
     blp_pvals = 2 * (1 - norm.cdf(np.abs(blp_tvals)))
     blp_ci_lo = blp_params - norm.ppf(1 - alpha / 2) * blp_se
     blp_ci_hi = blp_params + norm.ppf(1 - alpha / 2) * blp_se
 
-    # Find position of median of alpha_1 to pick a "median" machine learner
+    # GATES: Compute t-values, lower and upper (1-alpha) confidence intervals
+    gates_tvals = gates_params / gates_se
+    gates_pvals = 2 * (1 - norm.cdf(np.abs(gates_tvals)))
+    gates_ci_lo = gates_params - norm.ppf(1 - alpha / 2) * gates_se
+    gates_ci_hi = gates_params + norm.ppf(1 - alpha / 2) * gates_se
+
+    # Find position of median of alpha_2 to pick a "median" machine learner
     # Choose closest observation to force median is element of the set
     pos = np.argwhere(
         blp_params[:, 1]
@@ -58,6 +66,12 @@ def generic_ml(
         blp_ci_hi=np.median(blp_ci_hi, axis=0),
         ml_fitted_d0=ml_fitted_d0[int(pos)],
         ml_fitted_d1=ml_fitted_d1[int(pos)],
+        gates_params=np.median(gates_params, axis=0),
+        gates_se=np.median(gates_se, axis=0),
+        gates_tvals=np.median(gates_tvals, axis=0),
+        gates_pvals=np.median(gates_pvals, axis=0),
+        gates_ci_lo=np.median(gates_ci_lo, axis=0),
+        gates_ci_hi=np.median(gates_ci_hi, axis=0),
     )
 
 
@@ -66,13 +80,22 @@ def inference_algorithm(
[…]

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_sim.py

@@ -10,16 +10,12 @@ from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
 from sklearn.ensemble import RandomForestRegressor  # type: ignore[import-untyped]
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, RNG
-from topics_causal_inf.define_dgps import DGP3, DGP4
+from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN, RNG
 from topics_causal_inf.generic_ml.simulation import simulation
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS
 
 N_OBS = 10_000
 N_SIMS = 20
 N_SPLITS = 25
-
-DGP_TO_RUN = [DGP3, DGP4]
 
 
 class _Arguments(NamedTuple):
@@ -43,8 +39,8 @@ ID_TO_KWARGS = {
             CausalForest(n_estimators=100, max_samples=0.2),
         ),
     )
-    for dgp in DGP_TO_RUN
-    for dim in DIM_VALS
+    for dgp in DGPS_TO_RUN
+    for dim in DIMS_TO_RUN
     if not (dgp.name == "dgp5" and dim < 8)  # noqa: PLR2004
 }
 

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -6,9 +6,7 @@ from typing import Annotated, NamedTuple
 import pandas as pd  # type: ignore[import-untyped]
 from pytask import Product, task
 
-from topics_causal_inf.config import BLD
-from topics_causal_inf.generic_ml.task_generic_ml_sim import DGP_TO_RUN
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS
+from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN
 
 
 class _Arguments(NamedTuple):
@@ -22,10 +20,10 @@ ID_TO_KWARGS = {
         path_to_table=BLD / "generic_ml" / "tables" / f"generic_ml_{dgp.name}.tex",
         path_to_res=[
             BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl"
-            for dim in DIM_VALS
+            for dim in DIMS_TO_RUN
         ],
     )
-    for dgp in DGP_TO_RUN
+    for dgp in DGPS_TO_RUN
 }
 
 for id_, kwargs in ID_TO_KWARGS.items():

### 2024-08-14 src/topics_causal_inf/py_wa_replication/task_py_wa_replication.py

@@ -9,14 +9,11 @@ import pandas as pd  # type: ignore[import-untyped]
 from pytask import Product, task
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, RNG, WGAN_GEN
+from topics_causal_inf.config import BLD, DIMS_TO_RUN, RNG, WGAN_GEN
 from topics_causal_inf.define_dgps import DGP3, DGP4
 from topics_causal_inf.py_wa_replication.py_wa_replication import simulation
-from topics_causal_inf.wa_replication.sim_config import DIM_VALS
 
 DGPS_TO_RUN = [DGP3, DGP4]
-
-DIMS_TO_RUN = DIM_VALS
 
 
 class _Arguments(NamedTuple):

### 2024-08-14 src/topics_causal_inf/define_dgps.py
Add gates to simulation.

@@ -47,4 +47,11 @@ DGP5 = DGP(
     treatment_effect=partial(tau_heterog, x_range=np.arange(0, 8), a=12, b=1 / 2),
 )
 
-DGPS = [DGP1, DGP2, DGP3, DGP4, DGP5]
+DGP6 = DGP(
+    name="dgp6",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_constant, tau=0),
+)
+
+DGPS = [DGP1, DGP2, DGP3, DGP4, DGP5, DGP6]

### 2024-08-14 src/topics_causal_inf/generic_ml/simulation.py

@@ -22,10 +22,11 @@ def simulation(
     rng: np.random.Generator,
 ) -> pd.DataFrame:
     """Run simulation using the generic_ml approach and WA 2018 DGP."""
-    mse = np.zeros(n_sims)
+    mse_blp = np.zeros(n_sims)
+    gates = np.zeros((n_sims, 5))
 
     for i in range(n_sims):
-        mse[i] = _single_experiment(
+        mse_blp[i], gates[i, :] = _single_experiment(
             n_obs=n_obs,
             dim=dim,
             n_splits=n_splits,
@@ -34,12 +35,18 @@ def simulation(
             rng=rng,
         )
 
-    out = pd.DataFrame(mse, columns=["mse"])
+    out = pd.DataFrame(mse_blp, columns=["mse_blp"])
 
     out["dgp"] = dgp.name
     out["n_obs"] = n_obs
     out["n_sims"] = n_sims
     out["dim"] = dim
+
+    true_gates = _simulate_true_gates(dgp=dgp, dim=dim, rng=rng)
+
+    for i in range(5):
+        out[f"gate_{i}"] = gates[:, i]
+        out[f"true_gate_{i}"] = true_gates[i]
 
     return out
 
@@ -75,4 +82,27 @@ def _single_experiment(
 
     data_eval["true"] = dgp.treatment_effect(data_eval)
 
-    return np.mean((data_eval["pred"] - data_eval["true"]) ** 2)
+    return np.mean((data_eval["pred"] - data_eval["true"]) ** 2), res.gates_params
+
+
+def _simulate_true_gates(
+    dgp: DGP,
+    dim: int,
+    rng: np.random.Generator,
+    n_obs: int = 1_000_000,
+) -> np.ndarray:
+    """Simulate true gates for the given DGP."""
+    data = data_wager_athey_2018(
+        n_obs=n_obs,
+        dim=dim,
+        dgp=dgp,
+        rng=rng,
+    )
+
+    feature_names = [f"x{i}" for i in range(dim)]
+
+    data["s_z"] = dgp.treatment_effect(data[feature_names])
+
+    data["s_z_quintiles"] = pd.qcut(data["s_z"], q=5, labels=np.arange(5))
+
+    return data.groupby("s_z_quintiles").apply("mean")["s_z"]

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -41,8 +41,8 @@ for id_, kwargs in ID_TO_KWARGS.items():
         table = res.groupby(["dim"]).apply(
             lambda x: pd.Series(
                 {
-                    "mean": x["mse"].mean(),
-                    "std": x["mse"].std(),
+                    "mean": x["mse_blp"].mean(),
+                    "std": x["mse_blp"].std(),
                 },
             ),
         )

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_plot_gates.py
Add task for gates plots.

@@ -1,0 +1,61 @@
+from pathlib import Path
+from typing import Annotated
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+import plotly.express as px  # type: ignore[import-untyped]
+import plotly.graph_objects as go  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.config import BLD, DGPS_TO_RUN
+from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
+
+PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
+
+
+ID_TO_KWARGS_PLOTS = {
+    dgp.name: {
+        "path_to_plot": BLD
+        / "generic_ml"
+        / "plots"
+        / f"generic_ml_gates_{dgp.name}.png",
+    }
+    for dgp in DGPS_TO_RUN
+}
+
+for _id, kwargs in ID_TO_KWARGS_PLOTS.items():
+
+    @task(id=_id, kwargs=kwargs)
+    def task_generic_ml_plot_gates(
+        path_to_plot: Annotated[Path, Product],
+        path_to_results: list[Path] = PATHS_TO_RESULTS,
+    ):
+        """Task for generic_ml plot gates."""
+        # Load results
+        res = pd.concat([pd.read_pickle(path) for path in path_to_results])
+
+        # Plot Figure
+        fig = go.Figure()
+
+        colors = px.colors.qualitative.Plotly
+
+        for i in range(5):
+            color = colors[i % len(colors)]  # Cycle through the color palett
+            fig.add_trace(
+                go.Histogram(
+                    x=res[f"gate_{i}"],
+                    histnorm="probability",
+                    name=f"gate_{i}",
+                    marker_color=color,
+                ),
+            )
+            fig.add_vline(
+                x=np.mean(res[f"gate_{i}"]),
+                line_dash="dash",
+                line_color=color,
+            )
+
+        fig.update_layout(barmode="overlay")
+        fig.update_traces(opacity=0.75)
+
+        fig.write_image(path_to_plot)

### 2024-08-14 src/topics_causal_inf/config.py
Move all simulation settings to config.

@@ -22,3 +22,9 @@ RNG = np.random.default_rng()
 DIMS_TO_RUN = [10, 15, 30]
 
 DGPS_TO_RUN = [DGP3, DGP4, DGP5]
+
+NUM_TREES = 100
+SUBSAMPLE_SHARE = 0.2
+N_OBS = 10_000
+N_SIM = 20
+N_SPLITS = 10

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_sim.py

@@ -10,12 +10,18 @@ from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
 from sklearn.ensemble import RandomForestRegressor  # type: ignore[import-untyped]
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN, RNG
+from topics_causal_inf.config import (
+    BLD,
+    DGPS_TO_RUN,
+    DIMS_TO_RUN,
+    N_OBS,
+    N_SIM,
+    N_SPLITS,
+    NUM_TREES,
+    RNG,
+    SUBSAMPLE_SHARE,
+)
 from topics_causal_inf.generic_ml.simulation import simulation
-
-N_OBS = 10_000
-N_SIMS = 20
-N_SPLITS = 25
 
 
 class _Arguments(NamedTuple):
@@ -24,7 +30,7 @@ class _Arguments(NamedTuple):
     path_to_res: Path
     ml_learner: tuple[RegressorMixin, RegressorMixin]
     n_obs: int = N_OBS
-    n_sims: int = N_SIMS
+    n_sims: int = N_SIM
     n_splits: int = N_SPLITS
     rng: np.random.Generator = RNG
 
@@ -35,8 +41,8 @@ ID_TO_KWARGS = {
         dgp=dgp,
         path_to_res=BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl",
         ml_learner=(
-            RandomForestRegressor(n_estimators=100, max_samples=0.2),
-            CausalForest(n_estimators=100, max_samples=0.2),
+            RandomForestRegressor(n_estimators=NUM_TREES, max_samples=SUBSAMPLE_SHARE),
+            CausalForest(n_estimators=NUM_TREES, max_samples=SUBSAMPLE_SHARE),
         ),
     )
     for dgp in DGPS_TO_RUN

### 2024-08-14 src/topics_causal_inf/py_wa_replication/task_py_wa_replication.py

@@ -9,23 +9,30 @@ import pandas as pd  # type: ignore[import-untyped]
 from pytask import Product, task
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, DIMS_TO_RUN, RNG, WGAN_GEN
-from topics_causal_inf.define_dgps import DGP3, DGP4
+from topics_causal_inf.config import (
+    BLD,
+    DGPS_TO_RUN,
+    DIMS_TO_RUN,
+    N_OBS,
+    N_SIM,
+    NUM_TREES,
+    RNG,
+    SUBSAMPLE_SHARE,
+    WGAN_GEN,
+)
 from topics_causal_inf.py_wa_replication.py_wa_replication import simulation
-
-DGPS_TO_RUN = [DGP3, DGP4]
 
 
 class _Arguments(NamedTuple):
     dgp: DGP
     path_to_data: Path
     data_generator: str
-    n_sim: int = 20
-    n_obs: int = 10_000
+    n_sim: int = N_SIM
+    n_obs: int = N_OBS
     dim: int = 10
     rng: np.random.Generator = RNG
-    num_trees: int = 100
-    sample_fraction: float = 0.2
+    num_trees: int = NUM_TREES
+    sample_fraction: float = SUBSAMPLE_SHARE
 
 
 DG_TO_RUN = ["standard", "wgan"]

### 2024-08-14 src/topics_causal_inf/classes.py
Correctly set true GATES for DGPs with constant CATE.

@@ -1,3 +1,4 @@
+import functools
 from collections.abc import Callable
 from dataclasses import dataclass
 from typing import NamedTuple
@@ -31,6 +32,6 @@ class DGP(NamedTuple):
     """DGP for simulation."""
 
     name: str
-    main_effect: Callable[[pd.DataFrame], np.ndarray]
-    treatment_effect: Callable[[pd.DataFrame], np.ndarray]
-    propensity_score: Callable[[pd.DataFrame], np.ndarray]
+    main_effect: Callable[[pd.DataFrame], np.ndarray] | functools.partial
+    treatment_effect: Callable[[pd.DataFrame], np.ndarray] | functools.partial
+    propensity_score: Callable[[pd.DataFrame], np.ndarray] | functools.partial

### 2024-08-14 src/topics_causal_inf/config.py

@@ -4,7 +4,7 @@ from pathlib import Path
 
 import numpy as np
 
-from topics_causal_inf.define_dgps import DGP3, DGP4, DGP5
+from topics_causal_inf.define_dgps import DGP3, DGP4, DGP5, DGP6
 
 SRC = Path(__file__).parent.resolve()
 ROOT = SRC.joinpath("..", "..").resolve()
@@ -21,10 +21,10 @@ RNG = np.random.default_rng()
 
 DIMS_TO_RUN = [10, 15, 30]
 
-DGPS_TO_RUN = [DGP3, DGP4, DGP5]
+DGPS_TO_RUN = [DGP3, DGP4, DGP5, DGP6]
 
 NUM_TREES = 100
 SUBSAMPLE_SHARE = 0.2
 N_OBS = 10_000
-N_SIM = 20
-N_SPLITS = 10
+N_SIM = 2
+N_SPLITS = 2

### 2024-08-14 src/topics_causal_inf/generic_ml/simulation.py

@@ -92,6 +92,10 @@ def _simulate_true_gates(
     n_obs: int = 1_000_000,
 ) -> np.ndarray:
     """Simulate true gates for the given DGP."""
+    # Check whether dgp.treatment_effect.keywords["tau"] raises error
+    if "tau" in dgp.treatment_effect.keywords:  # type: ignore[union-attr]
+        return np.ones(5) * dgp.treatment_effect.keywords["tau"]  # type: ignore[union-attr]
+
     data = data_wager_athey_2018(
         n_obs=n_obs,
         dim=dim,

### 2024-08-14 src/topics_causal_inf/py_wa_replication/task_py_wa_replication.py

@@ -50,7 +50,7 @@ ID_TO_KWARGS = {
     for dgp in DGPS_TO_RUN
     for dim in DIMS_TO_RUN
     for dgen in DG_TO_RUN
-    if not (dgen == "wgan" and dim != 10)  # noqa: PLR2004
+    if not (dgen == "wgan" and (dim != 10 or dgp.name == "dgp6"))  # noqa: PLR2004
 }
 
 for id_, kwargs in ID_TO_KWARGS.items():

### 2024-08-14 tests/generic_ml/test_simulation.py

@@ -1,0 +1,9 @@
+from topics_causal_inf.config import RNG
+from topics_causal_inf.define_dgps import DGP3, DGP4, DGP5, DGP6
+from topics_causal_inf.generic_ml.simulation import _simulate_true_gates
+
+
+def test_simulate_true_gates_runs() -> None:
+    for dgp in [DGP3, DGP4, DGP5, DGP6]:
+        for dim in [10, 15, 30]:
+            _simulate_true_gates(dgp=dgp, dim=dim, rng=RNG)

### 2024-08-14 documents/task_documents.py
Add CLAN to generic_ml.

@@ -10,6 +10,7 @@ documents = ["paper"]
 
 for document in documents:
 
+    @pytask.mark.skip()
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
         document=BLD / "documents" / f"{document}.pdf",
@@ -26,6 +27,7 @@ for document in documents:
         "produces": ROOT / f"{document}.pdf",
     }
 
+    @pytask.mark.skip()
     @pytask.task(id=document, kwargs=kwargs)
     def task_copy_to_root(depends_on, produces):
         """Copy a document to the root directory for easier retrieval."""

### 2024-08-14 src/topics_causal_inf/classes.py

@@ -26,6 +26,7 @@ class GenericMLResult:
     gates_pvals: np.ndarray
     gates_ci_lo: np.ndarray
     gates_ci_hi: np.ndarray
+    clan: np.ndarray
 
 
 class DGP(NamedTuple):

### 2024-08-14 src/topics_causal_inf/config.py

@@ -26,5 +26,5 @@ DGPS_TO_RUN = [DGP3, DGP4, DGP5, DGP6]
 NUM_TREES = 100
 SUBSAMPLE_SHARE = 0.2
 N_OBS = 10_000
-N_SIM = 2
-N_SPLITS = 2
+N_SIM = 20
+N_SPLITS = 10

### 2024-08-14 src/topics_causal_inf/generic_ml/generic_ml.py

@@ -28,7 +28,7 @@ def generic_ml(
     # Prepare inputs
 
     # Inference algorithm
-    blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se = (
+    blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se, clan = (
         inference_algorithm(
             data,
             n_splits,
@@ -72,6 +72,7 @@ def generic_ml(
         gates_pvals=np.median(gates_pvals, axis=0),
         gates_ci_lo=np.median(gates_ci_lo, axis=0),
         gates_ci_hi=np.median(gates_ci_hi, axis=0),
+        clan=np.median(clan, axis=0),
     )
 
 
@@ -87,8 +88,13 @@ def inference_algorithm(
     list[RegressorMixin],
     np.ndarray,
     np.ndarray,
+    np.ndarray,
 ]:
     """Implements the inference algorithm; see Algorithm 1 in the paper."""
+    pattern = r"^x\d+$"
+    feature_names = data.filter(regex=pattern).columns.tolist()
+    dim = len(feature_names)
+
     blp_params = np.zeros((n_splits, 2))
     blp_se = np.zeros((n_splits, 2))
     lambda_hat = np.zeros(n_splits)
@@ -96,6 +102,7 @@ def inference_algorithm(
     ml_fitted_d1 = list(np.zeros(n_splits))
     gates_params = np.zeros((n_splits, 5))
     gates_se = np.zeros((n_splits, 5))
+    clan = np.zeros((n_splits, 2, dim))
 
     for i in range(n_splits):
         (
@@ -106,13 +113,14 @@ def inference_algorithm(
             ml_fitted_d1[i],
             gates_params[i, :],
             gates_se[i, :],
+            clan[i, :, :],
         ) = estimate_single_split(
             data,
             strategy,
             ml_learner,
         )
 
-    return blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se
+    return blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se, clan
 
 
 def estimate_single_split(
@@ -125,6 +133,7 @@ def estimate_single_split(
     float,
     RegressorMixin,
     RegressorMixin,
+    np.ndarray,
     np.ndarray,
     np.ndarray,
 ]:
@@ -159,6 +168,9 @@ def estimate_single_split(
     gates = gates_ht_transform(main)
[…]

### 2024-08-14 src/topics_causal_inf/generic_ml/simulation.py

@@ -6,7 +6,7 @@ from econml.grf import CausalForest  # type: ignore[import-untyped]
 from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.generic_ml.generic_ml import generic_ml
+from topics_causal_inf.generic_ml.generic_ml import classification_analysis, generic_ml
 from topics_causal_inf.utilities import (
     data_wager_athey_2018,
 )
@@ -24,9 +24,10 @@ def simulation(
     """Run simulation using the generic_ml approach and WA 2018 DGP."""
     mse_blp = np.zeros(n_sims)
     gates = np.zeros((n_sims, 5))
+    clans = np.zeros((n_sims, 2, dim))
 
     for i in range(n_sims):
-        mse_blp[i], gates[i, :] = _single_experiment(
+        mse_blp[i], gates[i, :], clans[i, :, :] = _single_experiment(
             n_obs=n_obs,
             dim=dim,
             n_splits=n_splits,
@@ -47,6 +48,13 @@ def simulation(
     for i in range(5):
         out[f"gate_{i}"] = gates[:, i]
         out[f"true_gate_{i}"] = true_gates[i]
+
+    true_clans = _simulate_true_clan(dgp=dgp, dim=dim, rng=rng)
+
+    for i in range(2):
+        for j in range(dim):
+            out[f"clan_{i}_{j}"] = clans[:, i, j]
+            out[f"true_clan_{i}_{j}"] = true_clans[i, j]
 
     return out
 
@@ -82,7 +90,11 @@ def _single_experiment(
 
     data_eval["true"] = dgp.treatment_effect(data_eval)
 
-    return np.mean((data_eval["pred"] - data_eval["true"]) ** 2), res.gates_params
+    return (
+        np.mean((data_eval["pred"] - data_eval["true"]) ** 2),
+        res.gates_params,
+        res.clan,
+    )
 
 
 def _simulate_true_gates(
@@ -109,4 +121,29 @@ def _simulate_true_gates(
 
     data["s_z_quintiles"] = pd.qcut(data["s_z"], q=5, labels=np.arange(5))
 
-    return data.groupby("s_z_quintiles").apply("mean")["s_z"]
+    return data.groupby("s_z_quintiles", observed=False).apply("mean")["s_z"]
+
+
+def _simulate_true_clan(
+    dgp: DGP,
+    dim: int,
[…]

### 2024-08-14 src/topics_causal_inf/generic_ml/task_generic_ml_plot_gates.py

@@ -7,6 +7,7 @@ import plotly.express as px  # type: ignore[import-untyped]
 import plotly.graph_objects as go  # type: ignore[import-untyped]
 from pytask import Product, task
 
+from topics_causal_inf.classes import DGP
 from topics_causal_inf.config import BLD, DGPS_TO_RUN
 from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
 
@@ -15,6 +16,7 @@ PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
 
 ID_TO_KWARGS_PLOTS = {
     dgp.name: {
+        "dgp": dgp,
         "path_to_plot": BLD
         / "generic_ml"
         / "plots"
@@ -27,12 +29,15 @@ for _id, kwargs in ID_TO_KWARGS_PLOTS.items():
 
     @task(id=_id, kwargs=kwargs)
     def task_generic_ml_plot_gates(
+        dgp: DGP,
         path_to_plot: Annotated[Path, Product],
         path_to_results: list[Path] = PATHS_TO_RESULTS,
     ):
         """Task for generic_ml plot gates."""
         # Load results
         res = pd.concat([pd.read_pickle(path) for path in path_to_results])
+
+        res = res[res["dgp"] == dgp.name]
 
         # Plot Figure
         fig = go.Figure()

### 2024-08-15 src/topics_causal_inf/generic_ml/simulation.py
Correct dim error and change paths.

@@ -133,7 +133,7 @@ def _simulate_true_clan(
     """Simulate true CLAN for the given DGP."""
     # Check whether dgp.treatment_effect.keywords["tau"] raises error
     if "tau" in dgp.treatment_effect.keywords:  # type: ignore[union-attr]
-        return np.ones(2) * dgp.treatment_effect.keywords["tau"]  # type: ignore[union-attr]
+        return np.ones((2, dim)) * dgp.treatment_effect.keywords["tau"]  # type: ignore[union-attr]
 
     data = data_wager_athey_2018(
         n_obs=n_obs,

### 2024-08-15 src/topics_causal_inf/py_wa_replication/task_py_wa_replication_tables.py

@@ -26,7 +26,10 @@ ID_TO_KWARGS_TABLES = {
     f"{dgp.name}": _Arguments(
         dgp=dgp,
         path_to_results=PATH_TO_RESULTS,
-        path_to_table=BLD / "tables" / f"py_wa_replication_{dgp.name}.tex",
+        path_to_table=BLD
+        / "py_wa_replication"
+        / "tables"
+        / f"py_wa_replication_{dgp.name}.tex",
     )
     for dgp in DGPS_TO_RUN
 }

### 2024-08-15 src/topics_causal_inf/generic_ml/simulation.py
Add plots for clans and correct true characteristics to 0.5 if tau is constant.

@@ -133,7 +133,7 @@ def _simulate_true_clan(
     """Simulate true CLAN for the given DGP."""
     # Check whether dgp.treatment_effect.keywords["tau"] raises error
     if "tau" in dgp.treatment_effect.keywords:  # type: ignore[union-attr]
-        return np.ones((2, dim)) * dgp.treatment_effect.keywords["tau"]  # type: ignore[union-attr]
+        return np.ones((2, dim)) * 0.5  # type: ignore[union-attr]
 
     data = data_wager_athey_2018(
         n_obs=n_obs,

### 2024-08-15 src/topics_causal_inf/generic_ml/task_generic_ml_plot_gates.py

@@ -1,66 +1,0 @@
-from pathlib import Path
-from typing import Annotated
-
-import numpy as np
-import pandas as pd  # type: ignore[import-untyped]
-import plotly.express as px  # type: ignore[import-untyped]
-import plotly.graph_objects as go  # type: ignore[import-untyped]
-from pytask import Product, task
-
-from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, DGPS_TO_RUN
-from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
-
-PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
-
-
-ID_TO_KWARGS_PLOTS = {
-    dgp.name: {
-        "dgp": dgp,
-        "path_to_plot": BLD
-        / "generic_ml"
-        / "plots"
-        / f"generic_ml_gates_{dgp.name}.png",
-    }
-    for dgp in DGPS_TO_RUN
-}
-
-for _id, kwargs in ID_TO_KWARGS_PLOTS.items():
-
-    @task(id=_id, kwargs=kwargs)
-    def task_generic_ml_plot_gates(
-        dgp: DGP,
-        path_to_plot: Annotated[Path, Product],
-        path_to_results: list[Path] = PATHS_TO_RESULTS,
-    ):
-        """Task for generic_ml plot gates."""
-        # Load results
-        res = pd.concat([pd.read_pickle(path) for path in path_to_results])
-
-        res = res[res["dgp"] == dgp.name]
-
-        # Plot Figure
-        fig = go.Figure()
-
-        colors = px.colors.qualitative.Plotly
-
-        for i in range(5):
-            color = colors[i % len(colors)]  # Cycle through the color palett
-            fig.add_trace(
-                go.Histogram(
-                    x=res[f"gate_{i}"],
-                    histnorm="probability",
-                    name=f"gate_{i}",
-                    marker_color=color,
-                ),
-            )
-            fig.add_vline(
-                x=np.mean(res[f"gate_{i}"]),
-                line_dash="dash",
-                line_color=color,
-            )
-
-        fig.update_layout(barmode="overlay")
-        fig.update_traces(opacity=0.75)
-
-        fig.write_image(path_to_plot)

### 2024-08-15 src/topics_causal_inf/generic_ml/task_generic_ml_plots.py

@@ -1,0 +1,152 @@
+from pathlib import Path
+from typing import Annotated
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+import plotly.express as px  # type: ignore[import-untyped]
+import plotly.graph_objects as go  # type: ignore[import-untyped]
+from plotly.subplots import make_subplots  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.config import BLD, DGPS_TO_RUN
+from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
+
+PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
+
+
+ID_TO_KWARGS_GATES = {
+    dgp.name: {
+        "dgp": dgp,
+        "path_to_plot": BLD
+        / "generic_ml"
+        / "plots"
+        / f"generic_ml_gates_{dgp.name}.png",
+    }
+    for dgp in DGPS_TO_RUN
+}
+
+for _id, kwargs in ID_TO_KWARGS_GATES.items():
+
+    @task(id=_id, kwargs=kwargs)
+    def task_generic_ml_plot_gates(
+        dgp: DGP,
+        path_to_plot: Annotated[Path, Product],
+        path_to_results: list[Path] = PATHS_TO_RESULTS,
+    ):
+        """Task for generic_ml plot gates."""
+        # Load results
+        res = pd.concat([pd.read_pickle(path) for path in path_to_results])
+
+        res = res[res["dgp"] == dgp.name]
+
+        # Plot Figure
+        fig = go.Figure()
+
+        colors = px.colors.qualitative.Plotly
+
+        for i in range(5):
+            color = colors[i % len(colors)]  # Cycle through the color palett
+            fig.add_trace(
+                go.Histogram(
+                    x=res[f"gate_{i}"],
+                    histnorm="probability",
+                    name=f"gate_{i}",
+                    marker_color=color,
+                ),
+            )
+            fig.add_vline(
+                x=np.mean(res[f"true_gate_{i}"]),
+                line_dash="dash",
+                line_color=color,
+            )
+
+        fig.update_layout(barmode="overlay")
+        fig.update_traces(opacity=0.75)
+
+        fig.write_image(path_to_plot)
+
+
[…]

### 2024-08-15 documents/task_documents.py
Fix latex compilation.

@@ -10,12 +10,17 @@ documents = ["paper"]
 
 for document in documents:
 
-    @pytask.mark.skip()
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
         document=BLD / "documents" / f"{document}.pdf",
         compilation_steps=cs.latexmk(
-            options=("--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd"),
+            options=(
+                "--pdf",
+                "--interaction=nonstopmode",
+                "--synctex=1",
+                "--cd",
+                "--f",
+            ),
         ),
     )
     @pytask.task(id=document)
@@ -27,7 +32,6 @@ for document in documents:
         "produces": ROOT / f"{document}.pdf",
     }
 
-    @pytask.mark.skip()
     @pytask.task(id=document, kwargs=kwargs)
     def task_copy_to_root(depends_on, produces):
         """Copy a document to the root directory for easier retrieval."""

### 2024-08-15 documents/paper.tex
Fix latex compilation.
Add to paper.

@@
 
-\begin{document}
+\title{Topics Causal Inference}
 
-\title{Topics Causal Inference}
-\thanks{Bonn University. Email: s93jbudd[at]uni-bonn.de}
-
-\author{Julian Budde}
+\author{Julian Budde\thanks{Bonn University. Email: s93jbudd[at]uni-bonn.de}}
 
@@
 }
+
+\begin{document}
 
@@
 \paragraph*{Approximations to the CATE}
-Another approach is taken in~\cite{semenova2021debiased}:
-They propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
+~\cite{semenova2021debiased} propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
 Then, the new target for inference becomes the \textit{partial} CATE $s_0(X)$.
@@
 
+\subsection{Simulation Setup}
+\paragraph*{Designing a High-Dimensional DGP}
+A key question is how to design a DGP that is both ``realistic'', in the sense that the results are relevant to empirical applications, and still poses a high-dimensional problem.
+The typical approach taken in the literature, e.g.\ in~\cite{wager2018estimation} is to postulate a DGP that has convenient analytical features and is easy to simulate.
+This simplifies the analysis but begs the question of relevance.
+Also, it leaves some degree of freedom to the researcher to choose a DGP that favourably displays the advantages of their proposed method.
+To address this issue, ``empirical'' methods for simulation design have been proposed, such as the WGAN approach in~\cite{athey2024wgan}.
+Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP.
+The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in classifying observations as real or generated.
+
+While this approach can be usefully employed to design specific simulations given a dataset at hand for some questions, it also has some limitations, specifically with respect to the problem studied here.
+In particular, we are concerned with a high-dimensional, non-sparse structure of the CATE functional --- this is what the generic ML method is robust against.
+However, for a given DGP with high-dimensional structure it cannot be possible train a generator that preserves this structure.
+Else, if we had an estimate of the DGP, we could simply construct a plug-in estimator of the CATE, but as argued before such an estimator cannot exist.
+
+For this reason I resort to the analytical DGP presented in~\cite{wager2018estimation} (WA in the following).
+To illustrate the difficulty with using a WGAN generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA dgp.
+To the best of my knowledge, such simulation results are not available in the original or any other paper.
+\footnote{They do, however, assess robustness of their simulation setup in several dimensions, including sampling variability, in Section 5.}
+The following paragraphs explain the simulation setup in more detail.
+
+\paragraph*{WA DGP}
+I use the third DGP discussed in~\cite{wager2018estimation} as my baseline DGP\@.
+Covariates are generated as $X~U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
+~\cite{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
+Instead, the CATE function is generated by
+\begin{equation*}
+    \tau(x) = \xi(X_1)\xi(X_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
+\end{equation*}
+That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
+The propensity score is kept constant at 0.5.
+
+\paragraph*{WGAN Approximation}
+[Describe how WGAN was performed.]
+
+\subsection{Results: WA Replication and WGAN}
+
+\subsection{Results: Generic Machine Learning}
+
 \clearpage
@@
 % \setstretch{1}
-\printbibliography{}
+\printbibliography
 \setstretch{1.5}

### 2024-08-15 src/topics_causal_inf/config.py
Rename py_wa_replication to wa_replication.

@@ -15,8 +15,6 @@ WGAN_GEN = SRC / "wgan_generated"
 
 DOCUMENTS = ROOT.joinpath("documents").resolve()
 
-TEMPLATE_GROUPS = ["marital_status", "highest_qualification"]
-
 RNG = np.random.default_rng()
 
 DIMS_TO_RUN = [10, 15, 30]

### 2024-08-15 src/topics_causal_inf/wa_replication/task_wa_replication.py

@@ -20,7 +20,7 @@ from topics_causal_inf.config import (
     SUBSAMPLE_SHARE,
     WGAN_GEN,
 )
-from topics_causal_inf.py_wa_replication.py_wa_replication import simulation
+from topics_causal_inf.wa_replication.wa_replication import simulation
 
 
 class _Arguments(NamedTuple):
@@ -41,9 +41,9 @@ ID_TO_KWARGS = {
     f"{dgp.name}_{dim}_{dgen}": _Arguments(
         dgp=dgp,
         path_to_data=BLD
-        / "py_wa_replication"
+        / "wa_replication"
         / "sims"
-        / f"py_wa_replication_{dgp.name}_dim{dim}_{dgen}.pkl",
+        / f"wa_replication_{dgp.name}_dim{dim}_{dgen}.pkl",
         dim=dim,
         data_generator=dgen,
     )
@@ -56,7 +56,7 @@ ID_TO_KWARGS = {
 for id_, kwargs in ID_TO_KWARGS.items():
 
     @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
-    def task_py_wa_replication_sims(
+    def task_wa_replication_sims(
         dgp: DGP,
         path_to_data: Annotated[Path, Product],
         n_sim: int,

### 2024-08-15 src/topics_causal_inf/wa_replication/task_wa_replication_tables.py

@@ -8,7 +8,7 @@ from pytask import Product, task
 
 from topics_causal_inf.classes import DGP
 from topics_causal_inf.config import BLD
-from topics_causal_inf.py_wa_replication.task_py_wa_replication import (
+from topics_causal_inf.wa_replication.task_wa_replication import (
     DGPS_TO_RUN,
     ID_TO_KWARGS,
 )
@@ -27,9 +27,9 @@ ID_TO_KWARGS_TABLES = {
         dgp=dgp,
         path_to_results=PATH_TO_RESULTS,
         path_to_table=BLD
-        / "py_wa_replication"
+        / "wa_replication"
         / "tables"
-        / f"py_wa_replication_{dgp.name}.tex",
+        / f"wa_replication_{dgp.name}.tex",
     )
     for dgp in DGPS_TO_RUN
 }
@@ -38,7 +38,7 @@ ID_TO_KWARGS_TABLES = {
 for id_, kwargs in ID_TO_KWARGS_TABLES.items():
 
     @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
-    def task_py_wa_replication_table(
+    def task_wa_replication_table(
         dgp: DGP,
         path_to_results: dict[str, Path],
         path_to_table: Annotated[Path, Product],

### 2024-08-15 tests/wa_replication/test_wa_replication.py

@@ -4,7 +4,7 @@ from pathlib import Path
 import pandas as pd  # type: ignore[import-untyped]
 from topics_causal_inf.config import RNG, WGAN_GEN
 from topics_causal_inf.define_dgps import DGP3
-from topics_causal_inf.py_wa_replication.py_wa_replication import (
+from topics_causal_inf.wa_replication.wa_replication import (
     simulation,
 )
 

### 2024-08-15 documents/paper.tex
Add plots and tables to paper.

@@
 \DeclareMathOperator{\E}{E}
+\usepackage{multirow}
 
@@
 To address this issue, ``empirical'' methods for simulation design have been proposed, such as the WGAN approach in~\cite{athey2024wgan}.
-Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP.
+Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP.\@
 The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in classifying observations as real or generated.
@@
 \paragraph*{WGAN Approximation}
-[Describe how WGAN was performed.]
+Describe how WGAN was performed.
 
@@
 
+\begin{table}
+    \caption{Simulation Results: WA Replication and WGAN}
+    \center
+    \begin{tabular}{lccccc}
+        & Dimension & MSE & SD & Coverage & SD \\
+        \toprule
+        \multicolumn{6}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+        \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
+        \midrule
+
+        \multicolumn{6}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+        \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
+        \midrule
+
+        \multicolumn{6}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+        \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
+        \midrule
+
+        \multicolumn{6}{l}{\textit{Panel D:\ Placebo}} \\
+        \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\
+        \bottomrule
+    \end{tabular}
+
+\end{table}
+
 \subsection{Results: Generic Machine Learning}
+
+\begin{table}
+    \caption{Simulation Results: ML Generic --- Cate Estimation}
+    \center
+    \begin{tabular}{lcc}
+        Dimension & MSE & SD \\
+        \toprule
+        \multicolumn{3}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+        \input{../bld/generic_ml/tables/generic_ml_dgp3.tex} \\
+        \midrule
+
+        \multicolumn{3}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+        \input{../bld/generic_ml/tables/generic_ml_dgp4.tex} \\
+        \midrule
+
+        \multicolumn{3}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+        \input{../bld/generic_ml/tables/generic_ml_dgp5.tex} \\
+        \midrule
+
+        \multicolumn{3}{l}{\textit{Panel D:\ Placebo}} \\
+        \input{../bld/generic_ml/tables/generic_ml_dgp6.tex} \\
+        \bottomrule
+    \end{tabular}
+
+\end{table}
+
+% Gates Figure
+\begin{figure}
+    \caption{Simulation Results: Generic ML --- GATES}\label{fig:gates}
+
+    \centering
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp3.png}
+         \caption{Original WA DGP 3}\label{fig_gates:dgp3}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp4.png}
+         \caption{High-dimensional --- 4 Covariates}\label{fig_gates:dgp4}
+     \end{subfigure}
+
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp5.png}
+         \caption{High-dimensional --- 8 Covariates}\label{fig_gates:dgp5}
+     \end{subfigure}
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp6.png}
+         \caption{Placebo}\label{fig_gates:dgp6}
+     \end{subfigure}
+
+\end{figure}
+
+% Gates Figure
+\begin{figure}
+    \caption{Simulation Results: Generic ML --- CLANS}\label{fig:clans}
+
+    \centering
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp3.png}
+         \caption{Original WA DGP 3}\label{fig_clans:dgp3}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp4.png}
+         \caption{High-dimensional --- 4 Covariates}\label{fig_clans:dgp4}
+     \end{subfigure}
+
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp5.png}
+         \caption{High-dimensional --- 8 Covariates}\label{fig_clans:dgp5}
+     \end{subfigure}
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp6.png}
+         \caption{Placebo}\label{fig_clans:dgp6}
+     \end{subfigure}
+
+\end{figure}
 

### 2024-08-15 src/topics_causal_inf/generic_ml/task_generic_ml_plots.py

@@ -51,7 +51,7 @@ for _id, kwargs in ID_TO_KWARGS_GATES.items():
                 go.Histogram(
                     x=res[f"gate_{i}"],
                     histnorm="probability",
-                    name=f"gate_{i}",
+                    name=f"Group {i}",
                     marker_color=color,
                 ),
             )
@@ -61,8 +61,23 @@ for _id, kwargs in ID_TO_KWARGS_GATES.items():
                 line_color=color,
             )
 
-        fig.update_layout(barmode="overlay")
+        fig.update_layout(
+            barmode="overlay",
+            legend={
+                "yanchor": "top",
+                "y": -0.2,  # Position the legend below the plot area
+                "xanchor": "center",
+                "x": 0.5,
+                "orientation": "h",
+            },
+        )
         fig.update_traces(opacity=0.75)
+
+        fig.update_layout(
+            showlegend=False,
+        )
+
+        fig.update_layout(margin={"l": 10, "r": 10, "t": 10, "b": 10})
 
         fig.write_image(path_to_plot)
 
@@ -148,5 +163,11 @@ for _id, kwargs in ID_TO_KWARGS_CLANS.items():
             fig.update_xaxes(range=[0, 1], row=row, col=col)
 
         fig.update_traces(opacity=0.75)
+
+        fig.update_layout(
+            showlegend=False,
+        )
+
+        fig.update_layout(margin={"l": 10, "r": 10, "t": 10, "b": 10})
 
         fig.write_image(path_to_plot)

### 2024-08-15 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -7,6 +7,7 @@ import pandas as pd  # type: ignore[import-untyped]
 from pytask import Product, task
 
 from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN
+from topics_causal_inf.utilities import clean_tex_table
 
 
 class _Arguments(NamedTuple):
@@ -48,7 +49,13 @@ for id_, kwargs in ID_TO_KWARGS.items():
         )
 
         # Save table
-        table.to_latex(
-            path_to_table,
+        out = table.to_latex(
             float_format="%.3f",
+            header=False,
         )
+
+        # Remove all rows starting with "\begin{tabular}" and "\end{tabular}"
+        out = clean_tex_table(out)
+
+        # Write to file
+        path_to_table.write_text(out)

### 2024-08-15 src/topics_causal_inf/utilities.py

@@ -1,5 +1,6 @@
 """Utilities used throughout the project."""
 
+import re
 from copy import copy
 
 import numpy as np
@@ -156,3 +157,18 @@ def _check_dgp_and_dim_compatible(dgp: DGP, dim: int) -> None:
         length = len(dgp.treatment_effect.keywords["x_range"])
         msg = f"dgp.treatment_effect x_range has length {length} > dim {dim}."
         raise ValueError(msg)
+
+
+def clean_tex_table(table: str) -> str:
+    """Clean table produced by df.to_latex for paper."""
+    out = re.sub(r"\\begin{tabular}.*\n", "", table)
+    out = re.sub(r"\\end{tabular}.*\n", "", out)
+    out = re.sub(r"\\toprule.*\n", "", out)
+    out = re.sub(r"\\midrule.*\n", "", out)
+    out = re.sub(r"\\bottomrule.*\n", "", out)
+    out = re.sub("_", " ", out)
+    out = re.sub("standard", "Standard DGP", out)
+    out = re.sub("wgan", "WGAN DGP", out)
+    out = re.sub(r"data generator.*\n", "", out)
+    out = re.sub(r"\\cline.*\n", "", out)
+    return re.sub(r"dim .*\n", "", out)

### 2024-08-15 src/topics_causal_inf/wa_replication/task_wa_replication_tables.py

@@ -8,6 +8,7 @@ from pytask import Product, task
 
 from topics_causal_inf.classes import DGP
 from topics_causal_inf.config import BLD
+from topics_causal_inf.utilities import clean_tex_table
 from topics_causal_inf.wa_replication.task_wa_replication import (
     DGPS_TO_RUN,
     ID_TO_KWARGS,
@@ -56,8 +57,7 @@ for id_, kwargs in ID_TO_KWARGS_TABLES.items():
             ["mean", "std"],
         )
 
-        res_to_table.to_latex(
-            path_to_table,
+        out = res_to_table.to_latex(
             formatters={
                 ("mse", "mean"): "{:,.2f}".format,
                 ("mse", "std"): "{:,.2f}".format,
@@ -66,3 +66,8 @@ for id_, kwargs in ID_TO_KWARGS_TABLES.items():
             },
             header=False,
         )
+
+        out = clean_tex_table(out)
+
+        # Write to file
+        path_to_table.write_text(out)

### 2024-08-15 tests/wa_replication/test_wa_replication.py
Remove wgan from tests.

@@ -11,14 +11,19 @@ from topics_causal_inf.wa_replication.wa_replication import (
 
 def test_simulation_runs() -> None:
     for dgp in [DGP3]:
-        for dgen in ["wgan", "standard"]:
-            wgan_train_data = pd.read_feather(WGAN_GEN / f"df_{dgp.name}.feather")
-            generators = pickle.load(
-                Path.open(WGAN_GEN / f"generators_{dgp.name}.pkl", "rb"),
-            )
-            data_wrappers = pickle.load(
-                Path.open(WGAN_GEN / f"data_wrapper_{dgp.name}.pkl", "rb"),
-            )
+        for dgen in ["standard"]:
+            if dgen == "wgan":
+                wgan_train_data = pd.read_feather(WGAN_GEN / f"df_{dgp.name}.feather")
+                generators = pickle.load(
+                    Path.open(WGAN_GEN / f"generators_{dgp.name}.pkl", "rb"),
+                )
+                data_wrappers = pickle.load(
+                    Path.open(WGAN_GEN / f"data_wrapper_{dgp.name}.pkl", "rb"),
+                )
+            else:
+                wgan_train_data = None
+                generators = None
+                data_wrappers = None
 
             simulation(
                 n_sim=2,

### 2024-08-15 documents/paper.tex
Add coverage for GATEs and BLP params.

@@
 \paragraph*{ML Method Selection}
-[Tbd.]
+Because the approach works with \textit{any} first stage machine learning method, a researcher might consider constructing proxies from different methods.
+To choose a method, the authors propose a criterion that maximizes the (absolute) correlation between the true CATE $s_0(Z)$ and the ML proxy $S(Z)$.
+Remember we can identify $\beta_2$ by the strategy discussed above. Because $\beta_2$ is the BLP of $s_0(Z)$ given $S(Z)$ we have
+\begin{equation*}
+    \beta_2 = \frac{Cov[s_0(Z), S(Z)]}{Var(S(Z))}.
+\end{equation*}
+By squaring and rescaling we are at the following criterion
+\begin{equation*}
+    \Lambda := |\beta_2|^2 Var(S(Z)) = Corr(s_0(Z), S(Z))^2 Var(s_0(Z)).
+\end{equation*}
+Hence, maximizing $\Lambda$ by choosing a proxy $S(Z)$, maximizes the desired correlation.
+Note the left-hand side can be estimated from the data, hence the criterion is feasible.
 
 \paragraph*{Split-based Inference}
-
+Finally, the paper proposes median-aggregating the results over multiple data splits to decrease dependence on the particular split.
+For example, for a given hypothesis of interest (for example values for BLP parameters $\beta_1, \beta_2$) we can consider properties of the \textit{median-aggregated p-value} defined by
+\begin{equation*}
+    p^+ = Med(p_A^+|Data)
+\end{equation*}
+where $Med(\cdot)$ denotes the median over datasplits and $p_A^+$ is a (one-sided) p-value in a given auxiliary sample $A$.
+Under regularity condition spelled out in the paper, it is possible to show that under $H_0$ p-values are uniform, that is inference is valid.
+\footnote{Depending on the exact regularity condition p-values might have to be multiplied by a factor of 2.
+The authors argue, however, that in their experience from simulations smaller p-values are typically appropriate.}
 \section{Simulation and Results}
@@
 
+% Generic ML: BLP and GATES Coverage
+\begin{table}
+    \caption{Simulation Results: ML Generic --- BLP and GATES Coverage}
+    \center
+    \begin{tabular}{lccccccc}
+        Dimension & \multicolumn{7}{c}{Coverage} \\
+        \toprule
+        \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+        \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp3.tex} \\
+        \midrule
+
+        \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+        \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp4.tex} \\
+        \midrule
+
+        \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+        \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp5.tex} \\
+        \midrule
+
+        \multicolumn{8}{l}{\textit{Panel D:\ Placebo}} \\
+        \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp6.tex} \\
+        \bottomrule
+    \end{tabular}
+
+\end{table}
+
 % Gates Figure

### 2024-08-15 src/topics_causal_inf/config.py

@@ -24,5 +24,5 @@ DGPS_TO_RUN = [DGP3, DGP4, DGP5, DGP6]
 NUM_TREES = 100
 SUBSAMPLE_SHARE = 0.2
 N_OBS = 10_000
-N_SIM = 20
-N_SPLITS = 10
+N_SIM = 40
+N_SPLITS = 20

### 2024-08-15 src/topics_causal_inf/generic_ml/simulation.py

@@ -25,9 +25,21 @@ def simulation(
     mse_blp = np.zeros(n_sims)
     gates = np.zeros((n_sims, 5))
     clans = np.zeros((n_sims, 2, dim))
+    blp_ci_lo = np.zeros((n_sims, 2))
+    blp_ci_hi = np.zeros((n_sims, 2))
+    gates_ci_lo = np.zeros((n_sims, 5))
+    gates_ci_hi = np.zeros((n_sims, 5))
 
     for i in range(n_sims):
-        mse_blp[i], gates[i, :], clans[i, :, :] = _single_experiment(
+        (
+            mse_blp[i],
+            gates[i, :],
+            clans[i, :, :],
+            blp_ci_lo[i],
+            blp_ci_hi[i],
+            gates_ci_lo[i, :],
+            gates_ci_hi[i, :],
+        ) = _single_experiment(
             n_obs=n_obs,
             dim=dim,
             n_splits=n_splits,
@@ -43,10 +55,26 @@ def simulation(
     out["n_sims"] = n_sims
     out["dim"] = dim
 
+    beta_1, beta_2 = _simulate_true_blp(
+        dgp=dgp,
+        dim=dim,
+        ml_learner=ml_learner,
+        rng=rng,
+    )
+
+    out["true_blp_beta_1"] = beta_1
+    out["true_blp_beta_2"] = beta_2
+    out["blp_ci_lo_beta_1"] = blp_ci_lo[:, 0]
+    out["blp_ci_hi_beta_1"] = blp_ci_hi[:, 0]
+    out["blp_ci_lo_beta_2"] = blp_ci_lo[:, 1]
+    out["blp_ci_hi_beta_2"] = blp_ci_hi[:, 1]
+
     true_gates = _simulate_true_gates(dgp=dgp, dim=dim, rng=rng)
 
     for i in range(5):
         out[f"gate_{i}"] = gates[:, i]
+        out[f"gate_ci_lo_{i}"] = gates_ci_lo[:, i]
+        out[f"gate_ci_hi_{i}"] = gates_ci_hi[:, i]
         out[f"true_gate_{i}"] = true_gates[i]
 
     true_clans = _simulate_true_clan(dgp=dgp, dim=dim, rng=rng)
@@ -94,6 +122,10 @@ def _single_experiment(
         np.mean((data_eval["pred"] - data_eval["true"]) ** 2),
         res.gates_params,
         res.clan,
+        res.blp_ci_hi,
+        res.blp_ci_lo,
[…]

### 2024-08-15 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -17,7 +17,7 @@ class _Arguments(NamedTuple):
 
 # For each dgp collect all results for all dim values
 ID_TO_KWARGS = {
-    dgp.name: _Arguments(
+    f"{dgp.name}_mse": _Arguments(
         path_to_table=BLD / "generic_ml" / "tables" / f"generic_ml_{dgp.name}.tex",
         path_to_res=[
             BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl"
@@ -44,6 +44,70 @@ for id_, kwargs in ID_TO_KWARGS.items():
                 {
                     "mean": x["mse_blp"].mean(),
                     "std": x["mse_blp"].std(),
+                },
+            ),
+        )
+
+        # Save table
+        out = table.to_latex(
+            float_format="%.3f",
+            header=False,
+        )
+
+        # Remove all rows starting with "\begin{tabular}" and "\end{tabular}"
+        out = clean_tex_table(out)
+
+        # Write to file
+        path_to_table.write_text(out)
+
+
+ID_TO_KWARGS_COVERAGE = {
+    f"{dgp.name}_blp_gates_coverage": _Arguments(
+        path_to_table=BLD
+        / "generic_ml"
+        / "tables"
+        / f"generic_ml_blp_gates_coverage_{dgp.name}.tex",
+        path_to_res=[
+            BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl"
+            for dim in DIMS_TO_RUN
+        ],
+    )
+    for dgp in DGPS_TO_RUN
+}
+
+
+for id_, kwargs in ID_TO_KWARGS_COVERAGE.items():
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_generic_ml_tables_coverage(
+        path_to_res: list[Path],
+        path_to_table: Annotated[Path, Product],
+    ) -> None:
+        """Task for generic_ml result tables."""
+        # Load results
+        res = pd.concat([pd.read_pickle(path) for path in path_to_res])
+
+        for i in range(5):
+            res[f"covers_gate_{i}"] = (
+                res[f"gate_ci_lo_{i}"] < res[f"true_gate_{i}"]
+            ) & (res[f"gate_ci_hi_{i}"] > res[f"true_gate_{i}"])
+
+        for beta in ["beta_1", "beta_2"]:
+            res[f"covers_blp_{beta}"] = (
[…]

### 2024-08-16 documents/paper.tex
Add estimated CATE for wgan MSE.

@@
 \begin{equation*}
-    \tau(x) = \xi(X_1)\xi(X_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
+    \tau_1(x) = \xi(x_1)\xi(x_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
 \end{equation*}
 That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
-The propensity score is kept constant at 0.5.
+The propensity score is kept constant at 0.5. I present results for various ambient dimensions $d$.
+
+I consider three additional DGPs: The first two DGPs increase the dimension of the CATE function, by increasing the number of arguments the function depends on.
+In particular, I use
+\begin{equation*}
+    \tau_2(x) = \prod_{i=1}^{4}\xi(x_i),
+\end{equation*}
+as well as
+\begin{equation*}
+    \tau_3(x) = \prod_{i=1}^{8}\xi(x_i).
+\end{equation*}
+I keep the rest of the DGP the same. Finally I also consider a \textit{placebo} setting with
+\begin{equation*}
+    \tau_4(x) = 0.
+\end{equation*}
+The first two extensions are used to investigate the performance of the causal forest and generic machine learning approach when heterogeneity is high-dimensional.
+The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
+A similar question is investigated in Figure 1 of~\cite{chernozhukov2023genml}.
 
 \paragraph*{WGAN Approximation}
-Describe how WGAN was performed.
+For generating the WGAN approximation to the DGP I follow the baseline approach described in ds-wgan package example.
+In particular, I train the generator and critic using the same hyperparameters.
+All fitting is performed twice, once for the distribution of $Y$ conditional on $X, D$ and once for the distribution of $X$ conditional on $D$.
+I train the model based on a sample of $N=100,000$ observations with dimension $d=10$ drawn from the baseline WA dgp as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
+I deliberately chose a large sample (at least larger than the $N=10,000$ observations used throughout simulations) to attenuate noise resulting from this original sample draw.
+To further address this it would be ideal to increase $N$ further or aggregate the results over different initial sample draws.
+Given the training time of these models I found both approaches to be infeasible for the scope of this paper.
+
+One additional difficulty in using WGANs for CATE estimation is to define the target parameter/the "truth" against to which we want to evaluate the estimates.
+In the simpler case of ATE estimation, one approach is to draw a large sample from the generator, which we term the population, and to compute the ATE in this population by
+\begin{equation*}
+    ATE := \frac{1}{N^{pop}} Y_i(1) - Y_i(0).
+\end{equation*}
+However, a similar approach is not available for the CATE.
+While we could compute a similar average conditional on $X$, if there are continuous covariates that would be a bad approximation to the true underlying CATE since we end up with individual treatment effects.
+If, on the other hand, we fit a model to the population we know from that even if the WGAN preserved the high-dimensional structure it would not be possible to consistently estimate it.
+For the analysis I compute MSE using two approaches: Once using the true CATE function of the WA DGP and once using an estimated CATE based on the generated population using a causal forest approach.
+The second approach in particular should result in understated MSE estimates: If the WGAN generator cannot approximate the high-dimensional structure, using an estimated CATE based on a generated population will understate the difficulty of the problem.
+The first approach is of course not available in practice since the true CATE is unknown.
 
@@
     \center
-    \begin{tabular}{lccccc}
-        & Dimension & MSE & SD & Coverage & SD \\
+    \begin{tabular}{lccccccc}
+        & Dimension & MSE (True) & SD & MSE (Pop) & SD & Coverage & SD \\
         \toprule
-        \multicolumn{6}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+        \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
         \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
@@
 
-        \multicolumn{6}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+        \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
         \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
@@
 
-        \multicolumn{6}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+        \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
         \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
@@
 
-        \multicolumn{6}{l}{\textit{Panel D:\ Placebo}} \\
+        \multicolumn{8}{l}{\textit{Panel D:\ Placebo}} \\
         \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\

### 2024-08-16 documents/task_documents.py

@@ -5,8 +5,39 @@ import shutil
 import pytask
 from pytask_latex import compilation_steps as cs
 from topics_causal_inf.config import BLD, DOCUMENTS, ROOT
+from topics_causal_inf.generic_ml.task_generic_ml_plots import (
+    ID_TO_KWARGS_CLANS,
+    ID_TO_KWARGS_GATES,
+)
+from topics_causal_inf.generic_ml.task_generic_ml_tables import (
+    ID_TO_KWARGS,
+    ID_TO_KWARGS_COVERAGE,
+)
+from topics_causal_inf.wa_replication.task_wa_replication_tables import (
+    ID_TO_KWARGS_TABLES,
+)
 
 documents = ["paper"]
+
+path_genml_tables = [args.path_to_table for _, args in ID_TO_KWARGS.items()]
+path_genml_tables_coverage = [
+    args.path_to_table for _, args in ID_TO_KWARGS_COVERAGE.items()
+]
+path_genml_plots_clans = [
+    args["path_to_plot"] for _, args in ID_TO_KWARGS_CLANS.items()
+]
+path_genml_plots_gates = [
+    args["path_to_plot"] for _, args in ID_TO_KWARGS_GATES.items()
+]
+path_wa_tables = [args.path_to_table for _, args in ID_TO_KWARGS_TABLES.items()]
+
+dependencies = (
+    path_genml_tables
+    + path_genml_tables_coverage
+    + path_genml_plots_clans
+    + path_genml_plots_gates
+    + path_wa_tables
+)
 
 for document in documents:
 
@@ -23,7 +54,7 @@ for document in documents:
             ),
         ),
     )
-    @pytask.task(id=document)
+    @pytask.task(id=document, kwargs={"depends_on": dependencies})
     def task_compile_document():
         """Compile the document specified in the latex decorator."""
 

### 2024-08-16 src/topics_causal_inf/config.py

@@ -25,4 +25,4 @@ NUM_TREES = 100
 SUBSAMPLE_SHARE = 0.2
 N_OBS = 10_000
 N_SIM = 40
-N_SPLITS = 20
+N_SPLITS = 10

### 2024-08-16 src/topics_causal_inf/wa_replication/task_wa_replication_tables.py

@@ -52,6 +52,8 @@ for id_, kwargs in ID_TO_KWARGS_TABLES.items():
 
         res = res[res["dgp"] == dgp.name]
         res = res.drop(columns=["dgp"])
+        if "mse_cf" in res.columns:
+            res = res.drop(columns=["mse_cf"])
 
         res_to_table = res.groupby(["data_generator", "dim"]).agg(
             ["mean", "std"],
@@ -61,6 +63,8 @@ for id_, kwargs in ID_TO_KWARGS_TABLES.items():
             formatters={
                 ("mse", "mean"): "{:,.2f}".format,
                 ("mse", "std"): "{:,.2f}".format,
+                ("mse_pop_cate", "mean"): "{:,.2f}".format,
+                ("mse_pop_cate", "std"): "{:,.2f}".format,
                 ("coverage", "mean"): "{:,.2f}".format,
                 ("coverage", "std"): "{:,.2f}".format,
             },

### 2024-08-16 src/topics_causal_inf/wa_replication/wa_replication.py

@@ -45,6 +45,7 @@ def simulation(
     if data_generator == "wgan":
         _check_wgan_args(wgan_train_data, generators, data_wrappers)
     mse = np.zeros(n_sim)
+    mse_pop_cate = np.zeros(n_sim)
     coverage = np.zeros(n_sim)
 
     for i in range(n_sim):
@@ -54,10 +55,12 @@ def simulation(
                 data_generators=generators,  # type: ignore[arg-type]
                 data_wrappers=data_wrappers,  # type: ignore[arg-type]
             )
+            pop_cate = _estimate_cf_cate(pop_data, dim)
         else:
             pop_data = None
+            pop_cate = None
 
-        mse[i], coverage[i] = _experiment(
+        mse[i], mse_pop_cate[i], coverage[i] = _experiment(
             n_obs=n_obs,
             dim=dim,
             num_trees=num_trees,
@@ -66,12 +69,14 @@ def simulation(
             data_generator=data_generator,
             alpha=alpha,
             pop_data=pop_data,
+            pop_cate=pop_cate,
             rng=rng,
         )
 
     return pd.DataFrame(
         {
             "mse": mse,
+            "mse_pop_cate": mse_pop_cate,
             "coverage": coverage,
             "dim": dim,
             "data_generator": data_generator,
@@ -90,7 +95,8 @@ def _experiment(
     rng: np.random.Generator,
     alpha: float = 0.05,
     pop_data: pd.DataFrame | None = None,
-) -> tuple[float, float]:
+    pop_cate: CausalForest | None = None,
+) -> tuple[float, float, float]:
     feature_names = [f"x{i}" for i in range(dim)]
 
     if data_generator == "standard":
@@ -129,11 +135,17 @@ def _experiment(
     # Calculate (expected) MSE
     mse = np.mean((pred - true) ** 2)
 
+    if data_generator == "wgan" and pop_cate is not None:
+        true_pop_cate = pop_cate.predict(data_test[feature_names]).flatten()
+        mse_pop_cate = np.mean((pred - true_pop_cate) ** 2)
+    else:
+        mse_pop_cate = np.nan
+
     # Calculate (expected) coverage
     ci_lo, ci_hi = cf.predict_interval(X=data_test[feature_names], alpha=alpha)
     coverage = np.mean((ci_lo.flatten() <= true) & (true <= ci_hi.flatten()))
 
[…]

### 2024-08-16 documents/paper.tex
Add estimated CATE for wgan MSE.
Draw single pop data for wgan; update tables and paper.

@@
 The authors argue, however, that in their experience from simulations smaller p-values are typically appropriate.}
-\section{Simulation and Results}
+\section{Simulation Setup}
 
@@
 \begin{equation*}
-    ATE := \frac{1}{N^{pop}} Y_i(1) - Y_i(0).
+    ATE := \frac{1}{N^{pop}} \sum_{i=1}^{N^{pop}} Y_i(1) - Y_i(0).
 \end{equation*}
-However, a similar approach is not available for the CATE.
+However, a similar approach is not available for the CATE\@.
 While we could compute a similar average conditional on $X$, if there are continuous covariates that would be a bad approximation to the true underlying CATE since we end up with individual treatment effects.
@@
 The first approach is of course not available in practice since the true CATE is unknown.
+
+\section{Simulation Results}
+Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to \cite{wager2018estimation}.
+This low number is chosen to make the problem computationally feasible.
+I use $100$ trees in all causal forest estimators and a sub-sampling share of $0.2$.
+These parameters are much smaller than in the original~\cite{wager2018estimation} to reduce the run-time of the simulation.
+I still find very similar results in the replication parts, as outlined below.
+I follow~\cite{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
+To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation of simulation runs.
+For generic machine learning simulations I use a standard random forest as the baseline learner for $b_0(Z) = \E[Y(0)|Z]$ and a causal forest with similar hyperparameters as above for estimating $s_0(Z)$.
 
@@
 
+\section{Conclusion}
+
 \clearpage

### 2024-08-16 src/topics_causal_inf/utilities.py

@@ -167,6 +167,7 @@ def clean_tex_table(table: str) -> str:
     out = re.sub(r"\\midrule.*\n", "", out)
     out = re.sub(r"\\bottomrule.*\n", "", out)
     out = re.sub("_", " ", out)
+    out = re.sub("NaN", "-", out)
     out = re.sub("standard", "Standard DGP", out)
     out = re.sub("wgan", "WGAN DGP", out)
     out = re.sub(r"data generator.*\n", "", out)

### 2024-08-16 src/topics_causal_inf/wa_replication/wa_replication.py

@@ -48,18 +48,18 @@ def simulation(
     mse_pop_cate = np.zeros(n_sim)
     coverage = np.zeros(n_sim)
 
-    for i in range(n_sim):
-        if data_generator == "wgan":
-            pop_data = _generate_pop_data(
-                wgan_train_data=wgan_train_data,
-                data_generators=generators,  # type: ignore[arg-type]
-                data_wrappers=data_wrappers,  # type: ignore[arg-type]
-            )
-            pop_cate = _estimate_cf_cate(pop_data, dim)
-        else:
-            pop_data = None
-            pop_cate = None
+    if data_generator == "wgan":
+        pop_data = _generate_pop_data(
+            wgan_train_data=wgan_train_data,
+            data_generators=generators,  # type: ignore[arg-type]
+            data_wrappers=data_wrappers,  # type: ignore[arg-type]
+        )
+        pop_cate = _estimate_cf_cate(pop_data, dim)
+    else:
+        pop_data = None
+        pop_cate = None
 
+    for i in range(n_sim):
         mse[i], mse_pop_cate[i], coverage[i] = _experiment(
             n_obs=n_obs,
             dim=dim,

### 2024-08-16 documents/paper.tex
Add table and figure notes.

@@
 \usepackage[utf8]{inputenc}
-\usepackage{float, afterpage, rotating, graphicx}
+% \usepackage{float} % Note: incompatible with floatrow
+\usepackage{afterpage, rotating, graphicx}
 \usepackage{epstopdf}
@@
 \usepackage{multirow}
+\usepackage{threeparttable}
+\usepackage[capposition=top]{floatrow}
 
@@
 For generic machine learning simulations I use a standard random forest as the baseline learner for $b_0(Z) = \E[Y(0)|Z]$ and a causal forest with similar hyperparameters as above for estimating $s_0(Z)$.
+When analysing group average treatment effects (GATES), I define five groups based on the treatment effect quintiles predicted by the ML proxy.
+To approximate the true GATES induced by the proxy I estimate this on a large sample of size $N=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
+I use a main, auxiliary splitshare of $0.5$ throughout the analysis.
 
@@
     \caption{Simulation Results: WA Replication and WGAN}
+    \begin{threeparttable}
         \center
@@
         \end{tabular}
-
+        \begin{tablenotes}
+            \small
+            \item \textbf{Notes}: This table reports simulation results for~\cite{wager2018estimation}.
+            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+            \textit{Dimension} refers to the number of covariates in the DGP\@.
+\textit{Standard} DGP refers to sampling directly from the DGP, while \textit{WGAN} DGP refers to sampling from a WGAN trained based on a sample from the given DGP.
+            MSE and coverage are calculated based on a test sample of $10,000$ observations.
+MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP.
+            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
+            For details see the text.
+        \end{tablenotes}
+    \end{threeparttable}
 \end{table}
@@
     \caption{Simulation Results: ML Generic --- Cate Estimation}
+    \begin{threeparttable}
     \center
@@
     \end{tabular}
+    \begin{tablenotes}
+        \small
+        \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml}.
+        In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
+        Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+        \textit{Dimension} refers to the number of covariates in the DGP\@.
+        MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
+        Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
+        For details see the text.
+    \end{tablenotes}
+\end{threeparttable}
 
@@
     \caption{Simulation Results: ML Generic --- BLP and GATES Coverage}
+    \begin{threeparttable}
         \center
@@
             Dimension & \multicolumn{7}{c}{Coverage} \\
+            & \multicolumn{2}{c}{Best Linear Predictor} & \multicolumn{5}{c}{Group Average Treatment Effects} \\
+            & $\beta_1$ & $\beta_2$ & $\gamma_1$ & $\gamma_2$ & $\gamma_3$ & $\gamma_4$ & $\gamma_5$ \\
             \toprule
@@
 
+        \begin{tablenotes}
+            \small
+            \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml} studying coverage for the BLP and GATES parameters (five groups).
+            In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
+            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+            \textit{Dimension} refers to the number of covariates in the DGP\@.
+            MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
+            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
+            For details see the text.
+        \end{tablenotes}
+    \end{threeparttable}
 \end{table}
@@
      \end{subfigure}
+
+     \floatfoot{\textbf{Notes}: This figure reports simulation results for GATES parameters of the generic machine learning approach.
+     For each of the four DGPs, it plots the distribution of each of the five GATE parameter estimates over the simulation runs.
+     The dashed line indicates the true GATES using a large population draw. Sample size $N=10,000$ and $40$ simulation runs.
+     }
 
@@
 
+     \floatfoot{\textbf{Notes}: This figure reports simulation results for CLAN of the generic machine learning approach.
+     For each of the four DGPs, it plots the distribution of the estimated characteristics for the first four covariates over the simulation runs.
+     The dashed line indicates the true characteristics of the least and most affected group estimated from a large sample.
+     Sample size $N=10,000$ and $40$ simulation runs.
+     }
+
 \end{figure}

### 2024-08-16 documents/paper.tex
Correct minor latex problems.

@@
 To address this issue, ``empirical'' methods for simulation design have been proposed, such as the WGAN approach in~\cite{athey2024wgan}.
-Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP.\@
+Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP\@.
 The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in classifying observations as real or generated.
@@
 For this reason I resort to the analytical DGP presented in~\cite{wager2018estimation} (WA in the following).
-To illustrate the difficulty with using a WGAN generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA dgp.
+To illustrate the difficulty with using a WGAN generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA DGP\@.
 To the best of my knowledge, such simulation results are not available in the original or any other paper.
@@
 All fitting is performed twice, once for the distribution of $Y$ conditional on $X, D$ and once for the distribution of $X$ conditional on $D$.
-I train the model based on a sample of $N=100,000$ observations with dimension $d=10$ drawn from the baseline WA dgp as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
+I train the model based on a sample of $N=100,000$ observations with dimension $d=10$ drawn from the baseline WA DGP as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
 I deliberately chose a large sample (at least larger than the $N=10,000$ observations used throughout simulations) to attenuate noise resulting from this original sample draw.
@@
 
-One additional difficulty in using WGANs for CATE estimation is to define the target parameter/the "truth" against to which we want to evaluate the estimates.
+One additional difficulty in using WGANs for CATE estimation is to define the target parameter/the ``truth'' against to which we want to evaluate the estimates.
 In the simpler case of ATE estimation, one approach is to draw a large sample from the generator, which we term the population, and to compute the ATE in this population by
@@
 \section{Simulation Results}
-Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to \cite{wager2018estimation}.
+Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to~\cite{wager2018estimation}.
 This low number is chosen to make the problem computationally feasible.
@@
             \textit{Dimension} refers to the number of covariates in the DGP\@.
-\textit{Standard} DGP refers to sampling directly from the DGP, while \textit{WGAN} DGP refers to sampling from a WGAN trained based on a sample from the given DGP.
+\textit{Standard} DGP refers to sampling directly from the DGP, while \textit{WGAN} DGP refers to sampling from a WGAN trained based on a sample from the given DGP\@.
             MSE and coverage are calculated based on a test sample of $10,000$ observations.
-MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP.
+MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP\@.
             Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).

### 2024-08-16 src/topics_causal_inf/utilities.py

@@ -167,7 +167,7 @@ def clean_tex_table(table: str) -> str:
     out = re.sub(r"\\midrule.*\n", "", out)
     out = re.sub(r"\\bottomrule.*\n", "", out)
     out = re.sub("_", " ", out)
-    out = re.sub("NaN", "-", out)
+    out = re.sub("NaN", "---", out)
     out = re.sub("standard", "Standard DGP", out)
     out = re.sub("wgan", "WGAN DGP", out)
     out = re.sub(r"data generator.*\n", "", out)

### 2024-08-16 documents/task_documents.py
Wager and Athey 2018 JASA Replication and Extension (#1)
Add baseline analysis for submission including WA 2018 replication (incl wgan DGP) and generic ML simulation for BLP, Gates, and CLAN.

@@ -5,20 +5,56 @@ import shutil
 import pytask
 from pytask_latex import compilation_steps as cs
 from topics_causal_inf.config import BLD, DOCUMENTS, ROOT
+from topics_causal_inf.generic_ml.task_generic_ml_plots import (
+    ID_TO_KWARGS_CLANS,
+    ID_TO_KWARGS_GATES,
+)
+from topics_causal_inf.generic_ml.task_generic_ml_tables import (
+    ID_TO_KWARGS,
+    ID_TO_KWARGS_COVERAGE,
+)
+from topics_causal_inf.wa_replication.task_wa_replication_tables import (
+    ID_TO_KWARGS_TABLES,
+)
 
-documents = ["paper", "presentation"]
+documents = ["paper"]
+
+path_genml_tables = [args.path_to_table for _, args in ID_TO_KWARGS.items()]
+path_genml_tables_coverage = [
+    args.path_to_table for _, args in ID_TO_KWARGS_COVERAGE.items()
+]
+path_genml_plots_clans = [
+    args["path_to_plot"] for _, args in ID_TO_KWARGS_CLANS.items()
+]
+path_genml_plots_gates = [
+    args["path_to_plot"] for _, args in ID_TO_KWARGS_GATES.items()
+]
+path_wa_tables = [args.path_to_table for _, args in ID_TO_KWARGS_TABLES.items()]
+
+dependencies = (
+    path_genml_tables
+    + path_genml_tables_coverage
+    + path_genml_plots_clans
+    + path_genml_plots_gates
+    + path_wa_tables
+)
 
 for document in documents:
 
-    @pytask.mark.skip()
     @pytask.mark.latex(
         script=DOCUMENTS / f"{document}.tex",
         document=BLD / "documents" / f"{document}.pdf",
         compilation_steps=cs.latexmk(
-            options=("--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd"),
+            options=(
+                "--pdf",
+                "--interaction=nonstopmode",
+                "--synctex=1",
+                "--cd",
+                "--f",
+            ),
         ),
     )
-    @pytask.task(id=document)
+    @pytask.task(id=document, kwargs={"depends_on": dependencies})
     def task_compile_document():
         """Compile the document specified in the latex decorator."""
 
@@ -27,7 +63,6 @@ for document in documents:
         "produces": ROOT / f"{document}.pdf",
     }
 
-    @pytask.mark.skip()
[…]

### 2024-08-16 src/topics_causal_inf/classes.py

@@ -1,0 +1,38 @@
+import functools
+from collections.abc import Callable
+from dataclasses import dataclass
+from typing import NamedTuple
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
+
+
+@dataclass
+class GenericMLResult:
+    """Result for generic_ml procedure."""
+
+    blp_params: np.ndarray
+    blp_se: np.ndarray
+    blp_tvals: np.ndarray
+    blp_pvals: np.ndarray
+    blp_ci_lo: np.ndarray
+    blp_ci_hi: np.ndarray
+    ml_fitted_d0: RegressorMixin
+    ml_fitted_d1: RegressorMixin
+    gates_params: np.ndarray
+    gates_se: np.ndarray
+    gates_tvals: np.ndarray
+    gates_pvals: np.ndarray
+    gates_ci_lo: np.ndarray
+    gates_ci_hi: np.ndarray
+    clan: np.ndarray
+
+
+class DGP(NamedTuple):
+    """DGP for simulation."""
+
+    name: str
+    main_effect: Callable[[pd.DataFrame], np.ndarray] | functools.partial
+    treatment_effect: Callable[[pd.DataFrame], np.ndarray] | functools.partial
+    propensity_score: Callable[[pd.DataFrame], np.ndarray] | functools.partial

### 2024-08-16 src/topics_causal_inf/config.py

@@ -2,12 +2,27 @@
 
 from pathlib import Path
 
+import numpy as np
+
+from topics_causal_inf.define_dgps import DGP3, DGP4, DGP5, DGP6
+
 SRC = Path(__file__).parent.resolve()
 ROOT = SRC.joinpath("..", "..").resolve()
 
 BLD = ROOT.joinpath("bld").resolve()
 
+WGAN_GEN = SRC / "wgan_generated"
 
 DOCUMENTS = ROOT.joinpath("documents").resolve()
 
-TEMPLATE_GROUPS = ["marital_status", "highest_qualification"]
+RNG = np.random.default_rng()
+
+DIMS_TO_RUN = [10, 15, 30]
+
+DGPS_TO_RUN = [DGP3, DGP4, DGP5, DGP6]
+
+NUM_TREES = 100
+SUBSAMPLE_SHARE = 0.2
+N_OBS = 10_000
+N_SIM = 40
+N_SPLITS = 10

### 2024-08-16 src/topics_causal_inf/define_dgps.py

@@ -1,0 +1,57 @@
+from functools import partial
+
+import numpy as np
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.utilities import (
+    main_constant,
+    main_linear,
+    prop_beta,
+    prop_cons,
+    tau_constant,
+    tau_heterog,
+)
+
+DGP1 = DGP(
+    name="dgp1",
+    main_effect=main_linear,
+    propensity_score=partial(prop_beta, a=2, b=4),
+    treatment_effect=partial(tau_constant, tau=0),
+)
+
+DGP2 = DGP(
+    name="dgp2",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_heterog, x_range=np.arange(0, 2), a=20, b=1 / 3),
+)
+
+DGP3 = DGP(
+    name="dgp3",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_heterog, x_range=np.arange(0, 2), a=12, b=1 / 2),
+)
+
+DGP4 = DGP(
+    name="dgp4",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_heterog, x_range=np.arange(0, 4), a=12, b=1 / 2),
+)
+
+DGP5 = DGP(
+    name="dgp5",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_heterog, x_range=np.arange(0, 8), a=12, b=1 / 2),
+)
+
+DGP6 = DGP(
+    name="dgp6",
+    main_effect=partial(main_constant, main=0),
+    propensity_score=partial(prop_cons, prop=0.5),
+    treatment_effect=partial(tau_constant, tau=0),
+)
+
+DGPS = [DGP1, DGP2, DGP3, DGP4, DGP5, DGP6]

### 2024-08-16 src/topics_causal_inf/generic_ml/generic_ml.py

@@ -1,0 +1,346 @@
+"""Simple generic ML implementation in Python."""
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+import statsmodels.api as sm  # type: ignore[import-untyped]
+from econml.grf import CausalForest  # type: ignore[import-untyped]
+from formulaic import model_matrix  # type: ignore[import-untyped]
+from scipy.stats import norm  # type: ignore[import-untyped]
+from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
+from sklearn.model_selection import train_test_split  # type: ignore[import-untyped]
+
+from topics_causal_inf.classes import GenericMLResult
+
+
+def generic_ml(
+    data: pd.DataFrame,
+    n_splits: int,
+    alpha: float,
+    ml_learner: tuple[RegressorMixin, RegressorMixin],
+    strategy: str = "blp_weighted_residual",
+) -> GenericMLResult:
+    """GenericML interface."""
+    # Input checks
+    _check_strategy(strategy)
+    _check_required_covs(data)
+    _check_pscore_range(data["p_z"])
+
+    # Prepare inputs
+
+    # Inference algorithm
+    blp_params, blp_se, ml_fitted_d0, ml_fitted_d1, gates_params, gates_se, clan = (
+        inference_algorithm(
+            data,
+            n_splits,
+            strategy,
+            ml_learner,
+        )
+    )
+
+    # BLP Compute t-values, lower and upper (1-alpha) confidence intervals
+    blp_tvals = blp_params / blp_se
+    blp_pvals = 2 * (1 - norm.cdf(np.abs(blp_tvals)))
+    blp_ci_lo = blp_params - norm.ppf(1 - alpha / 2) * blp_se
+    blp_ci_hi = blp_params + norm.ppf(1 - alpha / 2) * blp_se
+
+    # GATES: Compute t-values, lower and upper (1-alpha) confidence intervals
+    gates_tvals = gates_params / gates_se
+    gates_pvals = 2 * (1 - norm.cdf(np.abs(gates_tvals)))
+    gates_ci_lo = gates_params - norm.ppf(1 - alpha / 2) * gates_se
+    gates_ci_hi = gates_params + norm.ppf(1 - alpha / 2) * gates_se
+
+    # Find position of median of alpha_2 to pick a "median" machine learner
[…]

### 2024-08-16 src/topics_causal_inf/generic_ml/simulation.py

@@ -1,0 +1,201 @@
+"""Simulation function for generic_ml using WA 2018 DGP."""
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+from econml.grf import CausalForest  # type: ignore[import-untyped]
+from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.generic_ml.generic_ml import classification_analysis, generic_ml
+from topics_causal_inf.utilities import (
+    data_wager_athey_2018,
+)
+
+
+def simulation(
+    n_sims: int,
+    n_obs: int,
+    dim: int,
+    dgp: DGP,
+    n_splits: int,
+    ml_learner: tuple[RegressorMixin, RegressorMixin],
+    rng: np.random.Generator,
+) -> pd.DataFrame:
+    """Run simulation using the generic_ml approach and WA 2018 DGP."""
+    mse_blp = np.zeros(n_sims)
+    gates = np.zeros((n_sims, 5))
+    clans = np.zeros((n_sims, 2, dim))
+    blp_ci_lo = np.zeros((n_sims, 2))
+    blp_ci_hi = np.zeros((n_sims, 2))
+    gates_ci_lo = np.zeros((n_sims, 5))
+    gates_ci_hi = np.zeros((n_sims, 5))
+
+    for i in range(n_sims):
+        (
+            mse_blp[i],
+            gates[i, :],
+            clans[i, :, :],
+            blp_ci_lo[i],
+            blp_ci_hi[i],
+            gates_ci_lo[i, :],
+            gates_ci_hi[i, :],
+        ) = _single_experiment(
+            n_obs=n_obs,
+            dim=dim,
+            n_splits=n_splits,
+            dgp=dgp,
+            ml_learner=ml_learner,
+            rng=rng,
+        )
+
+    out = pd.DataFrame(mse_blp, columns=["mse_blp"])
+
+    out["dgp"] = dgp.name
+    out["n_obs"] = n_obs
+    out["n_sims"] = n_sims
+    out["dim"] = dim
+
+    beta_1, beta_2 = _simulate_true_blp(
+        dgp=dgp,
+        dim=dim,
+        ml_learner=ml_learner,
+        rng=rng,
+    )
+
+    out["true_blp_beta_1"] = beta_1
+    out["true_blp_beta_2"] = beta_2
[…]

### 2024-08-16 src/topics_causal_inf/generic_ml/task_generic_ml_plots.py

@@ -1,0 +1,173 @@
+from pathlib import Path
+from typing import Annotated
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+import plotly.express as px  # type: ignore[import-untyped]
+import plotly.graph_objects as go  # type: ignore[import-untyped]
+from plotly.subplots import make_subplots  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.config import BLD, DGPS_TO_RUN
+from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
+
+PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
+
+
+ID_TO_KWARGS_GATES = {
+    dgp.name: {
+        "dgp": dgp,
+        "path_to_plot": BLD
+        / "generic_ml"
+        / "plots"
+        / f"generic_ml_gates_{dgp.name}.png",
+    }
+    for dgp in DGPS_TO_RUN
+}
+
+for _id, kwargs in ID_TO_KWARGS_GATES.items():
+
+    @task(id=_id, kwargs=kwargs)
+    def task_generic_ml_plot_gates(
+        dgp: DGP,
+        path_to_plot: Annotated[Path, Product],
+        path_to_results: list[Path] = PATHS_TO_RESULTS,
+    ):
+        """Task for generic_ml plot gates."""
+        # Load results
+        res = pd.concat([pd.read_pickle(path) for path in path_to_results])
+
+        res = res[res["dgp"] == dgp.name]
+
+        # Plot Figure
+        fig = go.Figure()
+
+        colors = px.colors.qualitative.Plotly
+
+        for i in range(5):
+            color = colors[i % len(colors)]  # Cycle through the color palett
+            fig.add_trace(
+                go.Histogram(
+                    x=res[f"gate_{i}"],
+                    histnorm="probability",
+                    name=f"Group {i}",
+                    marker_color=color,
+                ),
+            )
+            fig.add_vline(
+                x=np.mean(res[f"true_gate_{i}"]),
+                line_dash="dash",
+                line_color=color,
+            )
+
+        fig.update_layout(
+            barmode="overlay",
+            legend={
+                "yanchor": "top",
+                "y": -0.2,  # Position the legend below the plot area
+                "xanchor": "center",
+                "x": 0.5,
+                "orientation": "h",
+            },
+        )
+        fig.update_traces(opacity=0.75)
+
[…]

### 2024-08-16 src/topics_causal_inf/generic_ml/task_generic_ml_sim.py

@@ -1,0 +1,77 @@
+"""Tasks for generic_ml simulation."""
+
+from pathlib import Path
+from typing import Annotated, NamedTuple
+
+import numpy as np
+from econml.grf import CausalForest  # type: ignore[import-untyped]
+from pytask import Product, task
+from sklearn.base import RegressorMixin  # type: ignore[import-untyped]
+from sklearn.ensemble import RandomForestRegressor  # type: ignore[import-untyped]
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.config import (
+    BLD,
+    DGPS_TO_RUN,
+    DIMS_TO_RUN,
+    N_OBS,
+    N_SIM,
+    N_SPLITS,
+    NUM_TREES,
+    RNG,
+    SUBSAMPLE_SHARE,
+)
+from topics_causal_inf.generic_ml.simulation import simulation
+
+
+class _Arguments(NamedTuple):
+    dim: int
+    dgp: DGP
+    path_to_res: Path
+    ml_learner: tuple[RegressorMixin, RegressorMixin]
+    n_obs: int = N_OBS
+    n_sims: int = N_SIM
+    n_splits: int = N_SPLITS
+    rng: np.random.Generator = RNG
+
+
+ID_TO_KWARGS = {
+    f"genml_{dgp.name}_{dim}": _Arguments(
+        dim=dim,
+        dgp=dgp,
+        path_to_res=BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl",
+        ml_learner=(
+            RandomForestRegressor(n_estimators=NUM_TREES, max_samples=SUBSAMPLE_SHARE),
+            CausalForest(n_estimators=NUM_TREES, max_samples=SUBSAMPLE_SHARE),
+        ),
+    )
+    for dgp in DGPS_TO_RUN
+    for dim in DIMS_TO_RUN
+    if not (dgp.name == "dgp5" and dim < 8)  # noqa: PLR2004
+}
+
+for id_, kwargs in ID_TO_KWARGS.items():
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_generic_ml_simulation(
+        dim: int,
+        dgp: DGP,
+        n_obs: int,
+        n_sims: int,
+        n_splits: int,
+        rng: np.random.Generator,
+        ml_learner: tuple[RegressorMixin, RegressorMixin],
+        path_to_res: Annotated[Path, Product],
+    ) -> None:
[…]

### 2024-08-16 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -1,0 +1,125 @@
+"""Tasks for generic_ml simulation."""
+
+from pathlib import Path
+from typing import Annotated, NamedTuple
+
+import pandas as pd  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN
+from topics_causal_inf.utilities import clean_tex_table
+
+
+class _Arguments(NamedTuple):
+    path_to_table: Path
+    path_to_res: list[Path]
+
+
+# For each dgp collect all results for all dim values
+ID_TO_KWARGS = {
+    f"{dgp.name}_mse": _Arguments(
+        path_to_table=BLD / "generic_ml" / "tables" / f"generic_ml_{dgp.name}.tex",
+        path_to_res=[
+            BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl"
+            for dim in DIMS_TO_RUN
+        ],
+    )
+    for dgp in DGPS_TO_RUN
+}
+
+for id_, kwargs in ID_TO_KWARGS.items():
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_generic_ml_tables(
+        path_to_res: list[Path],
+        path_to_table: Annotated[Path, Product],
+    ) -> None:
+        """Task for generic_ml result tables."""
+        # Load results
+        res = pd.concat([pd.read_pickle(path) for path in path_to_res])
+
+        # Create table
+        table = res.groupby(["dim"]).apply(
+            lambda x: pd.Series(
+                {
+                    "mean": x["mse_blp"].mean(),
+                    "std": x["mse_blp"].std(),
+                },
+            ),
+        )
+
+        # Save table
+        out = table.to_latex(
+            float_format="%.3f",
+            header=False,
+        )
+
+        # Remove all rows starting with "\begin{tabular}" and "\end{tabular}"
+        out = clean_tex_table(out)
+
+        # Write to file
+        path_to_table.write_text(out)
+
+
+ID_TO_KWARGS_COVERAGE = {
+    f"{dgp.name}_blp_gates_coverage": _Arguments(
+        path_to_table=BLD
+        / "generic_ml"
+        / "tables"
+        / f"generic_ml_blp_gates_coverage_{dgp.name}.tex",
+        path_to_res=[
[…]

### 2024-08-16 src/topics_causal_inf/utilities.py

@@ -1,0 +1,175 @@
+"""Utilities used throughout the project."""
+
+import re
+from copy import copy
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+from scipy.stats import beta  # type: ignore[import-untyped]
+from wgan import DataWrapper, Generator  # type: ignore[import-untyped]
+
+from topics_causal_inf.classes import DGP
+
+
+def data_wager_athey_2018(
+    n_obs: int,
+    dim: int,
+    dgp: DGP,
+    rng: np.random.Generator,
+) -> pd.DataFrame:
+    """Simulate data from one of the DGPs in Wager and Athey 2018 JASA.
+
+    Arguments:
+        n_obs: Number of observations.
+        dim: Dimension of covariates.
+        dgp: DGP to simulate data from.
+        rng: Random number generator.
+
+    Returns:
+        pd.DataFrame: Simulated data.
+    """
+    _check_dgp_and_dim_compatible(dgp, dim)
+
+    out = pd.DataFrame(index=range(n_obs))
+
+    for i in range(dim):
+        out[f"x{i}"] = rng.uniform(size=n_obs)
+
+    out["p_z"] = dgp.propensity_score(out)
+    out["d"] = rng.binomial(1, out["p_z"])
+
+    out["y"] = (
+        dgp.main_effect(out)
+        + dgp.treatment_effect(out) * out["d"]
+        + rng.normal(size=n_obs)
+    )
+
+    return out
+
+
+def tau_constant(data: pd.DataFrame, tau: float) -> np.ndarray:
+    """Treatment effect: tau(x) = 0."""
+    return np.ones(data.shape[0]) * tau
+
+
+def main_linear(data: pd.DataFrame) -> np.ndarray:
+    """Main effect: m(x) = 2x_1 - 1."""
+    return 2 * data["x1"] - 1
+
+
+def main_constant(data: pd.DataFrame, main: float) -> np.ndarray:
+    """Main effect: m(x) = main."""
+    return np.ones(data.shape[0]) * main
+
+
+def prop_beta(data: pd.DataFrame, a: float, b: float) -> np.ndarray:
+    """Propensity score: e(x) = 0.25 * (1 + beta_(a, b)(x_1))."""
+    return 0.25 * (1 + beta.pdf(data["x1"], a, b))
+
+
+def _xi(sr: pd.Series, a: float, b: float) -> pd.Series:
+    """Helper function: xi(x) = 1 + (1 + exp(-a * (x - b)))^-1."""
[…]

### 2024-08-16 src/topics_causal_inf/wa_replication/task_wa_replication.py

@@ -1,0 +1,98 @@
+"""Tasks for running WA 2018 replication simulations."""
+
+import pickle
+from pathlib import Path
+from typing import Annotated, NamedTuple
+
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.config import (
+    BLD,
+    DGPS_TO_RUN,
+    DIMS_TO_RUN,
+    N_OBS,
+    N_SIM,
+    NUM_TREES,
+    RNG,
+    SUBSAMPLE_SHARE,
+    WGAN_GEN,
+)
+from topics_causal_inf.wa_replication.wa_replication import simulation
+
+
+class _Arguments(NamedTuple):
+    dgp: DGP
+    path_to_data: Path
+    data_generator: str
+    n_sim: int = N_SIM
+    n_obs: int = N_OBS
+    dim: int = 10
+    rng: np.random.Generator = RNG
+    num_trees: int = NUM_TREES
+    sample_fraction: float = SUBSAMPLE_SHARE
+
+
+DG_TO_RUN = ["standard", "wgan"]
+
+ID_TO_KWARGS = {
+    f"{dgp.name}_{dim}_{dgen}": _Arguments(
+        dgp=dgp,
+        path_to_data=BLD
+        / "wa_replication"
+        / "sims"
+        / f"wa_replication_{dgp.name}_dim{dim}_{dgen}.pkl",
+        dim=dim,
+        data_generator=dgen,
+    )
+    for dgp in DGPS_TO_RUN
+    for dim in DIMS_TO_RUN
+    for dgen in DG_TO_RUN
+    if not (dgen == "wgan" and (dim != 10 or dgp.name == "dgp6"))  # noqa: PLR2004
+}
+
+for id_, kwargs in ID_TO_KWARGS.items():
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_wa_replication_sims(
+        dgp: DGP,
+        path_to_data: Annotated[Path, Product],
+        n_sim: int,
+        n_obs: int,
+        dim: int,
+        data_generator: str,
+        rng: np.random.Generator,
+        num_trees: int,
+        sample_fraction: float,
+    ) -> None:
+        """Task for WA 2018 replication simulations."""
+        if data_generator == "wgan":
+            wgan_train_data = pd.read_feather(WGAN_GEN / f"df_{dgp.name}.feather")
+            generators = pickle.load(
[…]

### 2024-08-16 src/topics_causal_inf/wa_replication/task_wa_replication_tables.py

@@ -1,0 +1,77 @@
+"""Tasks for running WA 2018 replication simulations."""
+
+from pathlib import Path
+from typing import Annotated, NamedTuple
+
+import pandas as pd  # type: ignore[import-untyped]
+from pytask import Product, task
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.config import BLD
+from topics_causal_inf.utilities import clean_tex_table
+from topics_causal_inf.wa_replication.task_wa_replication import (
+    DGPS_TO_RUN,
+    ID_TO_KWARGS,
+)
+
+
+class _Arguments(NamedTuple):
+    dgp: DGP
+    path_to_results: list[Path]
+    path_to_table: Path
+
+
+PATH_TO_RESULTS = [args.path_to_data for _, args in ID_TO_KWARGS.items()]
+
+ID_TO_KWARGS_TABLES = {
+    f"{dgp.name}": _Arguments(
+        dgp=dgp,
+        path_to_results=PATH_TO_RESULTS,
+        path_to_table=BLD
+        / "wa_replication"
+        / "tables"
+        / f"wa_replication_{dgp.name}.tex",
+    )
+    for dgp in DGPS_TO_RUN
+}
+
+
+for id_, kwargs in ID_TO_KWARGS_TABLES.items():
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_wa_replication_table(
+        dgp: DGP,
+        path_to_results: dict[str, Path],
+        path_to_table: Annotated[Path, Product],
+    ) -> None:
+        """Task for WA 2018 replication simulations."""
+        # Combine results from different dimensions and data generators into single df
+        res = pd.concat(
+            [pd.read_pickle(path) for path in path_to_results],
+        )
+
+        res = res[res["dgp"] == dgp.name]
+        res = res.drop(columns=["dgp"])
+        if "mse_cf" in res.columns:
+            res = res.drop(columns=["mse_cf"])
+
+        res_to_table = res.groupby(["data_generator", "dim"]).agg(
+            ["mean", "std"],
+        )
+
+        out = res_to_table.to_latex(
+            formatters={
+                ("mse", "mean"): "{:,.2f}".format,
+                ("mse", "std"): "{:,.2f}".format,
+                ("mse_pop_cate", "mean"): "{:,.2f}".format,
+                ("mse_pop_cate", "std"): "{:,.2f}".format,
+                ("coverage", "mean"): "{:,.2f}".format,
[…]

### 2024-08-16 src/topics_causal_inf/wa_replication/wa_replication.py

@@ -1,0 +1,201 @@
+import numpy as np
+import pandas as pd  # type: ignore[import-untyped]
+from econml.grf import CausalForest  # type: ignore[import-untyped]
+from wgan import DataWrapper, Generator  # type: ignore[import-untyped]
+
+from topics_causal_inf.classes import DGP
+from topics_causal_inf.utilities import data_wager_athey_2018, data_wgan
+
+
+def simulation(
+    n_sim: int,
+    n_obs: int,
+    dim: int,
+    num_trees: int,
+    sample_fraction: float,
+    dgp: DGP,
+    rng: np.random.Generator,
+    data_generator: str,
+    alpha: float = 0.05,
+    wgan_train_data: pd.DataFrame | None = None,
+    generators: list[Generator] | None = None,
+    data_wrappers: list[DataWrapper] | None = None,
+) -> pd.DataFrame:
+    """Run simulation using the generic_ml approach and WA 2018 DGP.
+
+    Arguments:
+        n_sim: Number of simulations.
+        n_obs: Number of observations.
+        dim: Dimension of covariates.
+        num_trees: Number of trees in the causal forest.
+        sample_fraction: Fraction of samples to use in each tree.
+        dgp: DGP to simulate data from.
+        rng: Random number generator.
+        data_generator: Data generator to use.
+        wgan_train_data: Data used to train the WGAN.
+        generators: Generators for WGAN data generator.
+        alpha: Significance level; 1-alpha is the (asymptotic) coverage probability.
+        data_wrappers: Data wrappers for WGAN data generator.
+
+    Returns:
+        pd.DataFrame: Mean squared error of treatment effect prediction.
+
+
+    """
+    if data_generator == "wgan":
+        _check_wgan_args(wgan_train_data, generators, data_wrappers)
+    mse = np.zeros(n_sim)
+    mse_pop_cate = np.zeros(n_sim)
+    coverage = np.zeros(n_sim)
+
+    if data_generator == "wgan":
+        pop_data = _generate_pop_data(
+            wgan_train_data=wgan_train_data,
+            data_generators=generators,  # type: ignore[arg-type]
+            data_wrappers=data_wrappers,  # type: ignore[arg-type]
+        )
+        pop_cate = _estimate_cf_cate(pop_data, dim)
+    else:
+        pop_data = None
+        pop_cate = None
+
+    for i in range(n_sim):
[…]

### 2024-08-16 tests/generic_ml/test_generic_ml.py

@@ -1,0 +1,87 @@
+"""Tests for generic_ml and related functions."""
+
+import pandas as pd  # type: ignore[import-untyped]
+import pytest
+from econml.grf import CausalForest  # type: ignore[import-untyped]
+from sklearn.ensemble import RandomForestRegressor  # type: ignore[import-untyped]
+from topics_causal_inf.config import RNG
+from topics_causal_inf.define_dgps import DGPS
+from topics_causal_inf.generic_ml.generic_ml import generic_ml, ml_proxy
+from topics_causal_inf.generic_ml.simulation import simulation
+from topics_causal_inf.utilities import data_wager_athey_2018
+
+ML_LEARNER = (RandomForestRegressor(), RandomForestRegressor())
+
+
+def test_negative_pscore_raises_error():
+    data = pd.DataFrame()
+    data["y"] = RNG.normal(size=100)
+    data["d"] = RNG.choice([0, 1], size=100)
+    data["x0"] = RNG.normal(size=100)
+    data["p_z"] = RNG.normal(loc=-10, size=100)
+
+    with pytest.raises(ValueError, match="Propensity score must be in the range"):
+        generic_ml(data, 2, 0.05, ml_learner=ML_LEARNER)
+
+
+def test_generic_ml_runs():
+    data = pd.DataFrame()
+    data["y"] = RNG.normal(size=100)
+    data["d"] = RNG.choice([0, 1], size=100)
+    data["x0"] = RNG.normal(size=100)
+    data["p_z"] = RNG.uniform(size=100)
+
+    for strategy in ["blp_weighted_residual", "blp_ht_transform"]:
+        generic_ml(
+            data=data,
+            n_splits=2,
+            alpha=0.05,
+            ml_learner=ML_LEARNER,
+            strategy=strategy,
+        )
+
+
+def test_generic_ml_returns_positive_param_se():
+    data = pd.DataFrame()
+    data["y"] = RNG.normal(size=100)
+    data["d"] = RNG.choice([0, 1], size=100)
+    data["x0"] = RNG.normal(size=100)
+    data["p_z"] = RNG.uniform(size=100)
+
+    out = generic_ml(data, 2, 0.05, ml_learner=ML_LEARNER)
+
+    assert (out.blp_se > 0).all()
+
+
+def test_ml_proxy_predictions_not_constant():
[…]

### 2024-08-16 tests/wa_replication/test_wa_replication.py

@@ -1,0 +1,40 @@
+import pickle
+from pathlib import Path
+
+import pandas as pd  # type: ignore[import-untyped]
+from topics_causal_inf.config import RNG, WGAN_GEN
+from topics_causal_inf.define_dgps import DGP3
+from topics_causal_inf.wa_replication.wa_replication import (
+    simulation,
+)
+
+
+def test_simulation_runs() -> None:
+    for dgp in [DGP3]:
+        for dgen in ["standard"]:
+            if dgen == "wgan":
+                wgan_train_data = pd.read_feather(WGAN_GEN / f"df_{dgp.name}.feather")
+                generators = pickle.load(
+                    Path.open(WGAN_GEN / f"generators_{dgp.name}.pkl", "rb"),
+                )
+                data_wrappers = pickle.load(
+                    Path.open(WGAN_GEN / f"data_wrapper_{dgp.name}.pkl", "rb"),
+                )
+            else:
+                wgan_train_data = None
+                generators = None
+                data_wrappers = None
+
+            simulation(
+                n_sim=2,
+                n_obs=1_000,
+                dim=4,
+                num_trees=20,
+                sample_fraction=0.1,
+                dgp=dgp,
+                rng=RNG,
+                data_generator=dgen,
+                generators=generators,
+                data_wrappers=data_wrappers,
+                wgan_train_data=wgan_train_data,
+            )

### 2024-08-17 documents/paper.tex
Compute wgan coverage for estimated CATE; add to paper.

@@
 I follow~\cite{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
+Throughout, I construct confidence intervals with $95\%$ nominal coverage.
 To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation of simulation runs.
@@
 \subsection{Results: WA Replication and WGAN}
+I report results for the causal forest approach in Table~\ref*{table:wa_rep_wgan}.
+First, the estimates for MSE --- using the true CATE function --- in column 2 are very close to the original results in~\cite{wager2018estimation}.
+I find a MSE for 0.02 for $d=10$ with an SD smaller than 0.01 while they find MSE estimates ranging from 0.02 to 0.03 for dimensions ranging from 2 to 8 and a similar SD\@.
+Notably, when I increase the dimensions to 15 and 30, respectively, I find comparable MSE estimates.
+This shows, that causal forests are able to perform well in a sparse setting where the subset of covariates that are pure noise is potentially large.
+The fourth row reports in Panel A reports simulation results based on an intermediate WGAN generator.
+The MSE is more than three times as large as the directly simulated DGP when using the true CATE to compute the MSE (\textit{MSE True}), but delivers a similar MSE estimate when using an estimated CATE based on a large population draw (\textit{MSE Pop}).
+Hence, using a WGAN generated DGP can in principle work well for simulation purposes when true CATE has a low-dimensional structure.
+Turning to coverage in the last columns, I find that coverage is slightly above nominal level for all dimensions, which is different to the original results reported in~\cite{wager2018estimation}, who report coverage levels far below nominal level (e.g.
+0.73 for $d=8$).
+% [TODO Wgan coverage level after new simulation run using estimated CATE.]
 
+Panels B and C report results for a CATE with higher dimension. As expected the performance of the Causal Forest significantly deteriorates.
+MSE significantly increases, in particular for Panel C where the CATE depends on 8 covariates.
+Even with only 4 covariates, coverage is sinificantly below the nominal level ranging from 0.83 to 0.76. For 8 covariates coverage is below 0.5 for all dimensions.
+The WGAN generator approach result in similar MSE estimates when using the estimated CATE to compute MSE (0.22 vs. 0.19 using the standard approach).
+However, for the more complicated CATE in Panel C, the WGAN approach suggests an MSE that is far below the actual MSE using the standard approach (44.99 vs. 133.81).
+This is consistent with the previous argument, that the generator cannot preserve the high-dimensional structure. Hence, the CATE estimated based on a large population draw will be ``too simple''.
+For the remaining simulations using the generic machine learning approach I stick to the standard approach.
+Lastly, Panel D reports result for a Placebo CATE which is zero everywhere. The MSE is small and coverage is above nominal level.
+Hence I find no indication for the claim made in connection with Figure 1 in~\cite{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.
+\footnote{They, however, used a random forest directly (i.e.\ with a different target function and without modifications like honesty).
+Also, they probably reported out-of-the-box estimates which, since random forests are meant to address problems with more than one covariate, are prone to overfitting in this case.
+Using a random forest with cross-validation easily avoids overfitting.}
 \begin{table}
-    \caption{Simulation Results: WA Replication and WGAN}
+    \caption{Simulation Results: WA Replication and WGAN}\label{table:wa_rep_wgan}
     \begin{threeparttable}
@@
             \toprule
-            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
             \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
@@
 
-            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
             \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
@@
 
-            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
             \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
@@
 MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP\@.
+            Coverage for $95\%$ confidence intervals reported.
             Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
@@
 \subsection{Results: Generic Machine Learning}
+I next turn to simulation results for the generic machine learning approach.
+In the auxiliary sample I use a random forest learner for the baseline conditional average $b_0(z)$ and a causal forest for the CATE $s_0(z)$.
+Results are median aggregated over 10 sample splits. For example, I use the median of the 10 parameter estimate for the heterogeneity loading $\hat{\beta_2^s}, s=1,\ldots,10$.
+I also use the trained model $s(z)$ from the median slope parameter to evaluate the performance, i.e.\ construct BLP predictions.
+
+\paragraph*{MSE}
+While this approach does not directly target the CATE, it is still useful to evaluate the MSE\@.
+After all, if the MSE is justified by some decision-theoretic approach, we might wonder whether trading off model misspecification for estimation precision results in smaller loss.
+The results in Table~\ref*{table:generic_ml_cate} show that the MSE for estimating the true CATE is significantly worse compared to the causal forest approach.
+While the BLP is of course not linear in the original covaiates, it seems like the gain estimation precision might still be offset by the approximation.
+This is expected for DGPs with a low-dimensional CATE in which causal forests perform well (Panels A and B), but I would have expected a relatively better performance in Panel C.
+One potential concern is that I used too few sample splits (the paper suggests 250 instead of 10), but this was computationally not feasible at this point.
+The generic ML does in fact perform better in the placebo setting in Panel D, where the MSE is only a tenth compared to that of the causal forest.
 
 \begin{table}
-    \caption{Simulation Results: ML Generic --- Cate Estimation}
+    \caption{Simulation Results: Generic ML --- Cate Estimation}\label{table:generic_ml_cate}
     \begin{threeparttable}
@@
         \toprule
-        \multicolumn{3}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+        \multicolumn{3}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
         \input{../bld/generic_ml/tables/generic_ml_dgp3.tex} \\
@@
 
-        \multicolumn{3}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+        \multicolumn{3}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
         \input{../bld/generic_ml/tables/generic_ml_dgp4.tex} \\
@@
 
-        \multicolumn{3}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+        \multicolumn{3}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
         \input{../bld/generic_ml/tables/generic_ml_dgp5.tex} \\
@@
         \textit{Dimension} refers to the number of covariates in the DGP\@.
-        MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
+        MSE is calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
         Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
@@
 \begin{table}
-    \caption{Simulation Results: ML Generic --- BLP and GATES Coverage}
+    \caption{Simulation Results: Generic ML --- BLP and GATES Coverage}
     \begin{threeparttable}
@@
             \toprule
-            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
+            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
             \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp3.tex} \\
@@
 
-            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
+            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
             \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp4.tex} \\
@@
 
-            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
+            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
             \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp5.tex} \\
@@
             \textit{Dimension} refers to the number of covariates in the DGP\@.
-            MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
+            Coverage is calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
+            Coverage for $95\%$ confidence intervals reported.
             Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp3.png}
-         \caption{Original WA DGP 3}\label{fig_gates:dgp3}
+         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_gates:dgp3}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp4.png}
-         \caption{High-dimensional --- 4 Covariates}\label{fig_gates:dgp4}
+         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_gates:dgp4}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp5.png}
-         \caption{High-dimensional --- 8 Covariates}\label{fig_gates:dgp5}
+         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_gates:dgp5}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp3.png}
-         \caption{Original WA DGP 3}\label{fig_clans:dgp3}
+         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_clans:dgp3}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp4.png}
-         \caption{High-dimensional --- 4 Covariates}\label{fig_clans:dgp4}
+         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_clans:dgp4}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp5.png}
-         \caption{High-dimensional --- 8 Covariates}\label{fig_clans:dgp5}
+         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_clans:dgp5}
      \end{subfigure}

### 2024-08-17 src/topics_causal_inf/wa_replication/wa_replication.py

@@ -143,7 +143,12 @@ def _experiment(
 
     # Calculate (expected) coverage
     ci_lo, ci_hi = cf.predict_interval(X=data_test[feature_names], alpha=alpha)
-    coverage = np.mean((ci_lo.flatten() <= true) & (true <= ci_hi.flatten()))
+    if data_generator == "wgan" and pop_cate is not None:
+        coverage = np.mean(
+            (ci_lo.flatten() <= true_pop_cate) & (true_pop_cate <= ci_hi.flatten()),
+        )
+    else:
+        coverage = np.mean((ci_lo.flatten() <= true) & (true <= ci_hi.flatten()))
 
     return mse, mse_pop_cate, coverage
 

### 2024-08-17 documents/paper.tex
Add simulation result description for generic ML.

@@
 
+\paragraph*{Coverage, GATES and CLANS}
+Table~\ref*{table:generic_ml_coverage} verifies coverage properties of the generic ML procedure for both BLP parameters $\beta_1, \beta_2$ and GATES $\mu_1, \ldots, \mu_5$.
+Groups for GATES are defined by quintiles of predicted treatment effects from low ($\mu_1$) to high ($\mu_5$).
+I mostly find coverage levels close to the nominal level of 95\%, although with only 40 simulation runs the accuracy is of course severely limited.
+% TODO describe coverage properties, currently some seem off?
+Figure~\ref*{fig:gates} illustrates the performance of the GATES procedure for the four different DGPs\@.
+It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($N=1,000,000$) sample draw.
+While quantitatively not precise, we can still see that generic ML procedure correctly estimates almost all GATES for all four DGPs.
+This is true even for the DGPs with increased heterogeneity dimension, as well as the placebo, where estimates are tightly centered around 0.
+
+Lastly, Figure~\ref*{fig:clans} shows estimates for the characteristics --- limited to covariates $x_1, \ldots, x_4$ --- of the least and most affected groups as predited by the ML proxy.
+If the CATE function does not depend on a covariate, the uniform distribution implies that the average characteristic for both groups is 0.5.
+This is confirmed in both Panels A (where $x_3$ and $x_4$ are noise) and the placebo DGP in Panel D (where all covariates are noise).
+The CLAN also correctly estimates the characteristics for least and most affected groups for covariates that determine heterogeneity, which can be seen in Panel A (for $x_3, x_4$) as well as Panels B and C.
+
 \begin{table}
@@
 \begin{table}
-    \caption{Simulation Results: Generic ML --- BLP and GATES Coverage}
+    \caption{Simulation Results: Generic ML --- BLP and GATES Coverage}\label{table:generic_ml_coverage}
     \begin{threeparttable}

### 2024-08-18 documents/paper.tex
Notes for conclusion.
Add BLP CI plot.

@@
 
+% TODO Describe causal forest?
+
 \subsection{Generic Machine Learning}
@@
 The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
-A similar question is investigated in Figure 1 of~\cite{chernozhukov2023genml}.
+A similar question is analyzed in Figure 1 of~\cite{chernozhukov2023genml}.
 
@@
 I mostly find coverage levels close to the nominal level of 95\%, although with only 40 simulation runs the accuracy is of course severely limited.
-% TODO describe coverage properties, currently some seem off?
+In particular, GATES and $\beta_1$ (corresponding to the ATE) have (near) correct coverage in most cases except for the very high-dimensional setting where coverage declines significantly for some parameters.
+Notably, $\beta_2$ has coverage of $0\%$ or close to it for almost all settings.
+To investigate this further, Figure~\ref*{fig:blp} plots the distribution of the lower and upper confidence interval points for both all four DGPs and both BLP parameters.
+While endpoints are to the left and right of $\beta_1$ in most cases except the high-dimensional ones, $\beta_2$ is always over-estimated.
+At this point I don't have a good explanation for this. Note the true parameter is always almost equal to $1$, since we have a highly heterogeneous treatment effects (except for the placebo setting).
+Hence, this could be a parameter at the boundary issue.
+
 Figure~\ref*{fig:gates} illustrates the performance of the GATES procedure for the four different DGPs\@.
@@
 
+% BLP Figure
+\begin{figure}
+    \caption{Simulation Results: Generic ML --- BLP Confidence Intervals}\label{fig:blp}
+
+    \centering
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp3.png}
+         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_blp:dgp3}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp4.png}
+         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_blp:dgp4}
+     \end{subfigure}
+
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp5.png}
+         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_blp:dgp5}
+     \end{subfigure}
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp6.png}
+         \caption{Placebo}\label{fig_blp:dgp6}
+     \end{subfigure}
+
+     \floatfoot{\textbf{Notes}: This figure reports simulation results for confidence intervals of BLP parameters of the generic machine learning approach.
+     For each of the four DGPs, it plots the distribution of both lower and upper confidence intervals for $\beta_1, \beta_2$ over the simulation runs.
+     Lower ends are displayed red, upper blue.
+     The dashed line indicates the true GATES using a large population draw. Sample size $N=10,000$ and $40$ simulation runs.
+     }
+
+\end{figure}
+
 % Gates Figure
@@
 
-% Gates Figure
+% CLANS Figure
 \begin{figure}
@@
 
+\begin{itemize}
+    \item Main difficulty: Informative DGP
+    \item Think of criterion relevant for decision
+    \item Computation power/time to run simulations
+    \item Combine WGAN and analytical CATE structure; still need to justify this.
+\end{itemize}
+
 \clearpage

### 2024-08-18 src/topics_causal_inf/generic_ml/task_generic_ml_plots.py

@@ -1,5 +1,5 @@
 from pathlib import Path
-from typing import Annotated
+from typing import Annotated, NamedTuple
 
 import numpy as np
 import pandas as pd  # type: ignore[import-untyped]
@@ -9,8 +9,15 @@ from plotly.subplots import make_subplots  # type: ignore[import-untyped]
 from pytask import Product, task
 
 from topics_causal_inf.classes import DGP
-from topics_causal_inf.config import BLD, DGPS_TO_RUN
+from topics_causal_inf.config import BLD, DGPS_TO_RUN, DIMS_TO_RUN
 from topics_causal_inf.generic_ml.task_generic_ml_sim import ID_TO_KWARGS
+
+
+class _Arguments(NamedTuple):
+    path_to_plot: Path
+    dgp: DGP
+    path_to_res: list[Path]
+
 
 PATHS_TO_RESULTS = [args.path_to_res for _, args in ID_TO_KWARGS.items()]
 
@@ -164,6 +171,81 @@ for _id, kwargs in ID_TO_KWARGS_CLANS.items():
 
         fig.update_traces(opacity=0.75)
 
+        fig.update_layout(
+            showlegend=False,
+        )
+
+        fig.update_layout(margin={"l": 10, "r": 10, "t": 10, "b": 10})
+
+        fig.write_image(path_to_plot)
+
+
+ID_TO_KWARGS_COVERAGE = {
+    f"{dgp.name}_blp_gates_coverage": _Arguments(
+        path_to_plot=BLD
+        / "generic_ml"
+        / "plots"
+        / f"generic_ml_blp_distribution_{dgp.name}.png",
+        dgp=dgp,
+        path_to_res=[
+            BLD / "generic_ml" / "sims" / f"generic_ml_{dgp.name}_dim{dim}.pkl"
+            for dim in DIMS_TO_RUN
+        ],
+    )
+    for dgp in DGPS_TO_RUN
+}
+
+
+for id_, kwargs in ID_TO_KWARGS_COVERAGE.items():  # type: ignore[assignment]
+
+    @task(id=id_, kwargs=kwargs)  # type: ignore[arg-type]
+    def task_generic_ml_blp_params(
+        path_to_res: list[Path],
+        dgp: DGP,
+        path_to_plot: Annotated[Path, Product],
+    ) -> None:
+        """Task for generic_ml BLP params plot."""
+        # Load results
[…]

### 2024-08-18 src/topics_causal_inf/generic_ml/task_generic_ml_tables.py

@@ -87,6 +87,15 @@ for id_, kwargs in ID_TO_KWARGS_COVERAGE.items():
         # Load results
         res = pd.concat([pd.read_pickle(path) for path in path_to_res])
 
+        for beta in ["beta_1", "beta_2"]:
+            if (res[f"blp_ci_lo_{beta}"] > res[f"blp_ci_hi_{beta}"]).mean() == 1:
+                res = res.rename(
+                    columns={
+                        f"blp_ci_lo_{beta}": f"blp_ci_hi_{beta}",
+                        f"blp_ci_hi_{beta}": f"blp_ci_lo_{beta}",
+                    },
+                )
+
         for i in range(5):
             res[f"covers_gate_{i}"] = (
                 res[f"gate_ci_lo_{i}"] < res[f"true_gate_{i}"]
@@ -99,8 +108,8 @@ for id_, kwargs in ID_TO_KWARGS_COVERAGE.items():
 
         # Create table
         # Get mean for all covers_gate_i and covers_blp_beta_i, each separately
-        params = [f"gate_{i}" for i in range(5)] + [
-            f"blp_{beta}" for beta in ["beta_1", "beta_2"]
+        params = [f"blp_{beta}" for beta in ["beta_1", "beta_2"]] + [
+            f"gate_{i}" for i in range(5)
         ]
 
         table = res.groupby(["dim"]).apply(

### 2024-08-18 documents/paper.tex
Add Introduction.

@@
 \section{Introduction}
+Estimating treatment effects has become a standard practice in applied econometris.
+Also, increasingly large datasets are utilized to investigate systematic heterogeneity.
+While this has the potential to increase our understanding of say the heterogneous effects of public policies,
+from a methodological perspective large datasets with many covariates come at the risk of increased researchers discretion.
+In this case, machine learning methods to systematically explore heterogeneity can be helpful to ``tie the researchers hands''.
+Furthermore, in high-dimensional settings with a potentially large number of irrelevant covariates, regularization built-in to many machine learning methods is required to deliver useful results.
+However, one concern is that many machine learning methods, like the very popular causal forests proposed in~\cite{wager2018estimation}, to estimate conditional average tretment effects rely on a low-dimensional structure of the conditional effects.
+Another approach proposed in~\cite{chernozhukov2023genml} is to only estimate low-dimensional features of conditional effects but utilizing machine learning methods in a first-step prediction exercise.
+The current paper investigates the performance of these two approaches for data generating processes both with low- and high-dimensional conditional treatment effect structures.
 

### 2024-08-18 documents/paper.tex
Add conclusion.

@@
 \documentclass[11pt, a4paper, leqno]{article}
-\usepackage{a4wide}
+\usepackage[margin=1in]{geometry}
+
+% \usepackage{a4wide}
 \usepackage[T1]{fontenc}
@@
 \section{Introduction}
-Estimating treatment effects has become a standard practice in applied econometris.
+Estimating treatment effects has become a standard practice in applied econometrics.
 Also, increasingly large datasets are utilized to investigate systematic heterogeneity.
@@
 \section{Conclusion}
-
-\begin{itemize}
-    \item Main difficulty: Informative DGP
-    \item Think of criterion relevant for decision
-    \item Computation power/time to run simulations
-    \item Combine WGAN and analytical CATE structure; still need to justify this.
-\end{itemize}
+One of the main difficulties for extending this project besides computational feasibility is the creation of an informative DGP\@.
+While the approach taken above allows to control the dimension of the CATE, the covariate structure is of course highly stylized (no discrete covariates, all uniform, all independent).
+I think most promising would be a hybrid approach that combines a WGAN generated DGP for the covariates and treatment assignment, with an analytical treatment effect structure.
+This still poses the challenge, however, of coming up with a high-dimensional CATE function that makes sense economically.
+Eventually, this all boils down to the question, how high-dimensional and non-sparse our objects of interest really are.
+Lastly, when moving to low-dimensional summaries of the CATE functional, it would be useful to have decision-theoretic framework to evaluate.
+This would allow to think more systematically about how to trade-off model approximations with statistical performance.
 

### 2024-08-18 documents/paper.tex
Fix typos and restructure text.

@@
     This term paper presents simulation results for two connected approaches to estimating conditional average treatment effects (CATE):
-    The \textit{causal forest} approach originally proposed in~\cite{wager2018estimation} and the \textit{generic machine learning} approach discussed in~\cite{chernozhukov2023genml}.
-    In a first part, I replicate simulation results for causal forest reported in the original paper that demonstrate the ability to estimate a sparse CATE function and perform valid inference.
-    I then continue to extend their analysis by lowering the degree of sparseness of the DGP and show that their method breaks down: RMSE is significantly increased and coverage is far below the nominal level.
+    The \textit{causal forest} approach originally proposed in~\citet{wager2018estimation} and the \textit{generic machine learning} approach discussed in~\citet{chernozhukov2023genml}.
+    In a first part, I replicate simulation results for the causal forest approach reported in the original paper that demonstrate the ability to estimate a sparse CATE function and perform valid inference.
+    I then continue to extend their analysis by lowering the degree of sparseness of the DGP and show that their method breaks down: MSE is significantly increased and coverage is far below the nominal level.
     I show that inference on the BLP is still possible using the generic machine learning approach.
-However, for the DGP I consider the RMSE is still considerably worse, illustrating a trade-off between estimation properties and informativeness of the inferential target.
-    As an additional point of interest I consider the design of high-dimensional monte carlo studies using the approach suggested by~\cite{athey2024wgan}.
+However, for the DGPs I consider the MSE is still considerably worse, illustrating a trade-off between estimation properties and informativeness of the inferential target.
+    As an additional point of interest I consider the design of high-dimensional monte carlo studies using the WGAN approach suggested by~\citet{athey2024wgan}.
 \end{abstract}
@@
 Estimating treatment effects has become a standard practice in applied econometrics.
-Also, increasingly large datasets are utilized to investigate systematic heterogeneity.
-While this has the potential to increase our understanding of say the heterogneous effects of public policies,
-from a methodological perspective large datasets with many covariates come at the risk of increased researchers discretion.
+Moreover, increasingly often large datasets are utilized to investigate systematic heterogeneity.
+While this has the potential to increase our understanding of, say, the heterogneous effects of public policies,
+from a methodological perspective large datasets with many covariates come at the risk of increased researcher discretion.
 In this case, machine learning methods to systematically explore heterogeneity can be helpful to ``tie the researchers hands''.
-Furthermore, in high-dimensional settings with a potentially large number of irrelevant covariates, regularization built-in to many machine learning methods is required to deliver useful results.
-However, one concern is that many machine learning methods, like the very popular causal forests proposed in~\cite{wager2018estimation}, to estimate conditional average tretment effects rely on a low-dimensional structure of the conditional effects.
-Another approach proposed in~\cite{chernozhukov2023genml} is to only estimate low-dimensional features of conditional effects but utilizing machine learning methods in a first-step prediction exercise.
+Furthermore, in high-dimensional settings with a potentially large number of irrelevant covariates, the regularization inherent to many machine learning methods is required to deliver useful results.
+However, one concern is that many machine learning methods to estimate conditional average tretment effects, like the very popular causal forests proposed in~\citet{wager2018estimation}, rely on a low-dimensional structure of the conditional effects.
+Another approach proposed in~\citet{chernozhukov2023genml} is to only estimate low-dimensional features of conditional effects but utilizing machine learning methods in a first-step prediction exercise.
 The current paper investigates the performance of these two approaches for data generating processes both with low- and high-dimensional conditional treatment effect structures.
@@
 Throughout we consider a setting with a binary treatment denoted by $D$ taking on values in $\{0, 1\}$.
-$Z$ will denote a vector of covariates that has dimension $p$. The sample size will be noted by $n$.
-We will consider generally consider a setting of \textit{unconfoundedness}, meaning that conditional on covariates the treatment is as-good-as randomly assigned.
-Denoting the outcomes obtained in states of the world with treatment realization $D=d$ by $Y(d)$, this means
+$Z$ will denote a vector of covariates that has dimension $d$. The sample size will be noted by $n$.
+We will generally consider a setting of \textit{unconfoundedness}, meaning that conditional on covariates the treatment is as-good-as randomly assigned.
+Denoting the potential outcomes obtained in states of the world with treatment realization $d$ by $Y(d)$, this means
 \begin{equation*}
@@
 \end{equation*}
-In the context of~\cite{chernozhukov2023genml} we will additionally make the assumption that the \textit{propensity score} $p(z) \equiv \Pr(D=1|Z=z)$ is a known function.
+In the context of~\citet{chernozhukov2023genml} we will additionally make the assumption that the \textit{propensity score} $p(z) \equiv \Pr(D=1|Z=z)$ is a known function.
 The leading case of this is a stratified experiment, with $p(z) = 0.5$ being the special case of a completely randomized experiment.
-Generally, the literature also often allows for an unknown propensity score which poses the problem of --- high-dimensional --- first-step estimation.
+Generally, the literature also often allows for an unknown propensity score which poses the problem of --- potentially high-dimensional --- first-step estimation.
 Note this is not the ``dimensionality'' issue considered here, which is about the complexity of conditional average treatment effects.
-The target of this literature is the \textit{conditional average treatment effect} (CATE), which is the average treatment effect for individuals with covariates $Z=z$:
+The target of this literature is the \textit{conditional average treatment effect} (CATE), which is the average treatment effect conditional on covariates $Z$:
 \begin{equation*}
@@
 Our goal is to estimate and perform inference on this quantity.
+In addition, we will also consider the baseline conditional average (BCA), which is defined as
+\begin{equation*}
+    BCA(Z) = b_0(Z) = \E[Y(0)|Z].
+\end{equation*}
 
-Ideally, we would like to have a procedure with the following \textit{properties}:
+Ideally, we would like to have a procedure for estimating the CATE with the following \textit{properties}:
 \begin{itemize}
@@
     \item[(2)] valid inference, i.e.\ confidence intervals, for $CATE(z)$;
-    \item[(3)] ability to adapt to high-dimensional covariate spaces (regimes where $p$ grows with $n$) without making sparsity assumptions;
+    \item[(3)] ability to adapt to high-dimensional covariate spaces (regimes where $d$ grows with $n$) without making sparsity assumptions;
     \item[(4)] no functional form restrictions on $s_0(z)$.
@@
 All proposed approaches then proceed by relaxing one or more of the requirements stated above.
-The following briefly summarizes some of the ideas, a more comprehensive list of reference can be found in~\cite{causalml2024}.
-I focus in particular on sparsity assumptions and approximations to the CATE function, because these are two ideas that are taken up in the generic machine learning approach discussed later.
+The following briefly summarizes some of the ideas, a more comprehensive list of reference can be found in~\citet{causalml2024}.
+I focus in particular on sparsity assumptions and approximations to the CATE function, because these are two ideas that are taken up in the generic machine learning approach.
 \paragraph*{Sparsity}
@@
 \begin{equation*}
-    \E[Y|Z] = \sum_{l=1}^{p_1}\alpha_l DX_l + \sum_{j=1}^{p_2}\beta_j W_j,
+    \E[Y|Z] = \sum_{l=1}^{d_1}\alpha_l DX_l + \sum_{j=1}^{d_2}\beta_j W_j,
 \end{equation*}
 where the covariates $Z$ are split into those determining systematic heterogeneity of treatment effects, $X$, and controls, $W$.
-In this model we have hence have $s_0(Z) = S_0(X) = \alpha'X$ under randomization. In these setups we can allow for both $p_1$ and $p_2$ to grow with the sample size.
+In this model we hence have $s_0(Z) = s_0(X) = \alpha'X$ under randomization. In these setups we can allow for both $d_1$ and $d_2$ to grow with the sample size.
 Hence, the CATE function can be high-dimensional at least for a subset of the covariates.
-For these linear models Lasso-type approaches adapted to causal inference (so called ``double'' lasso) will perform well, under a condition called \textit{approximate sparsity}.
+For these linear models lasso-type approaches adapted to causal inference (so called ``double'' lasso) will perform well, under a condition called \textit{approximate sparsity}.
 Intuitively, this condition is fulfilled by DGPs where the magnitude of the (ordered) coefficients $\alpha_l, \beta_l$, decays sufficiently fast.
@@
 \end{equation*}
-Assuming $p_1$, the dimension of heterogeneity, does not grow too fast and we have approximate sparsity, it is possible to show an asymptotic normal approximation, namely
+Assuming $d_1$, the dimension of heterogeneity, does not grow too fast and we have approximate sparsity, it is possible to show an asymptotic normal approximation, namely
 \begin{equation*}
@@
 \end{equation*}
-where $V$ can be consistently estimated. These results are shown in~\cite{belloni2015uniform} and~\cite{belloni2018uniformly}.
+where $V$ can be consistently estimated. These results are shown in~\citet{belloni2015uniform} and~\citet{belloni2018uniformly}.
 
 \paragraph*{Approximations to the CATE}
-~\cite{semenova2021debiased} propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
+~\citet{semenova2021debiased} propose to focus on a \textit{low-dimensional} subset $X$ of the covariates $Z$, which the researcher has to decide on ex ante.
 Then, the new target for inference becomes the \textit{partial} CATE $s_0(X)$.
 Their paper then simultaneously allows for \textit{unconfoundedness} which poses the challenge of estimating $p(Z)$ --- potentially using machine learning methods --- in a first step.
-In a second step, they the nfocus on a \textit{linear} approximation to the (partial) CATE\@.
+In a second step, they then focus on a \textit{linear} approximation to the (partial) CATE\@.
 Hence, they essentially use the following approximations to make the problem solvable:
@@
 \subsection{Generic Machine Learning}
-The generic ML approach allows want to place as little assumptions as possible on the data generating process, while maintaining valid inference.
+The generic ML approach wants to place as little assumptions as possible on the data generating process, while maintaining valid inference.
 This is achieved by (a) moving the inferential target to \textit{features} of the CATE (like the BLP) and (b) using sample splitting to derive inferential guarantees.
-Hence, this comes at the cost of potentially being able to only make statements about a (bad) approximation to the original inferential target as well as a loss in precision due to sample splitting.
+Hence, this comes at the cost of potentially being able to only make statements about an approximation to the original inferential target as well as a loss in precision due to sample splitting.
 However, there are a number of clear advantages: The researcher does not have to commit to a specific --- small --- set of covariates ex-ante or make sparsity assumptions.
-At the same time, the approach as discussed in~\cite{chernozhukov2023genml} does not allow for nuisance parameter estimation, i.e.\ requires a known propensity score.
+At the same time, the approach as discussed in~\citet{chernozhukov2023genml} does not allow for nuisance parameter estimation, i.e.\ requires a known propensity score.
 
@@
 \begin{equation*}
-    HY = \mu_1 + \mu_2(S(Z) - \E S(Z)) + \varepsilon, \qquad \E[(1, S(Z) - \E S(Z))'\varphi] = 0
+    HY = \mu_1 + \mu_2(S(Z) - \E S(Z)) + \varepsilon, \qquad \E[(1, S(Z) - \E S(Z))'\varepsilon] = 0
 \end{equation*}
 identifies the BLP parameters, i.e.\ $\mu_1=\beta_1$ and $\mu_2 = \beta_2$.
-Estmation can then proceed based on the empirical analogue.
+Estimation can then proceed based on the empirical analogue.
 While inference \textit{conditional} on the split is standard, unconditional inference has to rely on median aggregation of many splits (discussed below).
+Importantly, no assumptions that ensure validity of the machine learning methods need to be placed.
+Intuitively, if the first-stage machine learning proxy is pure noise --- which could be the case if the DGP is high-dimensional and dense ---
+$\beta_2 = 0$ and the BLP is just given by the ATE\@.
 
@@
 For example, we define five groups by the size of their predicted treatment effect $S(z)$.
-More generally, we can consider any finite partition of the real line $-\infty = I_0 < I_1 < \cdots < I_K = \infty$ and the define GATES by
+More generally, we can consider any finite partition of the real line $-\infty = I_0 < I_1 < \cdots < I_K = \infty$ and then define GATES by
 \begin{equation*}
@@
 
-\paragraph*{ML Method Selection}
-Because the approach works with \textit{any} first stage machine learning method, a researcher might consider constructing proxies from different methods.
-To choose a method, the authors propose a criterion that maximizes the (absolute) correlation between the true CATE $s_0(Z)$ and the ML proxy $S(Z)$.
-Remember we can identify $\beta_2$ by the strategy discussed above. Because $\beta_2$ is the BLP of $s_0(Z)$ given $S(Z)$ we have
-\begin{equation*}
-    \beta_2 = \frac{Cov[s_0(Z), S(Z)]}{Var(S(Z))}.
-\end{equation*}
-By squaring and rescaling we are at the following criterion
-\begin{equation*}
-    \Lambda := |\beta_2|^2 Var(S(Z)) = Corr(s_0(Z), S(Z))^2 Var(s_0(Z)).
-\end{equation*}
-Hence, maximizing $\Lambda$ by choosing a proxy $S(Z)$, maximizes the desired correlation.
-Note the left-hand side can be estimated from the data, hence the criterion is feasible.
+% \paragraph*{ML Method Selection}
+% Because the approach works with \textit{any} first stage machine learning method, a researcher might consider constructing proxies from different methods.
+% To choose a method, the authors propose a criterion that maximizes the (absolute) correlation between the true CATE $s_0(Z)$ and the ML proxy $S(Z)$.
+% Remember we can identify $\beta_2$ by the strategy discussed above. Because $\beta_2$ is the BLP of $s_0(Z)$ given $S(Z)$ we have
+% \begin{equation*}
+%     \beta_2 = \frac{Cov[s_0(Z), S(Z)]}{Var(S(Z))}.
+% \end{equation*}
+% By squaring and rescaling we are at the following criterion
+% \begin{equation*}
+%     \Lambda := |\beta_2|^2 Var(S(Z)) = Corr(s_0(Z), S(Z))^2 Var(s_0(Z)).
+% \end{equation*}
+% Hence, maximizing $\Lambda$ by choosing a proxy $S(Z)$, maximizes the desired correlation.
+% Note the left-hand side can be estimated from the data, hence the criterion is feasible.
 
@@
 where $Med(\cdot)$ denotes the median over datasplits and $p_A^+$ is a (one-sided) p-value in a given auxiliary sample $A$.
-Under regularity condition spelled out in the paper, it is possible to show that under $H_0$ p-values are uniform, that is inference is valid.
-\footnote{Depending on the exact regularity condition p-values might have to be multiplied by a factor of 2.
+Under regularity conditions spelled out in the paper, it is possible to show that under $H_0$ p-values are uniform, that is, inference is valid.\footnote{Depending on the exact regularity condition p-values might have to be multiplied by a factor of 2.
 The authors argue, however, that in their experience from simulations smaller p-values are typically appropriate.}
@@
 
-\subsection{Software Packages}
-For the baseline causal forest estimator I rely on the `CausalForest' implementation provided by the \href{https://econml.azurewebsites.net/}{EconML} Python package.
-This package implements a range of machine learning techniques for causal based on the statistical literature and making use of existing machine learning implementations in Python.
-For the generic ML part, I implemented most algorithms by hand since these only involve specifying linear regressions and sample splitting.
-Again, it is possible to flexibly use any of the machine learning estimators in this case, for example implemented in EconML or sklearn directly.
-For estimating DGPs using Wasserstein Generative Adversarial Networks (WGAN) I use the \href{https://github.com/gsbDBI/ds-wgan}{ds-wgan} implementation provided by the authors.
-Unfortunately, the code is not maintained and not compatible with recent versions of most basic Python libraries (e.g. NumPy 2.0).
-To avoid dependency conflicts with the other packages or manually debugging their code, I instead trained the generators in a Google Colab environment, which also comes at a significant speedup due to GPU availability.
-
-\subsection{Simulation Setup}
+\subsection{DGP and Parameters}
 \paragraph*{Designing a High-Dimensional DGP}
 A key question is how to design a DGP that is both ``realistic'', in the sense that the results are relevant to empirical applications, and still poses a high-dimensional problem.
-The typical approach taken in the literature, e.g.\ in~\cite{wager2018estimation} is to postulate a DGP that has convenient analytical features and is easy to simulate.
+The typical approach taken in the literature, e.g.\ in~\citet{wager2018estimation} is to postulate a DGP that has convenient analytical features and is easy to simulate.
 This simplifies the analysis but begs the question of relevance.
 Also, it leaves some degree of freedom to the researcher to choose a DGP that favourably displays the advantages of their proposed method.
-To address this issue, ``empirical'' methods for simulation design have been proposed, such as the WGAN approach in~\cite{athey2024wgan}.
+To address this issue, ``empirical'' methods for simulation design have been proposed, such as the approach in~\citet{athey2024wgan} based on Wasserstein Generative Adversarial Networks (WGANs).
 Without going into detail, the main idea is to train a \textit{generative} machine learning model, that can generate ``new'' observations by estimating the underlying DGP\@.
-The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in classifying observations as real or generated.
+The training/optimization procedure can be thought of as a two-player game in which a \textit{generator} and a \textit{discriminator} compete in generating and classifying observations as real or generated, respectively.
 
 While this approach can be usefully employed to design specific simulations given a dataset at hand for some questions, it also has some limitations, specifically with respect to the problem studied here.
-In particular, we are concerned with a high-dimensional, non-sparse structure of the CATE functional --- this is what the generic ML method is robust against.
-However, for a given DGP with high-dimensional structure it cannot be possible train a generator that preserves this structure.
+We are concerned with a high-dimensional, non-sparse structure of the CATE functional --- this is what the generic ML method is designed for.
+However, for a given DGP with high-dimensional structure it cannot be possible to train a generator that preserves this structure.
 Else, if we had an estimate of the DGP, we could simply construct a plug-in estimator of the CATE, but as argued before such an estimator cannot exist.
 
-For this reason I resort to the analytical DGP presented in~\cite{wager2018estimation} (WA in the following).
-To illustrate the difficulty with using a WGAN generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA DGP\@.
-To the best of my knowledge, such simulation results are not available in the original or any other paper.
-\footnote{They do, however, assess robustness of their simulation setup in several dimensions, including sampling variability, in Section 5.}
+For this reason I resort to the analytical DGP presented in~\citet{wager2018estimation} (WA in the following).
+To illustrate the difficulty with using a WGAN-generated DGP when the problem becomes high-dimensional, I also report results based on a generator that was trained on a draw from the WA DGP\@.
+To the best of my knowledge, such simulation results are not available in the original paper.\footnote{They do, however, assess robustness of their simulation setup in several dimensions, including sampling variability, in Section 5.}
 The following paragraphs explain the simulation setup in more detail.
@@
 \paragraph*{WA DGP}
-I use the third DGP discussed in~\cite{wager2018estimation} as my baseline DGP\@.
-Covariates are generated as $X~U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
-~\cite{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
+I use the third DGP discussed in~\citet{wager2018estimation} as my baseline DGP\@.
+Covariates are generated as $X\sim U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
+~\citet{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
 Instead, the CATE function is generated by
@@
 \end{equation*}
-That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
+That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise with respect to treatment effect estimation.
 The propensity score is kept constant at 0.5. I present results for various ambient dimensions $d$.
@@
 The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
-A similar question is analyzed in Figure 1 of~\cite{chernozhukov2023genml}.
+A similar question is analyzed in Figure 1 of~\citet{chernozhukov2023genml}.
 
 \paragraph*{WGAN Approximation}
-For generating the WGAN approximation to the DGP I follow the baseline approach described in ds-wgan package example.
+For generating the WGAN approximation to the DGP I follow the baseline approach described in the \textit{ds-wgan} package example.
 In particular, I train the generator and critic using the same hyperparameters.
 All fitting is performed twice, once for the distribution of $Y$ conditional on $X, D$ and once for the distribution of $X$ conditional on $D$.
-I train the model based on a sample of $N=100,000$ observations with dimension $d=10$ drawn from the baseline WA DGP as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
-I deliberately chose a large sample (at least larger than the $N=10,000$ observations used throughout simulations) to attenuate noise resulting from this original sample draw.
-To further address this it would be ideal to increase $N$ further or aggregate the results over different initial sample draws.
+I train the model based on a sample of $n=100,000$ observations with dimension $d=10$ drawn from the baseline WA DGP as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
+I deliberately chose a large sample (at least larger than the $n=10,000$ observations used throughout simulations) to attenuate noise resulting from this original sample draw.
+To further address this it would be ideal to increase $n$ further or aggregate the results over different initial sample draws.
 Given the training time of these models I found both approaches to be infeasible for the scope of this paper.
@@
 While we could compute a similar average conditional on $X$, if there are continuous covariates that would be a bad approximation to the true underlying CATE since we end up with individual treatment effects.
-If, on the other hand, we fit a model to the population we know from that even if the WGAN preserved the high-dimensional structure it would not be possible to consistently estimate it.
+An alternative would be to estimate the CATE function with one of the approaches outlined above.
+This should in principle be possible since the dimensionality of the problem is fixed while we can increase the population size arbitrarily.
+However, this obviously relies on the generator preserving the structure of the true CATE\@.
 For the analysis I compute MSE using two approaches: Once using the true CATE function of the WA DGP and once using an estimated CATE based on the generated population using a causal forest approach.
 The second approach in particular should result in understated MSE estimates: If the WGAN generator cannot approximate the high-dimensional structure, using an estimated CATE based on a generated population will understate the difficulty of the problem.
-The first approach is of course not available in practice since the true CATE is unknown.
+The first approach is of course not available in practice since the true CATE is unknown.\footnote{At the same time, an upward biased MSE should be expected if we cannot sufficiently reduce the noise when sampling from the original DGP and training the generator.}
+
+\subsection{Software Packages}
+For the baseline causal forest estimator I rely on the `CausalForest' implementation provided by the \href{https://econml.azurewebsites.net/}{EconML} Python package.
+This package implements a range of machine learning techniques for causal inference based on the statistical literature and makes use of existing machine learning implementations in Python.
+For the generic ML part, I implemented most algorithms myself since these only involve specifying linear regressions and sample splitting.
+Again, it is possible to flexibly use any of the machine learning estimators in this case, for example implemented in EconML or sklearn --- a popular machine leanring environment --- directly.
+For estimating DGPs using Wasserstein Generative Adversarial Networks (WGAN) I use the \href{https://github.com/gsbDBI/ds-wgan}{ds-wgan} implementation provided by the authors.
+Unfortunately, the code is not maintained and not compatible with recent versions of most basic Python libraries (e.g. NumPy 2.0).
+To avoid dependency conflicts with the other packages or manually debugging their code, I instead trained the generators in a Google Colab environment, which also comes at a significant speedup due to GPU availability.
 
 \section{Simulation Results}
-Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to~\cite{wager2018estimation}.
+Throughout the simulations I use $n=10,000$ observations and $40$ simulation repetitions, corresponding to~\citet{wager2018estimation}.
 This low number is chosen to make the problem computationally feasible.
 I use $100$ trees in all causal forest estimators and a sub-sampling share of $0.2$.
-These parameters are much smaller than in the original~\cite{wager2018estimation} to reduce the run-time of the simulation.
+These parameters are much smaller than in the original~\citet{wager2018estimation} to reduce the run-time of the simulation.
 I still find very similar results in the replication parts, as outlined below.
-I follow~\cite{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
+I follow~\citet{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
 Throughout, I construct confidence intervals with $95\%$ nominal coverage.
@@
 When analysing group average treatment effects (GATES), I define five groups based on the treatment effect quintiles predicted by the ML proxy.
-To approximate the true GATES induced by the proxy I estimate this on a large sample of size $N=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
+To approximate the true GATES induced by the proxy I estimate this on a large sample of size $n=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
 I use a main, auxiliary splitshare of $0.5$ throughout the analysis.
@@
 I report results for the causal forest approach in Table~\ref*{table:wa_rep_wgan}.
-First, the estimates for MSE --- using the true CATE function --- in column 2 are very close to the original results in~\cite{wager2018estimation}.
+First, the estimates for MSE --- using the true CATE function --- in column 2 are very close to the original results in~\citet{wager2018estimation}.
 I find a MSE for 0.02 for $d=10$ with an SD smaller than 0.01 while they find MSE estimates ranging from 0.02 to 0.03 for dimensions ranging from 2 to 8 and a similar SD\@.
@@
 The MSE is more than three times as large as the directly simulated DGP when using the true CATE to compute the MSE (\textit{MSE True}), but delivers a similar MSE estimate when using an estimated CATE based on a large population draw (\textit{MSE Pop}).
-Hence, using a WGAN generated DGP can in principle work well for simulation purposes when true CATE has a low-dimensional structure.
-Turning to coverage in the last columns, I find that coverage is slightly above nominal level for all dimensions, which is different to the original results reported in~\cite{wager2018estimation}, who report coverage levels far below nominal level (e.g.
+Hence, using a WGAN-generated DGP can in principle work well for simulation purposes when true CATE has a low-dimensional structure.
+Turning to coverage in the last columns, I find that coverage is slightly above nominal level for all dimensions, which is different to the original results reported in~\citet{wager2018estimation}, who report coverage levels far below nominal level (e.g.
 0.73 for $d=8$).
@@
 Lastly, Panel D reports result for a Placebo CATE which is zero everywhere. The MSE is small and coverage is above nominal level.
-Hence I find no indication for the claim made in connection with Figure 1 in~\cite{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.
+Hence I find no indication for the claim made in connection with Figure 1 in~\citet{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.
 \footnote{They, however, used a random forest directly (i.e.\ with a different target function and without modifications like honesty).
@@
             \small
-            \item \textbf{Notes}: This table reports simulation results for~\cite{wager2018estimation}.
-            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+            \item \textbf{Notes}: This table reports simulation results for~\citet{wager2018estimation}.
+            Results for $40$ simulation runs with number of observations $n=10,000$, number of trees $200$ and resampling share $0.2$.
             \textit{Dimension} refers to the number of covariates in the DGP\@.
@@
             Coverage for $95\%$ confidence intervals reported.
-            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
+            Panel A reports results for the third DGP in~\citet{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
             For details see the text.
@@
 Figure~\ref*{fig:gates} illustrates the performance of the GATES procedure for the four different DGPs\@.
-It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($N=1,000,000$) sample draw.
+It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($n=1,000,000$) sample draw.
 While quantitatively not precise, we can still see that generic ML procedure correctly estimates almost all GATES for all four DGPs.
@@
         \small
-        \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml}.
+        \item \textbf{Notes}: This table reports simulation results for~\citet{chernozhukov2023genml}.
         In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
-        Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+        Results for $40$ simulation runs with number of observations $n=10,000$, number of trees $200$ and resampling share $0.2$.
         \textit{Dimension} refers to the number of covariates in the DGP\@.
         MSE is calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
-        Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
+        Panel A reports results for the third DGP in~\citet{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
         For details see the text.
@@
             \small
-            \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml} studying coverage for the BLP and GATES parameters (five groups).
+            \item \textbf{Notes}: This table reports simulation results for~\citet{chernozhukov2023genml} studying coverage for the BLP and GATES parameters (five groups).
             In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
-            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
+            Results for $40$ simulation runs with number of observations $n=10,000$, number of trees $200$ and resampling share $0.2$.
             \textit{Dimension} refers to the number of covariates in the DGP\@.
@@
      Lower ends are displayed red, upper blue.
-     The dashed line indicates the true GATES using a large population draw. Sample size $N=10,000$ and $40$ simulation runs.
+     The dashed line indicates the true GATES using a large population draw. Sample size $n=10,000$ and $40$ simulation runs.
      }
@@
      For each of the four DGPs, it plots the distribution of each of the five GATE parameter estimates over the simulation runs.
-     The dashed line indicates the true GATES using a large population draw. Sample size $N=10,000$ and $40$ simulation runs.
+     The dashed line indicates the true GATES using a large population draw. Sample size $n=10,000$ and $40$ simulation runs.
      }
@@
      The dashed line indicates the true characteristics of the least and most affected group estimated from a large sample.
-     Sample size $N=10,000$ and $40$ simulation runs.
+     Sample size $n=10,000$ and $40$ simulation runs.
      }
@@
 While the approach taken above allows to control the dimension of the CATE, the covariate structure is of course highly stylized (no discrete covariates, all uniform, all independent).
-I think most promising would be a hybrid approach that combines a WGAN generated DGP for the covariates and treatment assignment, with an analytical treatment effect structure.
+I think most promising would be a hybrid approach that combines a WGAN-generated DGP for the covariates and treatment assignment, with an analytical treatment effect structure.
 This still poses the challenge, however, of coming up with a high-dimensional CATE function that makes sense economically.

### 2024-08-18 documents/paper.tex
Fix more typos.

@@
 I use $100$ trees in all causal forest estimators and a sub-sampling share of $0.2$.
-These parameters are much smaller than in the original~\citet{wager2018estimation} to reduce the run-time of the simulation.
+These parameters are much smaller than in the original simulations, again to reduce the run-time.
 I still find very similar results in the replication parts, as outlined below.
@@
 Throughout, I construct confidence intervals with $95\%$ nominal coverage.
-To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation of simulation runs.
+To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation over simulation runs.
 For generic machine learning simulations I use a standard random forest as the baseline learner for $b_0(Z) = \E[Y(0)|Z]$ and a causal forest with similar hyperparameters as above for estimating $s_0(Z)$.
 When analysing group average treatment effects (GATES), I define five groups based on the treatment effect quintiles predicted by the ML proxy.
-To approximate the true GATES induced by the proxy I estimate this on a large sample of size $n=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
-I use a main, auxiliary splitshare of $0.5$ throughout the analysis.
+To approximate the true GATES (as well as BLP and CLAN) induced by the proxy I estimate this on a large sample of size $n=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
+For the classification analysis (CLAN) I consider average covariate characteristics as the target. I use a main, auxiliary splitshare of $0.5$ and aggregation over 10 splits throughout the analysis.
 
@@
 I report results for the causal forest approach in Table~\ref*{table:wa_rep_wgan}.
-First, the estimates for MSE --- using the true CATE function --- in column 2 are very close to the original results in~\citet{wager2018estimation}.
+First, the estimates for MSE --- using the true CATE function for evaluation --- in column 2 are very close to the original results in~\citet{wager2018estimation}.
 I find a MSE for 0.02 for $d=10$ with an SD smaller than 0.01 while they find MSE estimates ranging from 0.02 to 0.03 for dimensions ranging from 2 to 8 and a similar SD\@.
@@
 This shows, that causal forests are able to perform well in a sparse setting where the subset of covariates that are pure noise is potentially large.
-The fourth row reports in Panel A reports simulation results based on an intermediate WGAN generator.
-The MSE is more than three times as large as the directly simulated DGP when using the true CATE to compute the MSE (\textit{MSE True}), but delivers a similar MSE estimate when using an estimated CATE based on a large population draw (\textit{MSE Pop}).
-Hence, using a WGAN-generated DGP can in principle work well for simulation purposes when true CATE has a low-dimensional structure.
+The fourth row of Panel A reports simulation results based on an intermediate WGAN generator.
+The MSE is more than three times as large compared to sampling directly from the DGP when using the true CATE to compute the MSE (\textit{MSE True}), but delivers a similar MSE estimate when using an estimated CATE based on a large population draw (\textit{MSE Pop}).
+Hence, using a WGAN-generated DGP can in principle work well for simulation purposes when the true CATE has a low-dimensional structure.
 Turning to coverage in the last columns, I find that coverage is slightly above nominal level for all dimensions, which is different to the original results reported in~\citet{wager2018estimation}, who report coverage levels far below nominal level (e.g.
@@
 
-Panels B and C report results for a CATE with higher dimension. As expected the performance of the Causal Forest significantly deteriorates.
+Panels B and C report results for a CATE with higher dimensionality. As expected, the performance of the Causal Forest significantly deteriorates.
 MSE significantly increases, in particular for Panel C where the CATE depends on 8 covariates.
-Even with only 4 covariates, coverage is sinificantly below the nominal level ranging from 0.83 to 0.76. For 8 covariates coverage is below 0.5 for all dimensions.
-The WGAN generator approach result in similar MSE estimates when using the estimated CATE to compute MSE (0.22 vs. 0.19 using the standard approach).
+Even with only 4 covariates, coverage is significantly below the nominal level ranging from 0.76 to 0.83. For 8 covariates coverage is below 0.5 for all dimensions.
+The WGAN generator approach results in similar MSE estimates when using the estimated CATE to compute MSE (0.22 vs. 0.19 using the standard approach).
 However, for the more complicated CATE in Panel C, the WGAN approach suggests an MSE that is far below the actual MSE using the standard approach (44.99 vs. 133.81).
@@
 For the remaining simulations using the generic machine learning approach I stick to the standard approach.
-Lastly, Panel D reports result for a Placebo CATE which is zero everywhere. The MSE is small and coverage is above nominal level.
-Hence I find no indication for the claim made in connection with Figure 1 in~\citet{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.
-\footnote{They, however, used a random forest directly (i.e.\ with a different target function and without modifications like honesty).
+Lastly, Panel D reports results for a Placebo CATE which is zero everywhere. The MSE is small and coverage is above nominal level.
+Hence I find no indication for the claim made in connection with Figure 1 in~\citet{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.\footnote{They, however, used a random forest directly (i.e.\ with a different target function and without modifications like honesty).
 Also, they probably reported out-of-the-box estimates which, since random forests are meant to address problems with more than one covariate, are prone to overfitting in this case.
 Using a random forest with cross-validation easily avoids overfitting.}
+
 \begin{table}
@@
 The results in Table~\ref*{table:generic_ml_cate} show that the MSE for estimating the true CATE is significantly worse compared to the causal forest approach.
-While the BLP is of course not linear in the original covaiates, it seems like the gain estimation precision might still be offset by the approximation.
+While the BLP is of course not linear in the original covariates, it seems like the gain in estimation precision might still be offset by the approximation.
 This is expected for DGPs with a low-dimensional CATE in which causal forests perform well (Panels A and B), but I would have expected a relatively better performance in Panel C.
@@
 Notably, $\beta_2$ has coverage of $0\%$ or close to it for almost all settings.
-To investigate this further, Figure~\ref*{fig:blp} plots the distribution of the lower and upper confidence interval points for both all four DGPs and both BLP parameters.
+To investigate this further, Figure~\ref*{fig:blp} plots the distribution of the lower and upper confidence interval points for both BLP parameters and all four DGPs.
 While endpoints are to the left and right of $\beta_1$ in most cases except the high-dimensional ones, $\beta_2$ is always over-estimated.
@@
 It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($n=1,000,000$) sample draw.
-While quantitatively not precise, we can still see that generic ML procedure correctly estimates almost all GATES for all four DGPs.
+While quantitatively not precise, we can still see that the generic ML procedure correctly estimates almost all GATES for all four DGPs.
 This is true even for the DGPs with increased heterogeneity dimension, as well as the placebo, where estimates are tightly centered around 0.
 
-Lastly, Figure~\ref*{fig:clans} shows estimates for the characteristics --- limited to covariates $x_1, \ldots, x_4$ --- of the least and most affected groups as predited by the ML proxy.
+Lastly, Figure~\ref*{fig:clans} shows estimates for the mean characteristics --- limited to covariates $x_1, \ldots, x_4$ --- of the least and most affected groups as predited by the ML proxy.
 If the CATE function does not depend on a covariate, the uniform distribution implies that the average characteristic for both groups is 0.5.
@@
 Eventually, this all boils down to the question, how high-dimensional and non-sparse our objects of interest really are.
-Lastly, when moving to low-dimensional summaries of the CATE functional, it would be useful to have decision-theoretic framework to evaluate.
+Lastly, when moving to low-dimensional summaries of the CATE functional, it would be useful to have a decision-theoretic framework to evaluate the performance of approaches with different targets, but for the same decision problem.
 This would allow to think more systematically about how to trade-off model approximations with statistical performance.
